### 2018-12-19 manuscript.tex
set up papaja manuscript.Rmd for reproducibility

\documentclass[man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Manuscript},
            pdfauthor={Sally E. Claridge, Daniel M. Charytonowicz, \& Adam A. Margolin},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[style=apa]{biblatex}

\addbibresource{manuscript.bib}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{Manuscript}
    \author{Sally E. Claridge\textsuperscript{1}, Daniel M.
Charytonowicz\textsuperscript{1}, \& Adam A.
Margolin\textsuperscript{1, 2}}
    \date{}

\shorttitle{Subtitle}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Genetics and Genomic Sciences, Icahn Institute for Data Science and Genomic Technology, Icahn School of Medicine at Mount Sinai, New York, NY\\\textsuperscript{2} Cancer Target Discovery and Development Network (Oregon Health and Science University, Portland, OR), National Cancer Institute, National Institutes of Health}
\keywords{keywords\newline\indent Word count: X}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers

\authornote{Add complete departmental affiliations for each
author here. Each new line herein must be indented, like this line.

Enter author note here.

Correspondence concerning this article should be addressed to Adam A.
Margolin, Icahn School of Medicine at Mount Sinai, Deparment of Genetics
and Genomic Sciences, One Gustave L. Levy Place, Box 1498, New York, NY
10029. E-mail:
\href{mailto:adam.margolin@mssm.edu}{\nolinkurl{adam.margolin@mssm.edu}}}

\abstract{
One or two sentences providing a \textbf{basic introduction} to the
field, comprehensible to a scientist in any discipline.

Two to three sentences of \textbf{more detailed background},
comprehensible to scientists in related disciplines.

One sentence clearly stating the \textbf{general problem} being
addressed by this particular study.

One sentence summarizing the main result (with the words ``\textbf{here
we show}'' or their equivalent).

Two or three sentences explaining what the \textbf{main result} reveals
in direct comparison to what was thought to be the case previously, or
how the main result adds to previous knowledge.

One or two sentences to put the results into a more \textbf{general
context}.

Two or three sentences to provide a \textbf{broader perspective},
readily comprehensible to a scientist in any discipline.

}

\begin{document}
\maketitle

\autocite{aksoy2017}

\section{Introduction}\label{introduction}

\section{Results}\label{results}

\section{Discussion}\label{discussion}

\newpage

\section{Methods}\label{methods}

We used R \autocite[Version 3.5.0;][]{R-base} and the R-packages
\emph{bsselectR} \autocite[Version 0.1.0;][]{R-bsselectR},
\emph{cowplot} \autocite[Version 0.9.3;][]{R-cowplot}, \emph{data.table}
\autocite[Version 1.11.4;][]{R-data.table}, \emph{dplyr}
\autocite[Version 0.7.8;][]{R-dplyr}, \emph{forcats} \autocite[Version
0.3.0;][]{R-forcats}, \emph{futile.logger} \autocite[Version
1.4.3;][]{R-futile.logger}, \emph{ggplot2} \autocite[Version
3.0.0;][]{R-ggplot2}, \emph{ggpubr} \autocite[Version
0.1.7.999;][]{R-ggpubr}, \emph{ggrepel} \autocite[Version
0.8.0;][]{R-ggrepel}, \emph{ggsignif} \autocite[Version
0.4.0;][]{R-ggsignif}, \emph{gplots} \autocite[Version
3.0.1;][]{R-gplots}, \emph{gridExtra} \autocite[Version
2.3;][]{R-gridExtra}, \emph{htmlwidgets} \autocite[Version
1.2;][]{R-htmlwidgets}, \emph{kableExtra} \autocite[Version
0.9.0;][]{R-kableExtra}, \emph{magrittr} \autocite[Version
1.5;][]{R-magrittr}, \emph{papaja} \autocite[Version
0.1.0.9842;][]{R-papaja}, \emph{plotly} \autocite[Version
4.8.0;][]{R-plotly}, \emph{purrr} \autocite[Version 0.2.5;][]{R-purrr},
\emph{readr} \autocite[Version 1.1.1;][]{R-readr}, \emph{reshape2}
\autocite[Version 1.4.3;][]{R-reshape2}, \emph{scales} \autocite[Version
1.0.0;][]{R-scales}, \emph{shiny} \autocite[Version 1.1.0;][]{R-shiny},
\emph{stringr} \autocite[Version 1.3.1;][]{R-stringr}, \emph{tibble}
\autocite[Version 1.4.2;][]{R-tibble}, \emph{tidyr} \autocite[Version
0.8.2;][]{R-tidyr}, \emph{tidyverse} \autocite[Version
1.2.1;][]{R-tidyverse}, and \emph{VennDiagram} \autocite[Version
1.6.20;][]{R-VennDiagram} for all our analyses.

\newpage

\section{References}\label{references}

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

Finally, the APA guidelines require a note at the start of the reference
section that explains what an asterisk means. This note can be added at
the end of the document as follows.

\hypertarget{refs}{}

\endgroup

\printbibliography

\end{document}