### 2018-12-19 manuscript_analysis.Rmd

---
title: "Manuscript"
author: "Sally Claridge"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: cerulean
    number_sections: true
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
    code_folding: hide
    pandoc_args: [
      "+RTS", "-K2048m",
      "-RTS"
    ]
---

Analysis of the Broad's Avana CRISPR (Broad Institute Cancer Dependency Map 2018, Meyers et al. 2017) and the Broad and Dana-Farber Cancer Institite's Achilles shRNA (MacFarland et al. 2018, Data Science 2018).

The Avana screen produced results using [CERES](https://depmap.org/ceres/) (Meyers et al. 2017) ([GitHub](https://github.com/cancerdatasci/ceres)), which generates gene dependency scores from sgRNA depletion scores from gene essentiality screens and eliminates bias arising from the effect of copy number variation on Cas9 DNA cleavage. The lower the CERES score, the higher the likelihood that the gene is essential in the associated cell line. Scores are scaled per cell line such that a score of 0 is the median effect of nonessential genes and -1 is the median effect of common core essential genes.

All annotation files (copy number, mutation status, and gene expression) (Consortium and Consortium 2015, Barretina et al. 2012) were downloaded from the [DepMap Data Portal](https://depmap.org/portal/download/).

# SET UP

--------------------------------------------------------------------------------

```{r setup, include=FALSE}
set.seed(1234)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# num_cores <- parallel::detectCores()
options(mc.cores = parallel::detectCores(), expressions = 5e5)
```

Edited `.Renviron` to avoid "Error: memory exhausted (limit reached?)" when melting the GE dataset, as per the suggestion from [StackOverflow](https://stackoverflow.com/questions/51295402/r-on-macos-error-vector-memory-exhausted-limit-reached)
```
cd ~
touch .Renviron
open .Renviron
# Added following line to .Renviron:
R_MAX_VSIZE=100Gb
```

```{r}
library(bsselectR)
library(cowplot)
theme_set(theme_light())
library(data.table)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(gplots)
library(gridExtra)
library(htmlwidgets)
library(kableExtra)
library(papaja)
library(plotly)
library(reshape2)
library(scales)
library(tidyverse)
library(VennDiagram)
# devtools::install_github("crsh/papaja")
```

# FUNCTIONS {.tabset .tabset-fade .tabset-pills}

--------------------------------------------------------------------------------

## Significance thresholding

```{r}
adj_signif <- function(df) {
  # Takes df from compare_means and creates a signif code column for adjusted p-vals
  df$p.signif <- ifelse(df$p.signif == "ns", NA, df$p.signif)
  df$p.signif.adj <- ifelse(df$p.adj <= 0.0001, "****",
                     ifelse(df$p.adj <= 0.001, "***",
                     ifelse(df$p.adj <= 0.01, "**",
                     ifelse(df$p.adj <= 0.05, "*", NA))))
  df$p.short <- formatC(df$p, format = "g", digits =  2)
  df$p.adj.short <- formatC(df$p.adj, format = "g", digits =  2)

  return(df)
}
```

## Wilcoxon tests by drug

```{r}
WilcoxonByDrug <- function(drug, dataset, data_name, grouping_var) {
  if(grouping_var == "lineage") {
    grouper <- c("Hugo_Symbol", "group_general_lineage_name")
  }
  else if(grouping_var == "gene") {
    grouper <- "Hugo_Symbol"
  }
  use_data <- filter(dataset, Drug == drug)
  sig <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = grouper, data = use_data, method = "wilcox.test", p.adjust.method = "BH")
  sig <- adj_signif(sig)
  sig <- sig[order(sig$p),]
  sig$Drug <- drug
  sig$Dataset <- toupper(data_name)
  return(sig)
}

# WilcoxonByDrugAllGenes <- function(melt_file, grid_file, data_name) {
#   druglist <- as.character(unique(melt_file$Drug))
#   sig_df <- NULL
#   for(i in 1:length(druglist)) {
#     print(paste0("Current drug in process: ", druglist[i], " (drug ", i, " of ", length(druglist), ")."))
#     flush.console()
#     melted <- filter(melt_file, Drug == druglist[i])
#     use_data <- merge(grid_file, melted, by = "accession_id", all.x = TRUE)
#     use_data <- merge(use_data, ccl_converter, by = "accession_id", all.x = TRUE)
#     use_data <- merge(use_data, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
#     use_data$Mutation_Status_Nonsilent <- ifelse(is.na(use_data$Mutation_Status_Nonsilent), "Wildtype", use_data$Mutation_Status_Nonsilent)
#     use_data$Hugo_Symbol <- factor(use_data$Hugo_Symbol)
#     use_data$AUC <- as.numeric(use_data$AUC)
#     use_data <- filter(use_data, !is.na(use_data$AUC))
#     use_data$Drug_Gene <- paste0(use_data$Drug, "_", use_data$Hugo_Symbol)
#     use_data_ptmuts <- filter(use_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
#     use_data_ptmuts$Keep_PtMut_Tests <- ifelse(use_data_ptmuts$Reference_Allele_Length == 1 | is.na(use_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(use_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
#     use_data_ptmuts <- filter(use_data_ptmuts, Keep_PtMut_Tests == TRUE)
#     
#     sig <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Hugo_Symbol", data = use_data_ptmuts, method = "wilcox.test", p.adjust.method = "BH")
#     sig <- adj_signif(sig)
#     sig <- sig[order(sig$p),]
#     sig <- filter(sig, p <= 0.05)
#     sig$Drug <- druglist[i]
#     sig$Dataset <- toupper(data_name)
#     sig_df <- rbind(sig_df, sig)
#   }
#   return(sig_df)
# }
```

## Pearson correlation by lineage
```{r}
PearsonByLineage <- function(lin, dataset, variable) {
  use_data <- filter(dataset, group_general_lineage_name == lin)
  if(variable == "GE") {
    use_data <- filter(use_data, !is.na(RPKM))
    use_data_summ <- use_data %>% count(Drug_Gene)
    use_data_summ <- filter(use_data_summ, n >= 3) # cor.test requires at least 3 non-NA data points
    use_data_filt <- filter(use_data, Drug_Gene %in% as.character(unique(use_data_summ$Drug_Gene)))
    if(nrow(use_data_filt) > 0) {
      pearson_res <- use_data_filt %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs")$estimate, Pearson_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs")$p.value)
      pearson_res <- merge(g2p_druggene_tally, pearson_res, all.y = TRUE)
      pearson_res$group_general_lineage_name <- lin
      pearson_res$Pearson_GE_Abs <- abs(pearson_res$Pearson_Corr_GE)
      pearson_res$Pearson_GE_Sign <- ifelse(sign(pearson_res$Pearson_Corr_GE) == 1, "+",
                                              ifelse(sign(pearson_res$Pearson_Corr_GE) == -1, "-", "0"))
      return(pearson_res)
    }
  }
  else if(variable == "CN") {
    use_data <- filter(use_data, !is.na(CN))
    use_data_summ <- use_data %>% count(Drug_Gene)
    use_data_summ <- filter(use_data_summ, n >= 3) # cor.test requires at least 3 non-NA data points
    use_data_filt <- filter(use_data, Drug_Gene %in% as.character(unique(use_data_summ$Drug_Gene)))
    if(nrow(use_data_filt) > 0) {
      pearson_res <- use_data_filt %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs")$estimate, Pearson_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs")$p.value)
      pearson_res <- merge(g2p_druggene_tally, pearson_res, all.y = TRUE)
      pearson_res$group_general_lineage_name <- lin
      pearson_res$Pearson_CN_Abs <- abs(pearson_res$Pearson_Corr_CN)
      pearson_res$Pearson_CN_Sign <- ifelse(sign(pearson_res$Pearson_Corr_CN) == 1, "+",
                                              ifelse(sign(pearson_res$Pearson_Corr_CN) == -1, "-", "0"))
      return(pearson_res)
    }
  }
}
```

## Save widget to new directory
```{r}
saveWidgetFix <- function (widget , file, ...) {
  ## A wrapper to saveWidget which compensates for arguable BUG in
  ## saveWidget which requires `file` to be in current working
  ## directory.
  wd <- getwd()
  on.exit(setwd(wd))
  outDir <- dirname(file)
  file <- basename(file)
  setwd(outDir);
  saveWidget(widget, file = file,...)
}
```

## Make lineage mut plots

```{r}
MutPlotsByLineage <- function(lineage_name) {

  use_data <- filter(dgl_mut_grid_summ, group_general_lineage_name == lineage_name)
  use_data$Dataset <- factor(use_data$Dataset, levels = c("CCLE", "CTRP", "GDSC"))
  ticker <- 0

  use_data_ccle <- filter(use_data, Dataset == "CCLE")
  use_data_ccle_summ <- use_data_ccle %>% count(InG2P)
  ccle_lin_p_ttest <- data.frame(statistic = NA,
                                 parameter =  NA,
                                 p.value = NA)
  if(nrow(use_data_ccle_summ) <= 1) {
    ccle_lin_p_ttest_res <-  NA
    ticker <- ticker + 1
    }
  else if(use_data_ccle_summ[use_data_ccle_summ$InG2P == "Yes",]$nn < 2) { ccle_lin_p_ttest_res <-  NA }
  else if(use_data_ccle_summ[use_data_ccle_summ$InG2P == "Yes",]$nn >= 2) {
    ccle_lin_p_ttest <- t.test(use_data_ccle$p ~ use_data_ccle$InG2P)
    ccle_lin_p_ttest_res <- paste0("t(", trunc(ccle_lin_p_ttest$parameter), ") = ", round(ccle_lin_p_ttest$statistic, 2), ", p = ", round(ccle_lin_p_ttest$p.value, 3))
  }

  use_data_ctrp <- filter(use_data, Dataset == "CTRP")
  use_data_ctrp_summ <- use_data_ctrp %>% count(InG2P)
  ctrp_lin_p_ttest <- data.frame(statistic = NA,
                                 parameter =  NA,
                                 p.value = NA)
  if(nrow(use_data_ctrp_summ) <= 1) {
    ctrp_lin_p_ttest_res <-  NA
    ticker <- ticker + 1
    }
  else if(use_data_ctrp_summ[use_data_ctrp_summ$InG2P == "Yes",]$nn < 2) { ctrp_lin_p_ttest_res <-  NA }
  else if(use_data_ctrp_summ[use_data_ctrp_summ$InG2P == "Yes",]$nn >= 2) {
    ctrp_lin_p_ttest <- t.test(use_data_ctrp$p ~ use_data_ctrp$InG2P)
    ctrp_lin_p_ttest_res <- paste0("t(", trunc(ctrp_lin_p_ttest$parameter), ") = ", round(ctrp_lin_p_ttest$statistic, 2), ", p = ", round(ctrp_lin_p_ttest$p.value, 3))
  }

  use_data_gdsc <- filter(use_data, Dataset == "GDSC")
  use_data_gdsc_summ <- use_data_gdsc %>% count(InG2P)
  gdsc_lin_p_ttest <- data.frame(statistic = NA,
                                 parameter =  NA,
                                 p.value = NA)
  if(nrow(use_data_gdsc_summ) <= 1) {
    gdsc_lin_p_ttest_res <-  NA
    ticker <- ticker + 1
    }
  else if(use_data_gdsc_summ[use_data_gdsc_summ$InG2P == "Yes",]$nn < 2) { gdsc_lin_p_ttest_res <-  NA }
  else if(use_data_gdsc_summ[use_data_gdsc_summ$InG2P == "Yes",]$nn >= 2) {
    gdsc_lin_p_ttest <- t.test(use_data_gdsc$p ~ use_data_gdsc$InG2P)
    gdsc_lin_p_ttest_res <- paste0("t(", trunc(gdsc_lin_p_ttest$parameter), ") = ", round(gdsc_lin_p_ttest$statistic, 2), ", p = ", round(gdsc_lin_p_ttest$p.value, 3))
  }

  y_value <- ifelse(ticker == 3, 10, max(use_data$n) + max(use_data$n) * 0.1)

  summ_text <- data.frame(Dataset = c("CCLE", "CTRP", "GDSC"),
                          Test_Results_p = c(ccle_lin_p_ttest_res, ctrp_lin_p_ttest_res, gdsc_lin_p_ttest_res),
                          t_statistic = c(round(ccle_lin_p_ttest$statistic, 3), round(ctrp_lin_p_ttest$statistic, 3), round(gdsc_lin_p_ttest$statistic, 3)),
                          df = c(trunc(ccle_lin_p_ttest$parameter), trunc(ctrp_lin_p_ttest$parameter), trunc(gdsc_lin_p_ttest$parameter)),
                          p_value = c(round(ccle_lin_p_ttest$p.value, 3), round(ctrp_lin_p_ttest$p.value, 3), round(gdsc_lin_p_ttest$p.value, 3)),
                          group_general_lineage_name = lineage_name)
  plot_point <- ggplot(data = use_data) +
    facet_wrap(~ Dataset, drop = FALSE) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, y_value)) +
    scale_x_continuous(breaks = seq(0, 1, by = 0.1), labels = seq(0, 1, by = 0.1)) +
    scale_y_continuous(breaks = pretty_breaks()) +
    geom_point(mapping = aes(label = Drug_Gene, color = InG2P, x = p, y = n), alpha = 0.5) +
    scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
    geom_text_repel(data = subset(use_data, InG2P == "Yes"), mapping = aes(x = p, y = n, label = Drug_Gene), size = 3, segment.size = 0.2, segment.color = "grey50") +
    geom_text(data = summ_text, mapping = aes(x = 0.5, y = y_value, label = Test_Results_p), nudge_y = 0.1) +
    labs(x = "Drug-gene association p-value between mutation groups", y = "Number of level A G2P associations") +
    theme(legend.position = "none", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank())
  plot_box <- ggplot(data = use_data) +
    facet_wrap(~ Dataset, drop = FALSE) +
    coord_cartesian(ylim = c(0, 1)) +
    scale_y_continuous(breaks = seq(0, 1, by = 0.1), labels = seq(0, 1, by = 0.1)) +
    geom_violin(mapping = aes(color = InG2P, y = p, x = InG2P), alpha = 0.5) +
    geom_boxplot(mapping = aes(color = InG2P, y = p, x = InG2P), alpha = 0.75, width = 0.25) +
    scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
    coord_flip() +
    labs(x = "In G2P?", title = toupper(lineage_name)) +
    theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))
  plot <- subplot(plot_box, plot_point, nrows = 2, titleX = TRUE, titleY = TRUE) %>% layout(title = toupper(lineage_name), margin = list(t = 75, l = 75), xaxis5 = list(title = "Gene-drug association p-value between mutation groups"), yaxis = list(title = "Level A G2P\nassociation", domain = c(0.8, 1)), yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))

  plot_point <- ggplot_gtable(ggplot_build(plot_point))
  plot_box <- ggplot_gtable(ggplot_build(plot_box))
  p_maxWidth <-  grid::unit.pmax(plot_point$widths, plot_box$widths)
  plot_point$widths <- as.list(p_maxWidth)
  plot_box$widths <- as.list(p_maxWidth)

  mut_plot <- cowplot::plot_grid(plot_box,
                     plot_point,
                     nrow = 2,
                     rel_heights = c(1, 4))

  saveWidgetFix(widget = plot, file = paste0("./plots/manuscript/plotlys_mut_by_lineage_grid/", lineage_name, "_mut_plotly_grid.html"))
  ggsave(paste0("./plots/manuscript/pngs_mut_by_lineage_grid/", lineage_name, "_mut_plot_grid.png"), mut_plot, device = "png", dpi = 450, width = 12, height = 6, units =   "in")

  return(summ_text)
}
```

## Make lineage GE/CN plots

```{r}
PearsonPlotsByLineage <- function(lineage) {
  print(paste0("Lineage: ", toupper(lineage)))
  use_data <- filter(dgl_signif_pearson_g2p_grid, group_general_lineage_name == lineage)
  use_data$Dataset <- factor(use_data$Dataset, levels = c("CCLE", "CTRP", "GDSC"))
  use_data$Pearson_GE_Label <- paste0(round(use_data$Pearson_Pval_GE, 3), "\nDrug_Gene: ", use_data$Drug_Gene)
  use_data$Pearson_CN_Label <- paste0(round(use_data$Pearson_Pval_CN, 3), "\nDrug_Gene: ", use_data$Drug_Gene)
  ticker_ge <- 0
  ticker_cn <- 0
  ticker <- 0

  use_data_ccle <- filter(use_data, Dataset == "CCLE")
  use_data_ccle_ge_summ <- use_data_ccle[!is.na(use_data_ccle$Pearson_GE_Abs),] %>% count(InG2P)
  use_data_ccle_cn_summ <- use_data_ccle[!is.na(use_data_ccle$Pearson_CN_Abs),] %>% count(InG2P)
  ccle_ge_ttest <- data.frame(statistic = NA, parameter = NA, p.value = NA)
  ccle_cn_ttest <- data.frame(statistic = NA, parameter = NA, p.value = NA)
  if(nrow(use_data_ccle_ge_summ) <= 1) {
    ccle_ge_res <-  NA
    ticker_ge <- ticker_ge + 1
  }
  else if(use_data_ccle_ge_summ[use_data_ccle_ge_summ$InG2P == "Yes",]$nn < 2) {
    ccle_ge_res <-  NA
  }
  else if(use_data_ccle_ge_summ[use_data_ccle_ge_summ$InG2P == "Yes",]$nn >= 2) {
    ccle_ge_ttest <- t.test(use_data_ccle$Pearson_GE_Abs ~ use_data_ccle$InG2P)
    ccle_ge_res <- paste0("t(", trunc(ccle_ge_ttest$parameter), ") = ", round(ccle_ge_ttest$statistic, 2), ", p = ", round(ccle_ge_ttest$p.value, 3))
  }
  if(nrow(use_data_ccle_cn_summ) <= 1) {
    ccle_cn_res <-  NA
    ticker_cn <- ticker_cn + 1
  }
  else if(use_data_ccle_cn_summ[use_data_ccle_cn_summ$InG2P == "Yes",]$nn < 2) {
    ccle_cn_res <-  NA
  }
  else if(use_data_ccle_cn_summ[use_data_ccle_cn_summ$InG2P == "Yes",]$nn >= 2) {
    ccle_cn_ttest <- t.test(use_data_ccle$Pearson_CN_Abs ~ use_data_ccle$InG2P)
    ccle_cn_res <- paste0("t(", trunc(ccle_cn_ttest$parameter), ") = ", round(ccle_cn_ttest$statistic, 2), ", p = ", round(ccle_cn_ttest$p.value, 3))
  }
  print("CCLE testing: DONE")

  use_data_ctrp <- filter(use_data, Dataset == "CTRP")
  use_data_ctrp_ge_summ <- use_data_ctrp[!is.na(use_data_ctrp$Pearson_GE_Abs),] %>% count(InG2P)
  use_data_ctrp_cn_summ <- use_data_ctrp[!is.na(use_data_ctrp$Pearson_CN_Abs),] %>% count(InG2P)
  ctrp_ge_ttest <- data.frame(statistic = NA, parameter = NA, p.value = NA)
  ctrp_cn_ttest <- data.frame(statistic = NA, parameter = NA, p.value = NA)
  if(nrow(use_data_ctrp_ge_summ) <= 1) {
    ctrp_ge_res <-  NA
    ticker_ge <- ticker_ge + 1
  }
  else if(use_data_ctrp_ge_summ[use_data_ctrp_ge_summ$InG2P == "Yes",]$nn < 2) {
    ctrp_ge_res <-  NA
  }
  else if(use_data_ctrp_ge_summ[use_data_ctrp_ge_summ$InG2P == "Yes",]$nn >= 2) {
    ctrp_ge_ttest <- t.test(use_data_ctrp$Pearson_GE_Abs ~ use_data_ctrp$InG2P)
    ctrp_ge_res <- paste0("t(", trunc(ctrp_ge_ttest$parameter), ") = ", round(ctrp_ge_ttest$statistic, 2), ", p = ", round(ctrp_ge_ttest$p.value, 3))
  }
  if(nrow(use_data_ctrp_cn_summ) <= 1) {
    ctrp_cn_res <-  NA
    ticker_cn <- ticker_cn + 1
  }
  else if(use_data_ctrp_cn_summ[use_data_ctrp_cn_summ$InG2P == "Yes",]$nn < 2) {
    ctrp_cn_res <-  NA
  }
  else if(use_data_ctrp_cn_summ[use_data_ctrp_cn_summ$InG2P == "Yes",]$nn >= 2) {
    ctrp_cn_ttest <- t.test(use_data_ctrp$Pearson_CN_Abs ~ use_data_ctrp$InG2P)
    ctrp_cn_res <- paste0("t(", trunc(ctrp_cn_ttest$parameter), ") = ", round(ctrp_cn_ttest$statistic, 2), ", p = ", round(ctrp_cn_ttest$p.value, 3))
  }
  print("CTRP testing: DONE")

  use_data_gdsc <- filter(use_data, Dataset == "GDSC")
  use_data_gdsc_ge_summ <- use_data_gdsc[!is.na(use_data_gdsc$Pearson_GE_Abs),] %>% count(InG2P)
  use_data_gdsc_cn_summ <- use_data_gdsc[!is.na(use_data_gdsc$Pearson_CN_Abs),] %>% count(InG2P)
  gdsc_ge_ttest <- data.frame(statistic = NA, parameter = NA, p.value = NA)
  gdsc_cn_ttest <- data.frame(statistic = NA, parameter = NA, p.value = NA)
  if(nrow(use_data_gdsc_ge_summ) <= 1) {
    gdsc_ge_res <-  NA
    ticker_ge <- ticker_ge + 1
  }
  else if(use_data_gdsc_ge_summ[use_data_gdsc_ge_summ$InG2P == "Yes",]$nn < 2) {
    gdsc_ge_res <-  NA
  }
  else if(use_data_gdsc_ge_summ[use_data_gdsc_ge_summ$InG2P == "Yes",]$nn >= 2) {
    gdsc_ge_ttest <- t.test(use_data_gdsc$Pearson_GE_Abs ~ use_data_gdsc$InG2P)
    gdsc_ge_res <- paste0("t(", trunc(gdsc_ge_ttest$parameter), ") = ", round(gdsc_ge_ttest$statistic, 2), ", p = ", round(gdsc_ge_ttest$p.value, 3))
  }
  if(nrow(use_data_gdsc_cn_summ) <= 1) {
    gdsc_cn_res <-  NA
    ticker_cn <- ticker_cn + 1
  }
  else if(use_data_gdsc_cn_summ[use_data_gdsc_cn_summ$InG2P == "Yes",]$nn < 2) {
    gdsc_cn_res <-  NA
  }
  else if(use_data_gdsc_cn_summ[use_data_gdsc_cn_summ$InG2P == "Yes",]$nn >= 2) {
    gdsc_cn_ttest <- t.test(use_data_gdsc$Pearson_CN_Abs ~ use_data_gdsc$InG2P)
    gdsc_cn_res <- paste0("t(", trunc(gdsc_cn_ttest$parameter), ") = ", round(gdsc_cn_ttest$statistic, 2), ", p = ", round(gdsc_cn_ttest$p.value, 3))
  }
  print("GDSC testing: DONE")

  y_value_ge <- ifelse(ticker_ge == 3, 10, max(use_data$n) + max(use_data$n) * 0.1)
  y_value_cn <- ifelse(ticker_cn == 3, 10, max(use_data$n) + max(use_data$n) * 0.1)

  summ_text <- data.frame(Dataset = c("CCLE", "CTRP", "GDSC"),
                          Test_Results_GE = c(ccle_ge_res, ctrp_ge_res, gdsc_ge_res),
                          t_statistic_GE = c(round(ccle_ge_ttest$statistic, 3), round(ctrp_ge_ttest$statistic, 3), round(gdsc_ge_ttest$statistic, 3)),
                          df_GE = c(trunc(ccle_ge_ttest$parameter), trunc(ctrp_ge_ttest$parameter), trunc(gdsc_ge_ttest$parameter)),
                          p_value_GE = c(round(ccle_ge_ttest$p.value, 3), round(ctrp_ge_ttest$p.value, 3), round(gdsc_ge_ttest$p.value, 3)),
                          Test_Results_CN = c(ccle_cn_res, ctrp_cn_res, gdsc_cn_res),
                          t_statistic_CN = c(round(ccle_cn_ttest$statistic, 3), round(ctrp_cn_ttest$statistic, 3), round(gdsc_cn_ttest$statistic, 3)),
                          df_CN = c(trunc(ccle_cn_ttest$parameter), trunc(ctrp_cn_ttest$parameter), trunc(gdsc_cn_ttest$parameter)),
                          p_value_CN = c(round(ccle_cn_ttest$p.value, 3), round(ctrp_cn_ttest$p.value, 3), round(gdsc_cn_ttest$p.value, 3)),
                          group_general_lineage_name = lineage)

  ## Gene expression
  if(sum(!is.na(use_data$Pearson_GE_Abs)) > 0) {
    plot_ge_point <- ggplot(data = use_data) +
      facet_wrap(~ Dataset, drop = FALSE) +
      scale_y_continuous(breaks = pretty_breaks()) +
      geom_point(mapping = aes(label = Pearson_GE_Label, color = InG2P, fill = InG2P, shape = Pearson_GE_Sign, x = Pearson_GE_Abs, y = n), alpha = 0.5) +
      scale_shape_manual(values = c(25, 24), labels = c("Negative", "Positive")) +
      scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
      scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
      geom_text_repel(data = subset(use_data, InG2P == "Yes" & n > 10), aes(label = Drug_Gene, x = Pearson_GE_Abs, y = n), size = 3, segment.size = 0.2, segment.color = "grey50") +
      geom_text(data = summ_text, mapping = aes(x = 0.5, y = y_value_ge, label = Test_Results_GE), nudge_y = 1) +
      labs(y = "Number of level A G2P associations", x = "Pearson correlation coefficient between drug-gene AUC and gene expression") +
      theme(legend.position = "none", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank())
    plot_ge_box <- ggplot(data = use_data) +
      facet_wrap(~ Dataset, drop = FALSE) +
      geom_violin(mapping = aes(color = InG2P, y = Pearson_GE_Abs, x = InG2P), alpha = 0.5) +
      geom_boxplot(mapping = aes(color = InG2P, y = Pearson_GE_Abs, x = InG2P), alpha = 0.5, width = 0.25) +
      scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
      coord_flip() +
      labs(x = "In G2P?", title = toupper(lineage)) +
      theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

    plot_ge <- subplot(plot_ge_box, plot_ge_point, nrows = 2, titleX = TRUE, titleY = TRUE) %>%
               layout(title = lineage, margin = list(t = 75, l = 75),
                      xaxis5 = list(title = "Pearson correlation coefficient between drug-gene AUC and gene expression"),
                      yaxis = list(title = "In G2P?", domain = c(0.8, 1)),
                      yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))

    png_ge_point <- ggplot_gtable(ggplot_build(plot_ge_point))
    png_ge_box <- ggplot_gtable(ggplot_build(plot_ge_box))
    p_maxWidth <-  grid::unit.pmax(png_ge_point$widths, png_ge_box$widths)
    png_ge_point$widths <- as.list(p_maxWidth)
    png_ge_box$widths <- as.list(p_maxWidth)

    png_ge <- cowplot::plot_grid(png_ge_box, png_ge_point, nrow = 2, rel_heights = c(1, 4))
    png_ge

    saveWidgetFix(widget = plot_ge, file = paste0("./plots/manuscript/plotlys_ge_by_lineage_grid/", lineage, "_ge_plotly_grid.html"))
    ggsave(paste0("./plots/manuscript/pngs_ge_by_lineage_grid/", lineage, "_ge_plot_grid.png"), png_ge, device = "png", dpi = 450, width = 12, height = 6, units = "in")
    print("Gene expression plots: DONE")
  }
  else if(sum(!is.na(use_data$Pearson_GE_Abs)) == 0) { print("Gene expression plots: None") }

  ## Copy number
  if(sum(!is.na(use_data$Pearson_CN_Abs)) > 0) {
    plot_cn_point <- ggplot(data = use_data) +
      facet_wrap(~ Dataset, drop = FALSE) +
      scale_y_continuous(breaks = pretty_breaks()) +
      geom_point(mapping = aes(label = Pearson_CN_Label, color = InG2P, fill = InG2P, shape = Pearson_CN_Sign, x = Pearson_CN_Abs, y = n), alpha = 0.5, position = position_jitter(width = 0, height = 0.15)) +
      scale_shape_manual(values = c(25, 24), labels = c("Negative", "Positive")) +
      scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
      scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
      geom_text_repel(data = subset(use_data, InG2P == "Yes"  & n > 10), aes(label = Drug_Gene, x = Pearson_CN_Abs, y = n), size = 3, segment.size = 0.2, segment.color = "grey50") +
      geom_text(data = summ_text, mapping = aes(x = 0.5, y = y_value_cn, label = Test_Results_CN), nudge_y = 1) +
      labs(y = "Number of level A G2P associations", x = "Pearson correlation coefficient between drug-gene AUC and copy number") +
      theme(legend.position = "none", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank())
    plot_cn_box <- ggplot(data = use_data) +
      facet_wrap(~ Dataset, drop = FALSE) +
      geom_violin(mapping = aes(color = InG2P, y = Pearson_CN_Abs, x = InG2P), alpha = 0.5) +
      geom_boxplot(mapping = aes(color = InG2P, y = Pearson_CN_Abs, x = InG2P), alpha = 0.5, width = 0.25) +
      scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
      coord_flip() +
      labs(x = "In G2P?", title = toupper(lineage)) +
      theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

    plot_cn <- subplot(plot_cn_box, plot_cn_point, nrows = 2, titleX = TRUE, titleY = TRUE) %>%
               layout(title = lineage, margin = list(t = 75, l = 75),
                      xaxis5 = list(title = "Pearson correlation coefficient between drug-gene AUC and copy number"),
                      yaxis = list(title = "In G2P?", domain = c(0.8, 1)),
                      yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))

    png_cn_point <- ggplot_gtable(ggplot_build(plot_cn_point))
    png_cn_box <- ggplot_gtable(ggplot_build(plot_cn_box))
    p_maxWidth <-  grid::unit.pmax(png_cn_point$widths, png_cn_box$widths)
    png_cn_point$widths <- as.list(p_maxWidth)
    png_cn_box$widths <- as.list(p_maxWidth)

    png_cn <- cowplot::plot_grid(png_cn_box, png_cn_point, nrow = 2, rel_heights = c(1, 4))
    png_cn

    saveWidgetFix(widget = plot_cn, file = paste0("./plots/manuscript/plotlys_cn_by_lineage_grid/", lineage, "_cn_plotly_grid.html"))
    ggsave(paste0("./plots/manuscript/pngs_cn_by_lineage_grid/", lineage, "_cn_plot_grid.png"), png_cn, device = "png", dpi = 450, width = 12, height = 6, units = "in")
    print("Copy number plots: DONE")
  }
  else if(sum(!is.na(use_data$Pearson_CN_Abs)) == 0) { print("Copy number plots: None") }

  return(summ_text)
}
```

## Make per-drug boxplots
```{r}
makeDrugBoxplots <- function(drug, dataset) {
  if(dataset == "CCLE" | dataset == "ccle") {
    use_data <- filter(ccle_data_ptmuts, Drug == drug)
    use_color <- as.character(use_data$Color_Nonsilent)
    names(use_color) <- use_data$Mutation_Status_Nonsilent
    sig <- ccle_signif_g2p_ByDrug[[drug]]
    label_text <- data.frame(p.signif = sig$p.signif, p.signif.adj = sig$p.signif.adj, Hugo_Symbol = sig$Hugo_Symbol)
    label_text$Hugo_Symbol <- factor(label_text$Hugo_Symbol, levels = g2p_genes)
  }
  else if(dataset == "CTRP" | dataset == "ctrp") {
    use_data <- filter(ctrp_data_ptmuts, Drug == drug)
    use_color <- as.character(use_data$Color_Nonsilent)
    names(use_color) <- use_data$Mutation_Status_Nonsilent
    sig <- ctrp_signif_g2p_ByDrug[[drug]]
    label_text <- data.frame(p.signif = sig$p.signif, p.signif.adj = sig$p.signif.adj, Hugo_Symbol = sig$Hugo_Symbol)
    label_text$Hugo_Symbol <- factor(label_text$Hugo_Symbol, levels = g2p_genes)
  }
  else if(dataset == "GDSC" | dataset == "gdsc") {
    use_data <- filter(gdsc_data_ptmuts, Drug == drug)
    use_color <- as.character(use_data$Color_Nonsilent)
    names(use_color) <- use_data$Mutation_Status_Nonsilent
    sig <- gdsc_signif_g2p_ByDrug[[drug]]
    label_text <- data.frame(p.signif = sig$p.signif, p.signif.adj = sig$p.signif.adj, Hugo_Symbol = sig$Hugo_Symbol)
    label_text$Hugo_Symbol <- factor(label_text$Hugo_Symbol, levels = g2p_genes)
  }
  else { return("Error: Invalid dataset.") }

  plot <- ggplot(data = use_data, aes(x = Mutation_Status_Nonsilent, y = AUC)) +
    facet_wrap(~ Hugo_Symbol, drop = FALSE, nrow = 1) +
    geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85), outlier.shape = 3, outlier.size = 0.5) +
    scale_fill_manual(values = use_color) +
    guides(color = FALSE) +
    geom_text(data = label_text, mapping = aes(x = 1.5, y = max(use_data$AUC), label = p.signif), nudge_y = 0.2) +
    theme(strip.text.x = element_text(size = 10, angle = 45), legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
    labs(fill = "Mutation Status", y = "AUC", x = "Mutation Status", title = paste0(toupper(dataset), ": ", drug))
  return(plot)
}
```

# DATA MANAGEMENT {.tabset .tabset-fade .tabset-pills}

--------------------------------------------------------------------------------

## D. Charytonowicz cell line converter

This comprehensive cancer cell line information curated by Daniel Charytonowicz.
```{r}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
lineage_converter <- unique(ccl_converter[, c("grouped_general_doid", "group_general_lineage_name")])
colnames(ccl_converter)[colnames(ccl_converter) == "Broad_ID"] <- "DepMap_ID"
```

## DepMap cell line metadata

```{r}
ccl_info <- read.delim("./data_munging/18Q4/DepMap-2018q4-celllines_20181115.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_info$Primary.Disease <- gsub("\\\\", "", ccl_info$Primary.Disease)
ccl_info$Primary.Disease <- gsub("Ewings", "Ewing's", ccl_info$Primary.Disease)
# From figshare
crispr_meta <- read.delim("./data_munging/18Q4/sample_info_18Q4_crispr_20181115.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[7] <- "CCLE_Name"
```

## CCLE genomic annotations

### Mutation status (Mut)

For mutation calling, get paired gene name and cell line fields in a data frame. For this analysis, we don't care about how many mutations there are per gene or what type of mutations there are, so I didn't save more information. I took unique gene-cell line combinations since we only cared about mutation presence/absence. Add a `Mutation_Status` column denoting all entries in MAF files as mutations present in the associated cell lines.

Merged mutation calls (coding region, germline filtered):
```{r, eval=FALSE}
maf_raw <- read.delim("./data_munging/18Q4/depmap_18Q4_mutation_calls_20181114.csv.gz", header = TRUE, sep = ",", row.names = 1)
colnames(maf_raw)[colnames(maf_raw) == "Tumor_Sample_Barcode"] <- "CCLE_Name"
# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele_Length <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1_Length <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- ifelse(maf_df$Variant_Classification == "Silent", "Wildtype", "Mutant")
maf_df <- unique(maf_df)
saveRDS(maf_df, "./data_munging/rds/maf_df_18Q4.rds", compress = "xz")
```

### Copy number (CN)

DepMap WES CN Data:
```{r, eval=FALSE}
cn <- read.delim("./data_munging/18Q4/public_18Q4_gene_cn_20181115.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
# Remove Entrez gene IDs from colnames
# Format: Hugo_Symbol (Entrez ID)
colnames(cn) <- gsub(" .*", "", colnames(cn))
# Melt
cn_melt <- melt(data = cn, id.vars = "DepMap_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Hugo_Symbol", value.name = "CN")
saveRDS(cn_melt, "./data_munging/rds/cn_melt_18Q4.rds", compress = "xz")
```

### Gene expression (GE)

18Q3, Gene expression data (Reads Per Kilobase of transcript, per Million mapped reads, RPKM).
```{r, eval=FALSE}
ge_18Q3 <- read.delim("./../htscreens_giant_files/CCLE_DepMap_18q3_RNAseq_RPKM_20180718.gct.gz", skip = 2, header = TRUE, sep = "\t", check.names = FALSE)
# Edit columns
ge_18Q3$Name <- NULL
colnames(ge_18Q3)[1] <- "Hugo_Symbol"
# Melt
ge_melt_18Q3 <- melt(data = ge_18Q3, id.vars = "Hugo_Symbol", measure.vars = colnames(ge_18Q3[2:ncol(ge_18Q3)]), value.name = "RPKM")
## Split variable column
ge_melt_18Q3 <- with(ge_melt_18Q3, cbind(Hugo_Symbol, colsplit(variable, pattern = " ", names = c("CCLE_Name", "DepMap_ID")), RPKM))
## Remove parentheses around Broad IDs
ge_melt_18Q3$DepMap_ID <- gsub("\\(|\\)", "", ge_melt_18Q3$DepMap_ID)
saveRDS(ge_melt_18Q3, "./../htscreens_giant_files/ge_melt_18Q3.rds", compress = "xz")
```

18Q4, CCLE RNAseq gene expression data (log2(TPM+1)):
```{r, eval=FALSE}
ge_18Q4 <- read.delim("./../htscreens_giant_files/CCLE_depMap_18Q4_TPM_v2_20181115.csv.gz", header = TRUE, sep = ",", check.names = FALSE)
# Remove Entrez gene IDs from colnames
# Format: Hugo_Symbol (Entrez ID)
colnames(ge_18Q4) <- gsub(" .*", "", colnames(ge_18Q4))
colnames(ge_18Q4)[1] <- "DepMap_ID"
# Melt
ge_melt_18Q4_all <- melt(ge_18Q4, id.vars = "DepMap_ID", measure.vars = colnames(ge_18Q4)[2:ncol(ge_18Q4)], variable.name = "Hugo_Symbol", value.name = "TPM")
ge_melt_18Q4 <- unique(ge_melt_18Q4)
saveRDS(ge_melt_18Q4, "./data_munging/rds/ge_melt_18Q4.rds", compress = "xz")
```

## G2P

```{r}
g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_NEW_DEC_2_2018_withGroupedLineages.csv", sep = "\t", header = TRUE)
g2p[, 1:2] <- NULL
g2p <- merge(g2p, lineage_converter, by = "grouped_general_doid", all.x = TRUE)

g2p$Drug <- tolower(g2p$Drug)
g2p[g2p == -1] <- NA
g2p$Ref[g2p$Ref == "None"] <- NA
g2p$Alt[g2p$Alt == "None"] <- NA
g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p$Drug_Gene_Lin <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene) | is.na(g2p$group_general_lineage_name), NA, paste0(g2p$Drug, "_", g2p$Gene, "_", g2p$group_general_lineage_name))
g2p$position <- paste0("g.chr", g2p$Chromosome, ":", g2p$End, g2p$Ref, ">", g2p$Alt)
g2p$position <- ifelse(grepl("NA", g2p$position), NA, g2p$position)
g2p_filt <- subset(g2p, grepl("^CID", g2p$DrugID))

g2p_dgls <-  unique(g2p_filt$Drug_Gene_Lin[!is.na(g2p_filt$Drug_Gene_Lin)])
g2p_druggenes <- unique(g2p_filt$Drug_Gene[!is.na(g2p_filt$Drug_Gene)])
g2p_genes <- unique(sapply(strsplit(g2p_druggenes, "_", fixed = TRUE), function(x) x[2]))
g2p_drugs <- unique(sapply(strsplit(g2p_druggenes, "_", fixed = TRUE), function(x) x[1]))
g2p_druggenes
g2p_genes
g2p_drugs
# Order and genes taken from pink heatmapo from Daniel's analyses
g2p_genes <- c("EGFR", "KRAS", "ERBB2", "ERBB3", "MET", "ALK", "ROS1", "PDGFRA", "ABL1", "KIT", "PDGFRB", "PDGFB", "BRAF", "G6PD", "RET", "BRCA1", "BRCA2", "UGT1A1", "TSC1", "TSC2", "AKT1", "IDH2", "DPYD", "ESR1", "CYP19A1", "TPMT", "FLT3", "NPM1", "DNMT3A", "PML", "PTCH1", "JAK2", "MYD88", "MGMT")
```

```{r}
# levels(g2p_filt$Phenotype.Description)
unique(g2p_filt$group_general_lineage_name)
levels(g2p_filt$group_general_lineage_name)

g2p_nonCID <- subset(g2p, !grepl("^CID", g2p$DrugID))
summ <- g2p_nonCID %>% group_by(Drug, DrugID) %>% tally() %>% summarize(DrugID = DrugID, n = n, percent = formatC(n / 402 * 100, digits = 6, format = "f"))
summ
```

```{r, eval=FALSE}
maf_df <- readRDS("./data_munging/rds/maf_df_18Q4.rds")
cn_melt <- readRDS("./data_munging/rds/cn_melt_18Q4.rds")
ge_melt_18Q3 <- readRDS("./../htscreens_giant_files/ge_melt_18Q3.rds")
# ge_melt_18Q4 <- readRDS("./data_munging/rds/ge_melt_18Q4.rds")

# maf_filt_dan <- filter(maf_df, Hugo_Symbol %in% g2p_genes)
# write.table(maf_filt_dan, file = "~/Desktop/maf_filt_dan_18Q4_2018112.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

## Drug screens

### Merge drug data and annotations

```{r, eval=FALSE}
# Drug metadata
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)
maf_df_g2p <- filter(maf_df, Hugo_Symbol %in% g2p_genes)
ge_filt_18Q3 <- filter(ge_melt_18Q3, Hugo_Symbol %in% g2p_genes)
# ge_filt_18Q4 <- filter(ge_melt_18Q4, Hugo_Symbol %in% g2p_genes)
cn_filt <- filter(cn_melt, Hugo_Symbol %in% g2p_genes)

# CCLE
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "ActArea")
ccle_grid <- expand.grid("accession_id" = unique(ccle_melt$accession_id), "Hugo_Symbol" = g2p_genes)
ccle_data <- merge(ccle_grid, ccle_melt, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, ccl_converter, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
ccle_data$Mutation_Status_Nonsilent <- ifelse(is.na(ccle_data$Mutation_Status_Nonsilent), "Wildtype", ccle_data$Mutation_Status_Nonsilent)
ccle_data <- merge(ccle_data, ge_filt_18Q3, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
# ccle_data <- merge(ccle_data, ge_filt_18Q4, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ccle_data <- merge(ccle_data, cn_filt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ccle_data$Hugo_Symbol <- factor(ccle_data$Hugo_Symbol)
ccle_data$ActArea <- as.numeric(ccle_data$ActArea)
# CCLE AUC is actually AOC
# assuming scale is 0-8, transform to range between 0 and 1
# get AUC by 1 - AOC
ccle_data$AUC <- 1 - (ccle_data$ActArea / 8)
ccle_data <- filter(ccle_data, !is.na(ccle_data$AUC))
ccle_data$AUC_zscore <- (ccle_data$AUC - mean(ccle_data$AUC)) / sd(ccle_data$AUC)
ccle_data$Color_Nonsilent <- ifelse(ccle_data$Mutation_Status_Nonsilent == "Wildtype", "palegreen3", "springgreen4")
ccle_data$Color_Nonsilent <- factor(ccle_data$Color_Nonsilent)
ccle_data$Drug_Gene <- paste0(ccle_data$Drug, "_", ccle_data$Hugo_Symbol)
ccle_data$Drug_Gene_Lin <- paste0(ccle_data$Drug, "_", ccle_data$Hugo_Symbol, "_", ccle_data$group_general_lineage_name)
saveRDS(ccle_data, "./data_munging/rds/ccle_data_18Q4_g2pgenes.rds", compress = "xz")

# CTRP
ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ctrp <- merge(drug_meta, ctrp, by = "CID")
ctrp_melt <- melt(ctrp, id.vars = c("CID", "Drug"), measure.vars = colnames(ctrp)[3:ncol(ctrp)], variable.name = "accession_id", value.name = "AUC")
ctrp_grid <- expand.grid("accession_id" = unique(ctrp_melt$accession_id), "Hugo_Symbol" = g2p_genes)
ctrp_data <- merge(ctrp_grid, ctrp_melt, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, ccl_converter, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
ctrp_data$Mutation_Status_Nonsilent <- ifelse(is.na(ctrp_data$Mutation_Status_Nonsilent), "Wildtype", ctrp_data$Mutation_Status_Nonsilent)
ctrp_data <- merge(ctrp_data, ge_filt_18Q3, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
# ctrp_data <- merge(ctrp_data, ge_filt_18Q4, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ctrp_data <- merge(ctrp_data, cn_filt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ctrp_data$Hugo_Symbol <- factor(ctrp_data$Hugo_Symbol)
ctrp_data$AUC <- as.numeric(ctrp_data$AUC)
ctrp_data <- filter(ctrp_data, !is.na(ctrp_data$AUC))
ctrp_data$AUC_zscore <- (ctrp_data$AUC - mean(ctrp_data$AUC)) / sd(ctrp_data$AUC)
ctrp_data$Color_Nonsilent <- ifelse(ctrp_data$Mutation_Status_Nonsilent == "Wildtype", "slategray3", "steelblue4")
ctrp_data$Color_Nonsilent <- factor(ctrp_data$Color_Nonsilent)
ctrp_data$Drug_Gene <- paste0(ctrp_data$Drug, "_", ctrp_data$Hugo_Symbol)
ctrp_data$Drug_Gene_Lin <- paste0(ctrp_data$Drug, "_", ctrp_data$Hugo_Symbol, "_", ctrp_data$group_general_lineage_name)
saveRDS(ctrp_data, "./data_munging/rds/ctrp_data_18Q4_g2pgenes.rds", compress = "xz")

# GDSC
gdsc <- read.delim("./data_munging/data_drug_auc_gdsc_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
gdsc <- merge(drug_meta, gdsc, by = "CID")
gdsc_melt <- melt(gdsc, id.vars = c("CID", "Drug"), measure.vars = colnames(gdsc)[3:ncol(gdsc)], variable.name = "accession_id", value.name = "AUC")
gdsc_grid <- expand.grid("accession_id" = unique(gdsc_melt$accession_id), "Hugo_Symbol" = g2p_genes)
gdsc_data <- merge(gdsc_grid, gdsc_melt, by = "accession_id", all.x = TRUE)
gdsc_data <- merge(gdsc_data, ccl_converter, by = "accession_id", all.x = TRUE)
gdsc_data <- merge(gdsc_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
gdsc_data$Mutation_Status_Nonsilent <- ifelse(is.na(gdsc_data$Mutation_Status_Nonsilent), "Wildtype", gdsc_data$Mutation_Status_Nonsilent)
gdsc_data <- merge(gdsc_data, ge_filt_18Q3, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
# gdsc_data <- merge(gdsc_data, ge_filt_18Q4, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
gdsc_data <- merge(gdsc_data, cn_filt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
gdsc_data$Hugo_Symbol <- factor(gdsc_data$Hugo_Symbol)
gdsc_data$AUC <- as.numeric(gdsc_data$AUC)
gdsc_data <- filter(gdsc_data, !is.na(gdsc_data$AUC))
gdsc_data$AUC_zscore <- (gdsc_data$AUC - mean(gdsc_data$AUC)) / sd(gdsc_data$AUC)
gdsc_data$Color_Nonsilent <- ifelse(gdsc_data$Mutation_Status_Nonsilent == "Wildtype", "paleturquoise3", "turquoise4")
gdsc_data$Color_Nonsilent <- factor(gdsc_data$Color_Nonsilent)
gdsc_data$Drug_Gene <- paste0(gdsc_data$Drug, "_", gdsc_data$Hugo_Symbol)
gdsc_data$Drug_Gene_Lin <- paste0(gdsc_data$Drug, "_", gdsc_data$Hugo_Symbol, "_", gdsc_data$group_general_lineage_name)
saveRDS(gdsc_data, "./data_munging/rds/gdsc_data_18Q4_g2pgenes.rds", compress = "xz")
```

### Filter drug data

```{r}
ccle_data <- readRDS("./data_munging/rds/ccle_data_18Q4_g2pgenes.rds")
ccle_data_ptmuts <- filter(ccle_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
ccle_data_ptmuts$Keep_PtMut_Tests <- ifelse(ccle_data_ptmuts$Reference_Allele_Length == 1 | is.na(ccle_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(ccle_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
ccle_data_ptmuts <- filter(ccle_data_ptmuts, Keep_PtMut_Tests == TRUE)
ccle_data_ptmuts$Dataset <- "CCLE"
ccle_data_ptmuts$Hugo_Symbol <- ordered(ccle_data_ptmuts$Hugo_Symbol, levels = g2p_genes)

ctrp_data <- readRDS("./data_munging/rds/ctrp_data_18Q4_g2pgenes.rds")
ctrp_data_ptmuts <- filter(ctrp_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
ctrp_data_ptmuts$Keep_PtMut_Tests <- ifelse(ctrp_data_ptmuts$Reference_Allele_Length == 1 | is.na(ctrp_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(ctrp_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
ctrp_data_ptmuts <- filter(ctrp_data_ptmuts, Keep_PtMut_Tests == TRUE)
ctrp_data_ptmuts$Dataset <- "CTRP"
ctrp_data_ptmuts$Hugo_Symbol <- ordered(ctrp_data_ptmuts$Hugo_Symbol, levels = g2p_genes)

gdsc_data <- readRDS("./data_munging/rds/gdsc_data_18Q4_g2pgenes.rds")
gdsc_data_ptmuts <- filter(gdsc_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
gdsc_data_ptmuts$Keep_PtMut_Tests <- ifelse(gdsc_data_ptmuts$Reference_Allele_Length == 1 | is.na(gdsc_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(gdsc_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
gdsc_data_ptmuts <- filter(gdsc_data_ptmuts, Keep_PtMut_Tests == TRUE)
gdsc_data_ptmuts$Dataset <- "GDSC"
gdsc_data_ptmuts$Hugo_Symbol <- ordered(gdsc_data_ptmuts$Hugo_Symbol, levels = g2p_genes)

# Filter drug screen datasets for G2P drug-gene associations
ccle_data_g2p <- filter(ccle_data_ptmuts, Drug_Gene %in% g2p_druggenes)
ccle_data_g2p$Mutation_Status_Nonsilent <- factor(ccle_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p <- filter(ctrp_data_ptmuts, Drug_Gene %in% g2p_druggenes)
ctrp_data_g2p$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p <- filter(gdsc_data_ptmuts, Drug_Gene %in% g2p_druggenes)
gdsc_data_g2p$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

# Filter drug screen datasets for G2P drug-gene-lineage associations
ccle_data_g2p_dgl <- filter(ccle_data_ptmuts, Drug_Gene_Lin %in% g2p_dgls)
ccle_data_g2p_dgl$Mutation_Status_Nonsilent <- factor(ccle_data_g2p_dgl$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p_dgl <- filter(ctrp_data_ptmuts, Drug_Gene_Lin %in% g2p_dgls)
ctrp_data_g2p_dgl$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p_dgl$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p_dgl <- filter(gdsc_data_ptmuts, Drug_Gene_Lin %in% g2p_dgls)
gdsc_data_g2p_dgl$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p_dgl$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

# Filter drug screen datasets for G2P drugs and genes
ccle_data_g2p_grid <- filter(ccle_data_ptmuts, Drug %in% g2p_drugs)
ccle_data_g2p_grid$Mutation_Status_Nonsilent <- factor(ccle_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p_grid <- filter(ctrp_data_ptmuts, Drug %in% g2p_drugs)
ctrp_data_g2p_grid$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p_grid <- filter(gdsc_data_ptmuts, Drug %in% g2p_drugs)
gdsc_data_g2p_grid$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
```

```{r}
test_ccle <- ccle_data_ptmuts %>% group_by(Hugo_Symbol, Mutation_Status_Nonsilent) %>% tally()
test_ctrp <- ctrp_data_ptmuts %>% group_by(Hugo_Symbol, Mutation_Status_Nonsilent) %>% tally()
test_gdsc <- gdsc_data_ptmuts %>% group_by(Hugo_Symbol, Mutation_Status_Nonsilent) %>% tally()
```

## CRISPR

### DepMap dependency probabilities

```{r, eval=FALSE}
dep <- read.delim("./data_munging/18Q4/gene_dependency_18Q4_20181115.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(dep) <- gsub(" .*", "", colnames(dep))
```

### Merge CRISPR data and annotations

The latest CRISPR CERES score data (18Q3, August 2018) was pulled from the [DepMap Data Portal](https://depmap.org/portal/download/) (Broad Institute Cancer Dependency Map 2018, Meyers et al. 2017).
```{r, eval=FALSE}
crispr <- read.delim("./data_munging/18Q4/gene_effect_18Q4_20181115.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr) <- gsub(" .*", "", colnames(crispr))
```

Merge annotation data:
```{r, eval=FALSE}
# Melt CRISPR dataset for merging
crispr_melt <- melt(crispr, id.vars = "DepMap_ID", measure.vars = colnames(crispr)[2:ncol(crispr)], variable.name = "Hugo_Symbol", value.name = "Score")
# Melt dependency probabilities dataset for merging
dep_melt <- melt(dep, id.vars = "DepMap_ID", measure.vars = colnames(dep)[2:ncol(dep)], variable.name = "Hugo_Symbol", value.name = "Dep_Prob")
# Merge dependency probabilities
crispr_melt <- merge(crispr_melt, dep_melt, by = c("DepMap_ID", "Hugo_Symbol"), all.x = TRUE)
# Merge cell line metadata
crispr_melt <- merge(crispr_melt, ccl_info, by = "DepMap_ID", all.x = TRUE)
crispr_melt <- merge(crispr_melt, crispr_meta, by = c("CCLE_Name", "DepMap_ID"), all.x = TRUE)

# Merge mutation annotations
maf_filt_crispr <- filter(maf_df, Hugo_Symbol %in% as.character(unique(crispr_melt$Hugo_Symbol)))
crispr_muts <- merge(crispr_melt, maf_filt_crispr, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
crispr_muts$Hugo_Symbol <- factor(crispr_muts$Hugo_Symbol)
crispr_muts <- crispr_muts %>% mutate(Mutation_Status_Nonsilent = if_else(is.na(Mutation_Status_Nonsilent), "Wildtype", Mutation_Status_Nonsilent))
# Summarize number of mutant and Wildtype cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Hugo_Symbol) %>%
  summarize(N_Nonsilent_Wildtype = sum(Mutation_Status_Nonsilent == "Wildtype"),
            N_Nonsilent_Mutant = sum(Mutation_Status_Nonsilent == "Mutant"))
# Merge test results back into full dataset, which restores information lost in the summarization
crispr_data <- merge(crispr_muts_summ, crispr_muts, by = "Hugo_Symbol")
# Add Color columns
crispr_data$Color_Nonsilent <- ifelse(crispr_data$Mutation_Status_Nonsilent == "Wildtype", "thistle", "darkorchid")
crispr_data$Color_Nonsilent <- factor(crispr_data$Color_Nonsilent)

# Cell line lineages
crispr_data <- merge(crispr_data, ccl_converter, by = c("CCLE_Name", "DepMap_ID"), all.x = TRUE)
levels(crispr_data$lineage_name) <- sort(levels(crispr_data$lineage_name), decreasing = TRUE)
# Copy number
crispr_data <- merge(crispr_data, cn_melt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
# Gene expression (RPKM)
ge_crispr_18Q3 <- filter(ge_melt_18Q3, Hugo_Symbol %in% as.character(unique(crispr_melt$Hugo_Symbol)))
# ge_crispr_18Q4 <- filter(ge_melt_18Q4, Hugo_Symbol %in% as.character(unique(crispr_melt$Hugo_Symbol)))
crispr_data <- merge(crispr_data, ge_crispr_18Q3, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
# crispr_data <- merge(crispr_data, ge_crispr_18Q4, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
saveRDS(crispr_data, "./../htscreens_giant_files/crispr_data_18Q4.rds", compress = "xz")
```

```{r}
crispr_data <- readRDS("./../htscreens_giant_files/crispr_data_18Q4.rds")
crispr_data_ptmuts <- filter(crispr_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele_Length == 1 | is.na(crispr_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)

# Filter CRISPR for G2P genes in G2P
crispr_data_g2p <- filter(crispr_data_ptmuts, Hugo_Symbol %in% g2p_genes)
crispr_data_g2p$Hugo_Symbol <- factor(crispr_data_g2p$Hugo_Symbol, levels = g2p_genes)
crispr_data_g2p$Mutation_Status_Nonsilent <- factor(crispr_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
```

# WILCOXON TESTS

--------------------------------------------------------------------------------

Wilcoxon tests for assessing effect of mutation status on gene essentiality scores (CRISPR) and AUC (CCLE, CTRP, and GDSC).

## CRISPR

Test G2P genes in lineage-agnostic context:
```{r, eval=FALSE}
crispr_signif_g2p_gene <- compare_means(Score ~ Mutation_Status_Nonsilent, group.by = c("Hugo_Symbol"), data = crispr_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_g2p_gene <- adj_signif(crispr_signif_g2p_gene)
crispr_signif_g2p_gene <- crispr_signif_g2p_gene[order(crispr_signif_g2p_gene$p),]
saveRDS(crispr_signif_g2p_gene, "./data_munging/rds/crispr_signif_g2p_gene.rds")
```

```{r}
crispr_signif_g2p_gene <- readRDS("./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_g2p_gene <- rbind(crispr_signif_g2p_gene, c("G6PD", rep(NA, 11)))
knitr::kable(crispr_signif_g2p_gene[, c("Hugo_Symbol", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results for Level A point mutations (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

Plot results of `crispr_signif_g2p_gene` tests:
```{r fig.height=7, fig.width=12}
crispr_data_signif_g2p_gene_filt <- filter(crispr_data_ptmuts, Hugo_Symbol %in% as.character(crispr_signif_g2p_gene$Hugo_Symbol))
crispr_data_signif_g2p_gene <- merge(crispr_data_signif_g2p_gene_filt, crispr_signif_g2p_gene, by = "Hugo_Symbol")
crispr_data_signif_g2p_gene$Hugo_Symbol <- ordered(crispr_data_signif_g2p_gene$Hugo_Symbol, levels = as.character(crispr_signif_g2p_gene$Hugo_Symbol))

crispr_data_signif_g2p_gene <- merge(crispr_data_signif_g2p_gene, crispr_data_signif_g2p_gene %>% group_by(Hugo_Symbol) %>% summarize(Text_y = max(Score)), by = "Hugo_Symbol")
crispr_data_signif_g2p_gene_text <- unique(data.frame(p.signif = crispr_data_signif_g2p_gene$p.signif, p.signif.adj = crispr_data_signif_g2p_gene$p.signif.adj,  Hugo_Symbol = crispr_data_signif_g2p_gene$Hugo_Symbol, Text_y = crispr_data_signif_g2p_gene$Text_y))

crispr_data_signif_g2p_gene_plot <- ggplot(data = crispr_data_signif_g2p_gene, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, nrow = 2, scales = "free_y") +
  geom_jitter(aes(color = Mutation_Status_Nonsilent), alpha = 0.5, width = 0.3) +
  geom_boxplot(color = "black", fill = NA, outlier.shape = NA) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = c("Mutant" = "darkorchid", "Wildtype" = "thistle")) +
  geom_text(data = crispr_data_signif_g2p_gene_text, mapping = aes(x = 1.5, y = Text_y, label = p.signif), nudge_y = 0.2) +
  theme(legend.position = "top", axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank()) +
  labs(y = "CERES Score",
       color = "Mutation Status",
       title = "CERES scores for select Level A G2P genes",
       subtitle = "Genes are sorted by increasing p-value.")
crispr_data_signif_g2p_gene_plot

# ggsave("./plots/manuscript/crispr_data_signif_g2p_gene_plot.png", crispr_data_signif_g2p_gene_plot, width = 12, height = 7, units = "in")
```

Test G2P genes in lineage-grouped context:
```{r, eval=FALSE}
crispr_signif_g2p_lineage <- compare_means(Score ~ Mutation_Status_Nonsilent, group.by = c("Hugo_Symbol", "group_general_lineage_name"), data = crispr_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_g2p_lineage <- adj_signif(crispr_signif_g2p_lineage)
crispr_signif_g2p_lineage <- crispr_signif_g2p_lineage[order(crispr_signif_g2p_lineage$p),]
saveRDS(crispr_signif_g2p_lineage, "./data_munging/rds/crispr_signif_g2p_lineage.rds")
# write.table(crispr_signif_g2p_lineage, file = "~/Desktop/crispr_signif_g2p_lineage.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

```{r}
crispr_signif_g2p_lineage <- readRDS("./data_munging/rds/crispr_signif_g2p_lineage.rds")
knitr::kable(filter(crispr_signif_g2p_lineage, p < 0.1)[, c("Hugo_Symbol", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results comparing non-silent mutant vs wildtype cell lines by lineage, p < 0.1 (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

Plot results of `crispr_signif_g2p_lineage` tests:
```{r fig.height=18, fig.width=15, eval=TRUE}
crispr_data_signif_g2p_lineage <- merge(crispr_data_signif_g2p_gene_filt, crispr_signif_g2p_lineage, by = c("Hugo_Symbol", "group_general_lineage_name"))
crispr_data_signif_g2p_lineage$Hugo_Symbol <- ordered(crispr_data_signif_g2p_lineage$Hugo_Symbol, levels = as.character(unique((crispr_signif_g2p_lineage$Hugo_Symbol))))

crispr_data_signif_g2p_lineage <- merge(crispr_data_signif_g2p_lineage, crispr_data_signif_g2p_lineage %>% group_by(Hugo_Symbol) %>% summarize(Text_y = max(Score)), by = "Hugo_Symbol")

crispr_data_signif_g2p_lineage_text <- unique(data.frame(p.signif = crispr_data_signif_g2p_lineage$p.signif, p.signif.adj = crispr_data_signif_g2p_lineage$p.signif.adj,  group_general_lineage_name = crispr_data_signif_g2p_lineage$group_general_lineage_name, Hugo_Symbol = crispr_data_signif_g2p_lineage$Hugo_Symbol, Text_y = crispr_data_signif_g2p_lineage$Text_y))

crispr_data_signif_g2p_lineage_plot <- ggplot(data = crispr_data_signif_g2p_lineage, aes(x = group_general_lineage_name, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, nrow = 6, scales = "free_y") +
  geom_point(aes(color = Mutation_Status_Nonsilent), alpha = 0.5) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = c("Mutant" = "darkorchid", "Wildtype" = "thistle")) +
  geom_text(data = crispr_data_signif_g2p_lineage_text, mapping = aes(x = group_general_lineage_name, y = Text_y, label = p.signif), angle = 90, vjust = 0.8, nudge_y = 0.1) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(y = "CERES Score",
       x = "Lineage",
       color = "Mutation Status",
       title = "CERES score by cell line lineage for significant Level A G2P genes",
       subtitle = "* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001\nGenes are sorted by increasing lineage-agnostic p-value.")
crispr_data_signif_g2p_lineage_plot

# write.table(crispr_data_signif_g2p_lineage, file = "~/Desktop/crispr_data_signif_g2p_lineage_20181213.tsv", quote = FALSE, sep = "\t")
# ggsave("./plots/manuscript/crispr_data_signif_g2p_lineage_plot.png", crispr_data_signif_g2p_lineage_plot, width = 15, height = 18, units = "in")
```

Count mutant/wildtype calls in each gene-lineage group:
```{r}
crispr_data_signif_g2p_lineage_summ <- crispr_data_signif_g2p_lineage %>% group_by(Hugo_Symbol, group_general_lineage_name, Mutation_Status_Nonsilent) %>% tally()

# write.table(crispr_data_signif_g2p_lineage_summ, file = "~/Desktop/crispr_data_signif_g2p_lineage_summ_20181005.tsv", quote = FALSE, sep = "\t")

knitr::kable(crispr_data_signif_g2p_lineage_summ, caption = "Mutant/Wildtype counts for gene-lineage combinations") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

## CCLE

Drug metadata:
```{r}
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)
```

Test all CRISPR genes in lineage-agnostic context:
```{r, eval=FALSE}
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "AUC")
ccle_grid <- expand.grid("accession_id" = unique(ccle_melt$accession_id), "Hugo_Symbol" = as.character(unique(crispr_data$Hugo_Symbol)))
system.time({ ccle_data_allgenes <- WilcoxonByDrugAllGenes(ccle_melt, ccle_grid, "CCLE") })
saveRDS(ccle_data_allgenes, "./data_munging/rds/ccle_data_18Q3_allgenes.rds", compress = "xz")

# user   system  elapsed 
# 1483.399  262.649 1901.698 
```

```{r}
ccle_data_allgenes <- readRDS("./data_munging/rds/ccle_data_18Q3_allgenes.rds")
ccle_data_allgenes <- ccle_data_allgenes[order(ccle_data_allgenes$p),]
knitr::kable(ccle_data_allgenes[1:1000, c("Hugo_Symbol", "Drug", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: by-drug Wilcoxon test results for point mutations in genes targeted in CRISPR screen (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

Test G2P genes in lineage-agnostic context:
```{r, eval=FALSE}
# G2P drug-gene associations only
ccle_signif_g2p <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = ccle_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
ccle_signif_g2p <- adj_signif(ccle_signif_g2p)
ccle_signif_g2p <- ccle_signif_g2p[order(ccle_signif_g2p$p),]
saveRDS(ccle_signif_g2p, "./data_munging/rds/ccle_signif_g2p_gene.rds")

# Grid comparisons of G2P drugs and genes
ccle_signif_g2p_grid <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = c("Drug_Gene"), data = ccle_data_g2p_grid, method = "wilcox.test", p.adjust.method = "BH")
ccle_signif_g2p_grid <- adj_signif(ccle_signif_g2p_grid)
ccle_signif_g2p_grid <- ccle_signif_g2p_grid[order(ccle_signif_g2p_grid$p),]
saveRDS(ccle_signif_g2p_grid, "./data_munging/rds/ccle_signif_g2p_grid.rds")
```

```{r}
ccle_signif_g2p <- readRDS("./data_munging/rds/ccle_signif_g2p_gene.rds")
knitr::kable(ccle_signif_g2p[, c("Drug_Gene", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: Wilcoxon test results for level A G2P drug-gene associations (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")

ccle_signif_g2p_grid <- readRDS("./data_munging/rds/ccle_signif_g2p_grid.rds")
knitr::kable(ccle_signif_g2p_grid[, c("Drug_Gene", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: Wilcoxon test results for point mutations in level A G2P genes and drugs (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

Test G2P genes in lineage-grouped context:
```{r, eval=FALSE}
ccle_signif_grid_lineage <- lapply(unique(ccle_data_g2p_grid$Drug), WilcoxonByDrug, dataset = ccle_data_g2p_grid, data_name = "ccle", grouping_var = "lineage")
names(ccle_signif_grid_lineage) <- unique(ccle_signif_grid_lineage$Drug)
ccle_signif_grid_lineage <- rbindlist(ccle_signif_grid_lineage, use.names = TRUE)
ccle_signif_grid_lineage <- ccle_signif_grid_lineage[order(ccle_signif_grid_lineage$p),]
saveRDS(ccle_signif_grid_lineage, "./data_munging/rds/ccle_signif_grid_lineage.rds")

ccle_signif_g2p_lineage <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene_Lin", data = ccle_data_g2p_dgl, method = "wilcox.test", p.adjust.method = "BH")
ccle_signif_g2p_lineage <- adj_signif(ccle_signif_g2p_lineage)
ccle_signif_g2p_lineage <- ccle_signif_g2p_lineage[order(ccle_signif_g2p_lineage$p),]
saveRDS(ccle_signif_g2p_lineage, "./data_munging/rds/ccle_signif_g2p_lineage.rds")
```

```{r}
ccle_signif_g2p_lineage <- readRDS("./data_munging/rds/ccle_signif_g2p_lineage.rds")
ccle_signif_grid_lineage <- readRDS("./data_munging/rds/ccle_signif_grid_lineage.rds")

knitr::kable(ccle_signif_g2p_lineage[, c("Drug_Gene_Lin", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: Wilcoxon test results for level A G2P drug-gene associations across lineages (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")

knitr::kable(ccle_signif_grid_lineage[1:500, c("Hugo_Symbol", "Drug", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: Wilcoxon test results for level A G2P drug-gene associations across lineages (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

## CTRP

Test all CRISPR genes in lineage-agnostic context:
```{r, eval=FALSE}
ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ctrp <- merge(drug_meta, ctrp, by = "CID")
ctrp_melt <- melt(ctrp, id.var  s = c("CID", "Drug"), measure.vars = colnames(ctrp)[3:ncol(ctrp)], variable.name = "accession_id", value.name = "AUC")
ctrp_grid <- expand.grid("accession_id" = unique(ctrp_melt$accession_id), "Hugo_Symbol" = as.character(unique(crispr_data$Hugo_Symbol)))
ctrp_data_allgenes <- WilcoxonByDrugAllGenes(ctrp_melt, ctrp_grid, "CTRP")
saveRDS(ctrp_data_allgenes, "./data_munging/rds/ctrp_data_18Q3_allgenes.rds", compress = "xz")
```

```{r}
ctrp_data_allgenes <- readRDS("./data_munging/rds/ctrp_data_18Q3_allgenes.rds")
ctrp_data_allgenes <- ctrp_data_allgenes[order(ctrp_data_allgenes$p),]
knitr::kable(ctrp_data_allgenes[1:1000, c("Hugo_Symbol", "Drug", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CTRP: by-drug Wilcoxon test results for point mutations in genes targeted in CRISPR screen (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

Test G2P genes in lineage-agnostic context:
```{r, eval=FALSE}
ctrp_signif_g2p <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = ctrp_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
ctrp_signif_g2p <- adj_signif(ctrp_signif_g2p)
ctrp_signif_g2p <- ctrp_signif_g2p[order(ctrp_signif_g2p$p),]
saveRDS(ctrp_signif_g2p, "./data_munging/rds/ctrp_signif_g2p_gene.rds")

# Grid comparisons of G2P drugs and genes
ctrp_signif_g2p_grid <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = ctrp_data_g2p_grid, method = "wilcox.test", p.adjust.method = "BH")
ctrp_signif_g2p_grid <- adj_signif(ctrp_signif_g2p_grid)
ctrp_signif_g2p_grid <- ctrp_signif_g2p_grid[order(ctrp_signif_g2p_grid$p),]
saveRDS(ctrp_signif_g2p_grid, "./data_munging/rds/ctrp_signif_g2p_grid.rds")
```

```{r}
ctrp_signif_g2p <- readRDS("./data_munging/rds/ctrp_signif_g2p_gene.rds")
knitr::kable(ctrp_signif_g2p[, c("Drug_Gene", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CTRP: Wilcoxon test results for level A G2P drug-gene associations (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")

ctrp_signif_g2p_grid <- readRDS("./data_munging/rds/ctrp_signif_g2p_grid.rds")
knitr::kable(ctrp_signif_g2p_grid[, c("Drug_Gene", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CTRP: Wilcoxon test results for point mutations in level A G2P genes and drugs (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

Test G2P genes in lineage-grouped context:
```{r, eval=FALSE}
ctrp_signif_grid_lineage <- lapply(unique(ctrp_data_g2p_grid$Drug), WilcoxonByDrug, dataset = ctrp_data_g2p_grid, data_name = "ctrp", grouping_var = "lineage")
names(ctrp_signif_grid_lineage) <- unique(ctrp_signif_grid_lineage$Drug)
ctrp_signif_grid_lineage <- rbindlist(ctrp_signif_grid_lineage, use.names = TRUE)
ctrp_signif_grid_lineage <- ctrp_signif_grid_lineage[order(ctrp_signif_grid_lineage$p),]
saveRDS(ctrp_signif_grid_lineage, "./data_munging/rds/ctrp_signif_grid_lineage.rds")

ctrp_signif_g2p_lineage <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene_Lin", data = ctrp_data_g2p_dgl, method = "wilcox.test", p.adjust.method = "BH")
ctrp_signif_g2p_lineage <- adj_signif(ctrp_signif_g2p_lineage)
ctrp_signif_g2p_lineage <- ctrp_signif_g2p_lineage[order(ctrp_signif_g2p_lineage$p),]
saveRDS(ctrp_signif_g2p_lineage, "./data_munging/rds/ctrp_signif_g2p_lineage.rds")
```

```{r}
ctrp_signif_g2p_lineage <- readRDS("./data_munging/rds/ctrp_signif_g2p_lineage.rds")
ctrp_signif_grid_lineage <- readRDS("./data_munging/rds/ctrp_signif_grid_lineage.rds")

knitr::kable(ctrp_signif_g2p_lineage[, c("Drug_Gene_Lin", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CTRP: Wilcoxon test results for level A G2P drug-gene associations across lineages (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")

knitr::kable(ctrp_signif_grid_lineage[1:500, c("Hugo_Symbol", "Drug", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CTRP: Wilcoxon test results for level A G2P drug-gene associations across lineages (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

## GDSC

Test all CRISPR genes in lineage-agnostic context:
```{r, eval=FALSE}
gdsc <- read.delim("./data_munging/data_drug_auc_gdsc_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
gdsc <- merge(drug_meta, gdsc, by = "CID")
gdsc_melt <- melt(gdsc, id.vars = c("CID", "Drug"), measure.vars = colnames(gdsc)[3:ncol(gdsc)], variable.name = "accession_id", value.name = "AUC")
gdsc_grid <- expand.grid("accession_id" = unique(gdsc_melt$accession_id), "Hugo_Symbol" = as.character(unique(crispr_data$Hugo_Symbol)))
system.time({ gdsc_data_allgenes <- WilcoxonByDrugAllGenes(gdsc_melt, gdsc_grid, "GDSC") })
saveRDS(gdsc_data_allgenes, "./data_munging/rds/gdsc_data_18Q3_allgenes.rds", compress = "xz")
```

```{r}
gdsc_data_allgenes <- readRDS("./data_munging/rds/gdsc_data_18Q3_allgenes.rds")
gdsc_data_allgenes <- gdsc_data_allgenes[order(gdsc_data_allgenes$p),]
knitr::kable(gdsc_data_allgenes[1:1000, c("Hugo_Symbol", "Drug", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "GDSC: by-drug Wilcoxon test results for point mutations in genes targeted in CRISPR screen (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

Test G2P genes in lineage-agnostic context:
```{r, eval=FALSE}
gdsc_signif_g2p <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = gdsc_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
gdsc_signif_g2p <- adj_signif(gdsc_signif_g2p)
gdsc_signif_g2p <- gdsc_signif_g2p[order(gdsc_signif_g2p$p),]
saveRDS(gdsc_signif_g2p, "./data_munging/rds/gdsc_signif_g2p_gene.rds")

# Grid comparisons of G2P drugs and genes
gdsc_signif_g2p_grid <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = gdsc_data_g2p_grid, method = "wilcox.test", p.adjust.method = "BH")
gdsc_signif_g2p_grid <- adj_signif(gdsc_signif_g2p_grid)
gdsc_signif_g2p_grid <- gdsc_signif_g2p_grid[order(gdsc_signif_g2p_grid$p),]
saveRDS(gdsc_signif_g2p_grid, "./data_munging/rds/gdsc_signif_g2p_grid.rds")
```

```{r}
gdsc_signif_g2p <- readRDS("./data_munging/rds/gdsc_signif_g2p_gene.rds")
knitr::kable(gdsc_signif_g2p[, c("Drug_Gene", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "GDSC: Wilcoxon test results for level A G2P drug-gene associations (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")

gdsc_signif_g2p_grid <- readRDS("./data_munging/rds/gdsc_signif_g2p_grid.rds")
knitr::kable(gdsc_signif_g2p_grid[, c("Drug_Gene", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "GDSC: Wilcoxon test results for point mutations in level A G2P genes and drugs (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

Test G2P genes in lineage-grouped context:
```{r, eval=FALSE}
gdsc_signif_grid_lineage <- lapply(unique(gdsc_data_g2p_grid$Drug), WilcoxonByDrug, dataset = gdsc_data_g2p_grid, data_name = "gdsc", grouping_var = "lineage")
names(gdsc_signif_grid_lineage) <- unique(gdsc_signif_grid_lineage$Drug)
gdsc_signif_grid_lineage <- rbindlist(gdsc_signif_grid_lineage, use.names = TRUE)
gdsc_signif_grid_lineage <- gdsc_signif_grid_lineage[order(gdsc_signif_grid_lineage$p),]
saveRDS(gdsc_signif_grid_lineage, "./data_munging/rds/gdsc_signif_grid_lineage.rds")

gdsc_signif_g2p_lineage <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene_Lin", data = gdsc_data_g2p_dgl, method = "wilcox.test", p.adjust.method = "BH")
gdsc_signif_g2p_lineage <- adj_signif(gdsc_signif_g2p_lineage)
gdsc_signif_g2p_lineage <- gdsc_signif_g2p_lineage[order(gdsc_signif_g2p_lineage$p),]
saveRDS(gdsc_signif_g2p_lineage, "./data_munging/rds/gdsc_signif_g2p_lineage.rds")
```

```{r}
gdsc_signif_g2p_lineage <- readRDS("./data_munging/rds/gdsc_signif_g2p_lineage.rds")
gdsc_signif_grid_lineage <- readRDS("./data_munging/rds/gdsc_signif_grid_lineage.rds")

knitr::kable(gdsc_signif_g2p_lineage[, c("Drug_Gene_Lin", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "GDSC: Wilcoxon test results for level A G2P drug-gene associations across lineages (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")

knitr::kable(gdsc_signif_grid_lineage[1:500, c("Hugo_Symbol", "Drug", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "GDSC: Wilcoxon test results for level A G2P drug-gene associations across lineages (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

# PEARSON/SPEARMAN CORRELATIONS

--------------------------------------------------------------------------------

Pearson correlations for assessing effect of gene expression and copy number on gene essentiality score and AUC.

## CRISPR

```{r}
crispr_pearson_res_GE <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>%
                         summarize(p = cor.test(y = Score, x = RPKM, method = "pearson", use = "complete.obs")$p.value,
                                   corr = cor.test(y = Score, x = RPKM, method = "pearson", use = "complete.obs")$estimate,
                                   Metric = "Gene Expression")
crispr_pearson_res_CN <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>%
                         summarize(p = cor.test(y = Score, x = CN, method = "pearson", use = "complete.obs")$p.value,
                                   corr = cor.test(y = Score, x = CN, method = "pearson", use = "complete.obs")$estimate,
                                   Metric = "Copy Number")
crispr_spearman_res_GE <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>%
                          summarize(p = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs")$p.value,
                                    corr = cor.test(y = Score, x = RPKM, method = "spearman", use = "complete.obs")$estimate,
                                    Metric = "Gene Expression")
crispr_spearman_res_CN <- crispr_data_g2p %>% group_by(Hugo_Symbol) %>%
                          summarize(p = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$p.value,
                                    corr = cor.test(y = Score, x = CN, method = "spearman", use = "complete.obs")$estimate,
                                    Metric = "Copy Number")
```

## Drug screens

```{r}
g2p_druggene_tally <- g2p_filt %>% group_by(Drug_Gene) %>% tally() %>% drop_na()

# CCLE
ccle_cor_res <- ccle_data_g2p_grid %>% group_by(Drug_Gene) %>%
                    summarize(Pearson_Corr_GE = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$estimate,
                              Pearson_Pval_GE = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$p.value,
                              Pearson_Corr_CN = cor.test(y = AUC, x = CN, method = "pearson", use = "complete.obs")$estimate,
                              Pearson_Pval_CN = cor.test(y = AUC, x = CN, method = "pearson", use = "complete.obs")$p.value,
                              Spearman_Corr_GE = cor.test(y = AUC, x = RPKM, method = "spearman", use = "complete.obs")$estimate,
                              Spearman_Pval_GE = cor.test(y = AUC, x = RPKM, method = "spearman", use = "complete.obs")$p.value,
                              Spearman_Corr_CN = cor.test(y = AUC, x = CN, method = "spearman", use = "complete.obs")$estimate,
                              Spearman_Pval_CN = cor.test(y = AUC, x = CN, method = "spearman", use = "complete.obs")$p.value)
ccle_cor_res <- merge(g2p_druggene_tally, ccle_cor_res, by = "Drug_Gene", all.y = TRUE)
ccle_cor_res$n <- ifelse(is.na(ccle_cor_res$n), 0, ccle_cor_res$n)
ccle_cor_res$InG2P <- ifelse(ccle_cor_res$n != 0, "Yes", "No")
ccle_cor_res$Dataset <- "CCLE"
ccle_pearson_ttest_GE <- t.test(ccle_cor_res$Pearson_Corr_GE ~ ccle_cor_res$InG2P)
ccle_pearson_ttest_res_GE <- paste0("t(", round(ccle_pearson_ttest_GE$parameter, 0), ") = ", round(ccle_pearson_ttest_GE$statistic, 2), ", p = ", round(ccle_pearson_ttest_GE$p.value, 2))
ccle_pearson_ttest_CN <- t.test(ccle_cor_res$Pearson_Corr_CN ~ ccle_cor_res$InG2P)
ccle_pearson_ttest_res_CN <- paste0("t(", round(ccle_pearson_ttest_CN$parameter, 0), ") = ", round(ccle_pearson_ttest_CN$statistic, 2), ", p = ", round(ccle_pearson_ttest_CN$p.value, 2))
ccle_spearman_ttest_GE <- t.test(ccle_cor_res$Spearman_Corr_GE ~ ccle_cor_res$InG2P)
ccle_spearman_ttest_res_GE <- paste0("t(", round(ccle_spearman_ttest_GE$parameter, 0), ") = ", round(ccle_spearman_ttest_GE$statistic, 2), ", p = ", round(ccle_spearman_ttest_GE$p.value, 2))
ccle_spearman_ttest_CN <- t.test(ccle_cor_res$Spearman_Corr_CN ~ ccle_cor_res$InG2P)
ccle_spearman_ttest_res_CN <- paste0("t(", round(ccle_spearman_ttest_CN$parameter, 0), ") = ", round(ccle_spearman_ttest_CN$statistic, 2), ", p = ", formatC(ccle_spearman_ttest_CN$p.value, 2, format = "f"))

# CTRP
ctrp_cor_res <- ctrp_data_g2p_grid %>% group_by(Drug_Gene) %>%
                    summarize(Pearson_Corr_GE = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$estimate,
                              Pearson_Pval_GE = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$p.value,
                              Pearson_Corr_CN = cor.test(y = AUC, x = CN, method = "pearson", use = "complete.obs")$estimate,
                              Pearson_Pval_CN = cor.test(y = AUC, x = CN, method = "pearson", use = "complete.obs")$p.value,
                              Spearman_Corr_GE = cor.test(y = AUC, x = RPKM, method = "spearman", use = "complete.obs")$estimate,
                              Spearman_Pval_GE = cor.test(y = AUC, x = RPKM, method = "spearman", use = "complete.obs")$p.value,
                              Spearman_Corr_CN = cor.test(y = AUC, x = CN, method = "spearman", use = "complete.obs")$estimate,
                              Spearman_Pval_CN = cor.test(y = AUC, x = CN, method = "spearman", use = "complete.obs")$p.value)
ctrp_cor_res <- merge(g2p_druggene_tally, ctrp_cor_res, by = "Drug_Gene", all.y = TRUE)
ctrp_cor_res$n <- ifelse(is.na(ctrp_cor_res$n), 0, ctrp_cor_res$n)
ctrp_cor_res$InG2P <- ifelse(ctrp_cor_res$n != 0, "Yes", "No")
ctrp_cor_res$Dataset <- "CTRP"
ctrp_pearson_ttest_GE <- t.test(ctrp_cor_res$Pearson_Corr_GE ~ ctrp_cor_res$InG2P)
ctrp_pearson_ttest_res_GE <- paste0("t(", round(ctrp_pearson_ttest_GE$parameter, 0), ") = ", round(ctrp_pearson_ttest_GE$statistic, 2), ", p ", ifelse(round(ctrp_pearson_ttest_GE$p.value, 3) == 0, "< 0.001", paste0("= ", round(ctrp_pearson_ttest_GE$p.value, 3))))
ctrp_pearson_ttest_CN <- t.test(ctrp_cor_res$Pearson_Corr_CN ~ ctrp_cor_res$InG2P)
ctrp_pearson_ttest_res_CN <- paste0("t(", round(ctrp_pearson_ttest_CN$parameter, 0), ") = ", round(ctrp_pearson_ttest_CN$statistic, 2), ", p = ", round(ctrp_pearson_ttest_CN$p.value, 3))
ctrp_spearman_ttest_GE <- t.test(ctrp_cor_res$Spearman_Corr_GE ~ ctrp_cor_res$InG2P)
ctrp_spearman_ttest_res_GE <- paste0("t(", round(ctrp_spearman_ttest_GE$parameter, 0), ") = ", round(ctrp_spearman_ttest_GE$statistic, 2), ", p = ", round(ctrp_spearman_ttest_GE$p.value, 2))
ctrp_spearman_ttest_CN <- t.test(ctrp_cor_res$Spearman_Corr_CN ~ ctrp_cor_res$InG2P)
ctrp_spearman_ttest_res_CN <- paste0("t(", round(ctrp_spearman_ttest_CN$parameter, 0), ") = ", round(ctrp_spearman_ttest_CN$statistic, 2), ", p = ", round(ctrp_spearman_ttest_CN$p.value, 2))

# GDSC
gdsc_cor_res <- gdsc_data_g2p_grid %>% group_by(Drug_Gene) %>%
                    summarize(Pearson_Corr_GE = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$estimate,
                              Pearson_Pval_GE = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$p.value,
                              Pearson_Corr_CN = cor.test(y = AUC, x = CN, method = "pearson", use = "complete.obs")$estimate,
                              Pearson_Pval_CN = cor.test(y = AUC, x = CN, method = "pearson", use = "complete.obs")$p.value,
                              Spearman_Corr_GE = cor.test(y = AUC, x = RPKM, method = "spearman", use = "complete.obs")$estimate,
                              Spearman_Pval_GE = cor.test(y = AUC, x = RPKM, method = "spearman", use = "complete.obs")$p.value,
                              Spearman_Corr_CN = cor.test(y = AUC, x = CN, method = "spearman", use = "complete.obs")$estimate,
                              Spearman_Pval_CN = cor.test(y = AUC, x = CN, method = "spearman", use = "complete.obs")$p.value)
gdsc_cor_res <- merge(g2p_druggene_tally, gdsc_cor_res, by = "Drug_Gene", all.y = TRUE)
gdsc_cor_res$n <- ifelse(is.na(gdsc_cor_res$n), 0, gdsc_cor_res$n)
gdsc_cor_res$InG2P <- ifelse(gdsc_cor_res$n != 0, "Yes", "No")
gdsc_cor_res$Dataset <- "GDSC"
gdsc_pearson_ttest_GE <- t.test(gdsc_cor_res$Pearson_Corr_GE ~ gdsc_cor_res$InG2P)
gdsc_pearson_ttest_res_GE <- paste0("t(", round(gdsc_pearson_ttest_GE$parameter, 0), ") = ", round(gdsc_pearson_ttest_GE$statistic, 2), ", p = ", round(gdsc_pearson_ttest_GE$p.value, 3))
gdsc_pearson_ttest_CN <- t.test(gdsc_cor_res$Pearson_Corr_CN ~ gdsc_cor_res$InG2P)
gdsc_pearson_ttest_res_CN <- paste0("t(", round(gdsc_pearson_ttest_CN$parameter, 0), ") = ", round(gdsc_pearson_ttest_CN$statistic, 2), ", p = ", round(gdsc_pearson_ttest_CN$p.value, 3))
gdsc_spearman_ttest_GE <- t.test(gdsc_cor_res$Spearman_Corr_GE ~ gdsc_cor_res$InG2P)
gdsc_spearman_ttest_res_GE <- paste0("t(", round(gdsc_spearman_ttest_GE$parameter, 0), ") = ", round(gdsc_spearman_ttest_GE$statistic, 2), ", p = ", round(gdsc_spearman_ttest_GE$p.value, 2))
gdsc_spearman_ttest_CN <- t.test(gdsc_cor_res$Spearman_Corr_CN ~ gdsc_cor_res$InG2P)
gdsc_spearman_ttest_res_CN <- paste0("t(", round(gdsc_spearman_ttest_CN$parameter, 0), ") = ", round(gdsc_spearman_ttest_CN$statistic, 2), ", p = ", round(gdsc_spearman_ttest_CN$p.value, 2))

# All
g2p_cor <- rbind(ccle_cor_res, ctrp_cor_res, gdsc_cor_res)
g2p_cor$Pearson_GE_Abs <- abs(g2p_cor$Pearson_Corr_GE)
g2p_cor$Pearson_GE_Sign <- ifelse(sign(g2p_cor$Pearson_Corr_GE) == 1, "+", ifelse(sign(g2p_cor$Pearson_Corr_GE) == -1, "-", "0"))
g2p_cor$Pearson_CN_Abs <- abs(g2p_cor$Pearson_Corr_CN)
g2p_cor$Pearson_CN_Sign <- ifelse(sign(g2p_cor$Pearson_Corr_CN) == 1, "+", ifelse(sign(g2p_cor$Pearson_Corr_CN) == -1, "-", "0"))
g2p_cor$Spearman_GE_Abs <- abs(g2p_cor$Spearman_Corr_GE)
g2p_cor$Spearman_GE_Sign <- ifelse(sign(g2p_cor$Spearman_Corr_GE) == 1, "+", ifelse(sign(g2p_cor$Spearman_Corr_GE) == -1, "-", "0"))
g2p_cor$Spearman_CN_Abs <- abs(g2p_cor$Spearman_Corr_CN)
g2p_cor$Spearman_CN_Sign <- ifelse(sign(g2p_cor$Spearman_Corr_CN) == 1, "+", ifelse(sign(g2p_cor$Spearman_Corr_CN) == -1, "-", "0"))
g2p_cor_text <- data.frame(Dataset = c("CCLE", "CTRP", "GDSC"),
                               Pearson_Test_Results_GE = c(ccle_pearson_ttest_res_GE, ctrp_pearson_ttest_res_GE, gdsc_pearson_ttest_res_GE),
                               Pearson_Test_Results_CN = c(ccle_pearson_ttest_res_CN, ctrp_pearson_ttest_res_CN, gdsc_pearson_ttest_res_CN),
                               Spearman_Test_Results_GE = c(ccle_spearman_ttest_res_GE, ctrp_spearman_ttest_res_GE, gdsc_spearman_ttest_res_GE),
                               Spearman_Test_Results_CN = c(ccle_spearman_ttest_res_CN, ctrp_spearman_ttest_res_CN, gdsc_spearman_ttest_res_CN))
```

# SUMMARIES

--------------------------------------------------------------------------------

## CRISPR: Mutation status counts

```{r}
crispr_data_g2p_summ <- unique(crispr_data_g2p[, c("Hugo_Symbol", "N_Nonsilent_Mutant", "N_Nonsilent_Wildtype")])
crispr_data_g2p_summ$Percent_Mutant <- crispr_data_g2p_summ$N_Nonsilent_Mutant / (crispr_data_g2p_summ$N_Nonsilent_Mutant + crispr_data_g2p_summ$N_Nonsilent_Wildtype) * 100
rownames(crispr_data_g2p_summ) <- NULL

# write.table(crispr_data_g2p_summ, file = "~/Desktop/crispr_data_g2p_summ_20181003.tsv", quote = FALSE, sep = "\t")

knitr::kable(crispr_data_g2p_summ, caption = "CRISPR: Level A G2P genes mutation status summary") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

## G2P tallies and summary stats

```{r}
print("Drug-Gene")
g2p_druggene_tally <- g2p_filt %>% group_by(Drug_Gene) %>% tally() %>% drop_na()
nrow(g2p_druggene_tally)
range(g2p_druggene_tally$n)
mean(g2p_druggene_tally$n)
sd(g2p_druggene_tally$n)

print("Drug")
g2p_drug_tally <- g2p_filt %>% group_by(Drug) %>% tally() %>% drop_na()
range(g2p_drug_tally$n)
mean(g2p_drug_tally$n)
sd(g2p_drug_tally$n)

print("Gene")
g2p_gene_tally <- g2p_filt %>% group_by(Gene) %>% tally() %>% drop_na()
range(g2p_gene_tally$n)
mean(g2p_gene_tally$n)
sd(g2p_gene_tally$n)

print("Drug-Gene-Lineage")
g2p_dgl_tally <- g2p_filt %>% group_by(Drug, Gene, Drug_Gene, group_general_lineage_name) %>% tally() %>% drop_na()
nrow(g2p_dgl_tally)
length(unique(g2p_dgl_tally$Drug))
length(unique(g2p_dgl_tally$Gene))
unique(g2p_filt$Phenotype.Description)
```

# PANCANCER

--------------------------------------------------------------------------------

## Mut

CRISPR mutation status plots of p-values:
```{r fig.height=1.5, fig.width=12}
crispr_signif_g2p_gene$y_label <- "CRISPR"
crispr_signif_g2p_gene$Hugo_Symbol <- factor(crispr_signif_g2p_gene$Hugo_Symbol, levels = g2p_genes)
crispr_signif_g2p_gene$p <- as.numeric(as.character(crispr_signif_g2p_gene$p))

crispr_mut_heatmap_plot <- ggplot(data = crispr_signif_g2p_gene, aes(x = Hugo_Symbol, y = y_label)) +
  geom_tile(aes(fill = p)) +
  scale_fill_gradientn(colors = c("darkgreen", "honeydew", "white"), values = c(0, 0.05, 1), na.value = "darksalmon") +
  scale_x_discrete(drop = FALSE) +
  theme(axis.title.y = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(x = "Gene")
crispr_mut_heatmap_plot

# write.table(crispr_signif_g2p_gene, file = "~/Desktop/crispr_signif_g2p_gene_20181203.csv", quote = FALSE, sep = ",", row.names = FALSE)
# ggsave("./plots/manuscript/crispr_mut_heatmap_plot_example.png", crispr_mut_heatmap_plot, device = "png", dpi = 450, width = 12, height = 1.5, units = "in")
```

```{r}
ccle_mut_summ <- merge(g2p_druggene_tally, ccle_signif_g2p_grid[, c("Drug_Gene", "p")], by = "Drug_Gene", all = TRUE)
ccle_mut_summ$n <- ifelse(is.na(ccle_mut_summ$n), 0, ccle_mut_summ$n)
ccle_mut_summ$InG2P <- ifelse(ccle_mut_summ$n != 0, "Yes", "No")
ccle_mut_summ$Dataset <- "CCLE"
ccle_p_ttest <- t.test(ccle_mut_summ$p ~ ccle_mut_summ$InG2P)
ccle_p_ttest_res <- paste0("t(", round(ccle_p_ttest$parameter, 0), ") = ", round(ccle_p_ttest$statistic, 2), ", p = ", round(ccle_p_ttest$p.value, 2))

ctrp_mut_summ <- merge(g2p_druggene_tally, ctrp_signif_g2p_grid[, c("Drug_Gene", "p")], by = "Drug_Gene", all = TRUE)
ctrp_mut_summ$n <- ifelse(is.na(ctrp_mut_summ$n), 0, ctrp_mut_summ$n)
ctrp_mut_summ$InG2P <- ifelse(ctrp_mut_summ$n != 0, "Yes", "No")
ctrp_mut_summ$Dataset <- "CTRP"
ctrp_p_ttest <- t.test(ctrp_mut_summ$p ~ ctrp_mut_summ$InG2P)
ctrp_p_ttest_res <- paste0("t(", round(ctrp_p_ttest$parameter, 0), ") = ", round(ctrp_p_ttest$statistic, 2), ", p = ", formatC(ctrp_p_ttest$p.value, digits = 2, format = "f"))

gdsc_mut_summ <- merge(g2p_druggene_tally, gdsc_signif_g2p_grid[, c("Drug_Gene", "p")], by = "Drug_Gene", all = TRUE)
gdsc_mut_summ$n <- ifelse(is.na(gdsc_mut_summ$n), 0, gdsc_mut_summ$n)
gdsc_mut_summ$InG2P <- ifelse(gdsc_mut_summ$n != 0, "Yes", "No")
gdsc_mut_summ$Dataset <- "GDSC"
gdsc_p_ttest <- t.test(gdsc_mut_summ$p ~ gdsc_mut_summ$InG2P)
gdsc_p_ttest_res <- paste0("t(", round(gdsc_p_ttest$parameter, 0), ") = ", round(gdsc_p_ttest$statistic, 2), ", p = ", formatC(gdsc_p_ttest$p.value, digits = 2, format = "f"))

g2p_mut_all_summ <- rbind(ccle_mut_summ, ctrp_mut_summ, gdsc_mut_summ)
g2p_mut_all_summ$p_log10 <- -log10(g2p_mut_all_summ$p)
g2p_mut_all_summ_text <- data.frame(Dataset = c("CCLE", "CTRP", "GDSC"), Test_Results_p = c(ccle_p_ttest_res, ctrp_p_ttest_res, gdsc_p_ttest_res))
g2p_mut_all_summ$Drug_Gene_Ordered <- factor(g2p_mut_all_summ$Drug_Gene, levels = as.character(unique(g2p_mut_all_summ$Drug_Gene)), labels = sapply(strsplit(x = as.character(unique(g2p_mut_all_summ$Drug_Gene)), split = "_"), function(x) paste0(x[1], "\n", x[2])))
```

```{r fig.height=5, fig.width=12}
plot_p_point <- ggplot(data = g2p_mut_all_summ, aes(x = p, y = n, label = Drug_Gene)) +
  facet_wrap(~ Dataset, scales = "free_x") +
  geom_point(mapping = aes(color = InG2P), alpha = 0.5) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_y_continuous(breaks = seq(0, 120, by = 10), labels = seq(0, 120, by = 10)) +
  geom_text_repel(data = subset(g2p_mut_all_summ, InG2P == "Yes" & n > 10), size = 3, segment.size = 0.2, segment.color = "grey50") +
  geom_text(data = g2p_mut_all_summ_text, mapping = aes(x = 0.5, y = 115, label = Test_Results_p)) +
  labs(x = "Gene-drug association p-value between mutation groups", y = "Number of level A G2P associations") +
  theme(legend.position = "none", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank())

plot_p_box <- ggplot(data = g2p_mut_all_summ) +
  facet_wrap(~ Dataset, scales = "free_x") +
  geom_violin(mapping = aes(color = InG2P, fill = InG2P, y = p, x = InG2P), alpha = 0.35) +
  geom_boxplot(mapping = aes(color = InG2P, y = p, x = InG2P), width = 0.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  coord_flip() +
  labs(x = "In G2P?") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

plot_p_point <- ggplot_gtable(ggplot_build(plot_p_point))
plot_p_box <- ggplot_gtable(ggplot_build(plot_p_box))
p_maxWidth <-  grid::unit.pmax(plot_p_point$widths, plot_p_box$widths)
plot_p_point$widths <- as.list(p_maxWidth)
plot_p_box$widths <- as.list(p_maxWidth)
g2p_mut_plot <- cowplot::plot_grid(plot_p_box,
                   plot_p_point,
                   nrow = 2,
                   rel_heights = c(1, 4))
g2p_mut_plot

# ggsave("./plots/manuscript/g2p_mut_plot.png", g2p_mut_plot, device = "png", dpi = 450, width = 12, height = 5, units = "in")
```

```{r, eval=FALSE}
plot_p_point <- ggplot(data = g2p_mut_all_summ) +
  facet_wrap(~ Dataset, scales = "free_x") +
  geom_point(mapping = aes(label = Drug_Gene, color = InG2P, x = p, y = n), alpha = 0.5) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_y_continuous(breaks = seq(0, 120, by = 10), labels = seq(0, 120, by = 10)) +
  geom_text(data = g2p_mut_all_summ_text, mapping = aes(x = 0.5, y = 115, label = Test_Results_p)) +
  labs(x = "Gene-drug association p-value between mutation groups", y = "Number of level A G2P associations") +
  theme(legend.position = "none", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank())

plot_p_box <- ggplot(data = g2p_mut_all_summ) +
  facet_wrap(~ Dataset, scales = "free_x") +
  geom_violin(mapping = aes(color = InG2P, y = p, x = InG2P), alpha = 0.5) +
  geom_boxplot(mapping = aes(color = InG2P, y = p, x = InG2P), alpha = 0.75, width = 0.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  coord_flip() +
  labs(x = "Level A G2P\nassociation") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

plot <- subplot(plot_p_box, plot_p_point, nrows = 2, titleX = TRUE, titleY = TRUE) %>% layout(margin = list(l = 75), xaxis5 = list(title = "Gene-drug association p-value between mutation groups"), yaxis = list(title = "In G2P?", domain = c(0.8, 1)), yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))

saveWidgetFix(widget = plot, file = "./plots/manuscript/g2p_mut_plot.html")
```

## GE

### Pearson

```{r fig.height=5, fig.width=12}
plot_ge_pearson_point <- ggplot(data = g2p_cor, aes(x = Pearson_GE_Abs, y = n, label = Drug_Gene)) +
  facet_wrap(~ Dataset) +
  geom_point(mapping = aes(fill = InG2P, color = InG2P, shape = Pearson_GE_Sign), alpha = 0.5, size = 1.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_y_continuous(breaks = seq(0, 120, by = 10), labels = seq(0, 120, by = 10)) +
  scale_x_continuous(breaks = seq(0, 0.65, by = 0.1), labels = seq(0, 0.65, by = 0.1)) +
  coord_cartesian(xlim = c(0, 0.65)) +
  scale_shape_manual(values = c(25, 24), labels = c("Negative", "Positive")) +
  geom_text_repel(data = subset(g2p_cor, InG2P == "Yes" & n > 10), size = 3, segment.size = 0.2, segment.color = "grey50") +
  geom_text(data = g2p_cor_text, mapping = aes(x = 0.325, y = 115, label = Pearson_Test_Results_GE)) +
  labs(x = "Pearson correlation coefficient (r) between drug-gene AUC and gene expression", y = "Number of level A G2P associations") +
  theme(legend.position = "bottom", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank()) +
  guides(color = FALSE, fill = FALSE, shape = FALSE)

plot_ge_pearson_box <- ggplot(data = g2p_cor) +
  facet_wrap(~ Dataset) +
  geom_violin(mapping = aes(color = InG2P, fill = InG2P, y = Pearson_GE_Abs, x = InG2P), alpha = 0.35) +
  geom_boxplot(mapping = aes(color = InG2P, y = Pearson_GE_Abs, x = InG2P), width = 0.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_y_continuous(breaks = seq(0, 0.65, by = 0.1), labels = seq(0, 0.65, by = 0.1)) +
  coord_flip(ylim = c(0, 0.65)) +
  labs(x = "In G2P?") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

plot_ge_pearson_point <- ggplot_gtable(ggplot_build(plot_ge_pearson_point))
plot_ge_pearson_box <- ggplot_gtable(ggplot_build(plot_ge_pearson_box))
p_maxWidth <-  grid::unit.pmax(plot_ge_pearson_point$widths[2:3], plot_ge_pearson_box$widths[2:3])
plot_ge_pearson_point$widths[2:3] <- p_maxWidth
plot_ge_pearson_box$widths[2:3] <- p_maxWidth

g2p_ge_pearson_plot <- cowplot::plot_grid(plot_ge_pearson_box, plot_ge_pearson_point, nrow = 2, rel_heights = c(1, 4))
g2p_ge_pearson_plot

# ggsave("./plots/manuscript/g2p_ge_pearson_plot.png", g2p_ge_pearson_plot, device = "png", dpi = 450, width = 12, height = 5, units = "in")
```

```{r, eval=FALSE}
plot_ge_point <- ggplot(data = g2p_cor) +
  facet_wrap(~ Dataset) +
  geom_point(mapping = aes(label = Drug_Gene, fill = InG2P, color = InG2P, shape = Pearson_GE_Sign, x = Pearson_GE_Abs, y = n, size = Pearson_Pval_GE), alpha = 0.5, size = 1.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_shape_manual(values = c(25, 24), labels = c("Negative", "Positive")) +
  scale_y_continuous(breaks = seq(0, 120, by = 10), labels = seq(0, 120, by = 10)) +
  scale_x_continuous(breaks = pretty_breaks()) +
  geom_text(data = g2p_cor_text, mapping = aes(x = 0.3, y = 115, label = Pearson_Test_Results_GE)) +
  labs(x = "Pearson correlation coefficient (r) between drug-gene AUC and gene expression", y = "Number of level A G2P associations") +
  theme(legend.position = "bottom", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank()) +
  guides(color = FALSE, fill = FALSE) +
  labs(shape = NULL)

plot_ge_box <- ggplot(data = g2p_cor) +
  facet_wrap(~ Dataset) +
  geom_violin(mapping = aes(color = InG2P, y = Pearson_GE_Abs, x = InG2P), alpha = 0.5) +
  geom_boxplot(mapping = aes(color = InG2P, y = Pearson_GE_Abs, x = InG2P), alpha = 0.5, width = 0.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  coord_flip() +
  labs(x = "In G2P?") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

ge_pearson_plot <- subplot(plot_ge_box, plot_ge_point, nrows = 2, heights = c(0.2, 0.8), titleX = TRUE, titleY = TRUE) %>% layout(margin = list(l = 75), xaxis5 = list(title = "Pearson correlation coefficient between drug-gene AUC and gene expression"), yaxis = list(title = "In G2P?", domain = c(0.8, 1)), yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))

saveWidgetFix(widget = ge_pearson_plot, file = "./plots/manuscript/g2p_ge_pearson_plot.html")
```

### Spearman

```{r fig.height=5, fig.width=12}
plot_ge_spearman_point <- ggplot(data = g2p_cor, aes(x = Spearman_GE_Abs, y = n, label = Drug_Gene)) +
  facet_wrap(~ Dataset) +
  geom_point(mapping = aes(fill = InG2P, color = InG2P, shape = Spearman_GE_Sign), alpha = 0.5, size = 1.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_y_continuous(breaks = seq(0, 120, by = 10), labels = seq(0, 120, by = 10)) +
  scale_x_continuous(breaks = seq(0, 0.65, by = 0.1), labels = seq(0, 0.65, by = 0.1)) +
  coord_cartesian(xlim = c(0, 0.6)) +
  scale_shape_manual(values = c(25, 24), labels = c("Negative", "Positive")) +
  geom_text_repel(data = subset(g2p_cor, InG2P == "Yes" & n > 10), size = 3, segment.size = 0.2, segment.color = "grey50") +
  geom_text(data = g2p_cor_text, mapping = aes(x = 0.3, y = 115, label = Spearman_Test_Results_GE)) +
  labs(x = "Spearman correlation coefficient (r_s) between drug-gene AUC and gene expression", y = "Number of level A G2P associations") +
  theme(legend.position = "bottom", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank()) +
  guides(color = FALSE, fill = FALSE, shape = FALSE)

plot_ge_spearman_box <- ggplot(data = g2p_cor) +
  facet_wrap(~ Dataset) +
  geom_violin(mapping = aes(color = InG2P, fill = InG2P, y = Spearman_GE_Abs, x = InG2P), alpha = 0.35) +
  geom_boxplot(mapping = aes(color = InG2P, y = Spearman_GE_Abs, x = InG2P), width = 0.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_y_continuous(breaks = seq(0, 0.65, by = 0.1), labels = seq(0, 0.65, by = 0.1)) +
  coord_flip(ylim = c(0, 0.6)) +
  labs(x = "In G2P?") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

plot_ge_spearman_point <- ggplot_gtable(ggplot_build(plot_ge_spearman_point))
plot_ge_spearman_box <- ggplot_gtable(ggplot_build(plot_ge_spearman_box))
p_maxWidth <-  grid::unit.pmax(plot_ge_spearman_point$widths[2:3], plot_ge_spearman_box$widths[2:3])
plot_ge_spearman_point$widths[2:3] <- p_maxWidth
plot_ge_spearman_box$widths[2:3] <- p_maxWidth

g2p_ge_spearman_plot <- cowplot::plot_grid(plot_ge_spearman_box, plot_ge_spearman_point, nrow = 2, rel_heights = c(1, 4))
g2p_ge_spearman_plot

# ggsave("./plots/manuscript/g2p_ge_spearman_plot.png", g2p_ge_spearman_plot, device = "png", dpi = 450, width = 12, height = 5, units = "in")
```

```{r, eval=FALSE}
plot_ge_point <- ggplot(data = g2p_cor) +
  facet_wrap(~ Dataset) +
  geom_point(mapping = aes(label = Drug_Gene, fill = InG2P, color = InG2P, shape = Spearman_GE_Sign, x = Spearman_GE_Abs, y = n, size = Spearman_Pval_GE), alpha = 0.5, size = 1.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_shape_manual(values = c(25, 24), labels = c("Negative", "Positive")) +
  scale_y_continuous(breaks = seq(0, 120, by = 10), labels = seq(0, 120, by = 10)) +
  scale_x_continuous(breaks = pretty_breaks()) +
  geom_text(data = g2p_cor_text, mapping = aes(x = 0.3, y = 115, label = Spearman_Test_Results_GE)) +
  labs(x = "Spearman correlation coefficient (r_s) between drug-gene AUC and gene expression", y = "Number of level A G2P associations") +
  theme(legend.position = "bottom", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank()) +
  guides(color = FALSE, fill = FALSE) +
  labs(shape = NULL)

plot_ge_box <- ggplot(data = g2p_cor) +
  facet_wrap(~ Dataset) +
  geom_violin(mapping = aes(color = InG2P, y = Spearman_GE_Abs, x = InG2P), alpha = 0.5) +
  geom_boxplot(mapping = aes(color = InG2P, y = Spearman_GE_Abs, x = InG2P), alpha = 0.5, width = 0.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  coord_flip() +
  labs(x = "In G2P?") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

ge_spearman_plot <- subplot(plot_ge_box, plot_ge_point, nrows = 2, heights = c(0.2, 0.8), titleX = TRUE, titleY = TRUE) %>% layout(margin = list(l = 75), xaxis5 = list(title = "Spearman correlation coefficient between drug-gene AUC and gene expression"), yaxis = list(title = "In G2P?", domain = c(0.8, 1)), yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))

saveWidgetFix(widget = ge_spearman_plot, file = "./plots/manuscript/g2p_ge_spearman_plot.html")
```

## CN

### Pearson

```{r fig.height=5, fig.width=12}
plot_cn_pearson_point <- ggplot(data = g2p_cor, aes(x = Pearson_CN_Abs, y = n, label = Drug_Gene)) +
  facet_wrap(~ Dataset) +
  geom_point(mapping = aes(fill = InG2P, color = InG2P, shape = Pearson_CN_Sign), alpha = 0.5, size = 1.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_y_continuous(breaks = seq(0, 120, by = 10), labels = seq(0, 120, by = 10)) +
  scale_x_continuous(breaks = seq(0, 0.65, by = 0.1), labels = seq(0, 0.65, by = 0.1)) +
  coord_cartesian(xlim = c(0, 0.5)) +
  scale_shape_manual(values = c(25, 24), labels = c("Negative", "Positive")) +
  geom_text_repel(data = subset(g2p_cor, InG2P == "Yes" & n > 10), size = 3, segment.size = 0.2, segment.color = "grey50") +
  geom_text(data = g2p_cor_text, mapping = aes(x = 0.25, y = 115, label = Pearson_Test_Results_CN)) +
  labs(x = "Pearson correlation coefficient (r) between drug-gene AUC and copy number", y = "Number of level A G2P associations") +
  theme(legend.position = "bottom", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank()) +
  guides(color = FALSE, fill = FALSE, shape = FALSE)

plot_cn_pearson_box <- ggplot(data = g2p_cor) +
  facet_wrap(~ Dataset) +
  geom_violin(mapping = aes(color = InG2P, fill = InG2P, y = Pearson_CN_Abs, x = InG2P), alpha = 0.35) +
  geom_boxplot(mapping = aes(color = InG2P, y = Pearson_CN_Abs, x = InG2P), width = 0.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_y_continuous(breaks = seq(0, 0.65, by = 0.1), labels = seq(0, 0.65, by = 0.1)) +
  coord_flip(ylim = c(0, 0.5)) +
  labs(x = "In G2P?") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

plot_cn_pearson_point <- ggplot_gtable(ggplot_build(plot_cn_pearson_point))
plot_cn_pearson_box <- ggplot_gtable(ggplot_build(plot_cn_pearson_box))
p_maxWidth <-  grid::unit.pmax(plot_cn_pearson_point$widths[2:3], plot_cn_pearson_box$widths[2:3])
plot_cn_pearson_point$widths[2:3] <- p_maxWidth
plot_cn_pearson_box$widths[2:3] <- p_maxWidth

g2p_cn_pearson_plot <- cowplot::plot_grid(plot_cn_pearson_box, plot_cn_pearson_point, nrow = 2, rel_heights = c(1, 4))
g2p_cn_pearson_plot

# ggsave("./plots/manuscript/g2p_cn_pearson_plot.png", g2p_cn_pearson_plot, device = "png", dpi = 450, width = 12, height = 5, units = "in")
```

```{r, eval=FALSE}
plot_cn_point <- ggplot(data = g2p_cor) +
  facet_wrap(~ Dataset) +
  geom_point(mapping = aes(label = Drug_Gene, fill = InG2P, color = InG2P, shape = Pearson_CN_Sign, x = Pearson_CN_Abs, y = n, size = Pearson_Pval_CN), alpha = 0.5, size = 1.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_shape_manual(values = c(25, 24), labels = c("Negative", "Positive")) +
  scale_y_continuous(breaks = seq(0, 120, by = 10), labels = seq(0, 120, by = 10)) +
  scale_x_continuous(breaks = pretty_breaks()) +
  geom_text(data = g2p_cor_text, mapping = aes(x = 0.3, y = 115, label = Pearson_Test_Results_CN)) +
  labs(x = "Pearson correlation coefficient (r) between drug-gene AUC and copy number", y = "Number of level A G2P associations") +
  theme(legend.position = "bottom", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank()) +
  guides(color = FALSE, fill = FALSE) +
  labs(shape = NULL)

plot_cn_box <- ggplot(data = g2p_cor) +
  facet_wrap(~ Dataset) +
  geom_violin(mapping = aes(color = InG2P, y = Pearson_CN_Abs, x = InG2P), alpha = 0.5) +
  geom_boxplot(mapping = aes(color = InG2P, y = Pearson_CN_Abs, x = InG2P), alpha = 0.5, width = 0.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  coord_flip() +
  labs(x = "In G2P?") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

cn_pearson_plot <- subplot(plot_cn_box, plot_cn_point, nrows = 2, heights = c(0.2, 0.8), titleX = TRUE, titleY = TRUE) %>% layout(margin = list(l = 75), xaxis5 = list(title = "Pearson correlation coefficient between drug-gene AUC and copy number"), yaxis = list(title = "In G2P?", domain = c(0.8, 1)), yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))

saveWidgetFix(widget = ge_pearson_plot, file = "./plots/manuscript/g2p_cn_pearson_plot.html")
```

### Spearman

```{r fig.height=5, fig.width=12}
plot_cn_spearman_point <- ggplot(data = g2p_cor, aes(x = Spearman_CN_Abs, y = n, label = Drug_Gene)) +
  facet_wrap(~ Dataset) +
  geom_point(mapping = aes(fill = InG2P, color = InG2P, shape = Spearman_CN_Sign), alpha = 0.5, size = 1.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_y_continuous(breaks = seq(0, 120, by = 10), labels = seq(0, 120, by = 10)) +
  scale_x_continuous(breaks = seq(0, 0.4, by = 0.05), labels = seq(0, 0.4, by = 0.05)) +
  coord_cartesian(xlim = c(0, 0.35)) +
  scale_shape_manual(values = c(25, 24), labels = c("Negative", "Positive")) +
  geom_text_repel(data = subset(g2p_cor, InG2P == "Yes" & n > 10), size = 3, segment.size = 0.2, segment.color = "grey50") +
  geom_text(data = g2p_cor_text, mapping = aes(x = 0.175, y = 115, label = Spearman_Test_Results_CN)) +
  labs(x = "Spearman correlation coefficient (r_s) between drug-gene AUC and copy number", y = "Number of level A G2P associations") +
  theme(legend.position = "bottom", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank()) +
  guides(color = FALSE, fill = FALSE, shape = FALSE)

plot_cn_spearman_box <- ggplot(data = g2p_cor) +
  facet_wrap(~ Dataset) +
  geom_violin(mapping = aes(color = InG2P, fill = InG2P, y = Spearman_CN_Abs, x = InG2P), alpha = 0.35) +
  geom_boxplot(mapping = aes(color = InG2P, y = Spearman_CN_Abs, x = InG2P), width = 0.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_y_continuous(breaks = seq(0, 0.4, by = 0.05), labels = seq(0, 0.4, by = 0.05)) +
  coord_flip(ylim = c(0, 0.35)) +
  labs(x = "In G2P?") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

plot_cn_spearman_point <- ggplot_gtable(ggplot_build(plot_cn_spearman_point))
plot_cn_spearman_box <- ggplot_gtable(ggplot_build(plot_cn_spearman_box))
p_maxWidth <-  grid::unit.pmax(plot_cn_spearman_point$widths[2:3], plot_cn_spearman_box$widths[2:3])
plot_cn_spearman_point$widths[2:3] <- p_maxWidth
plot_cn_spearman_box$widths[2:3] <- p_maxWidth

g2p_cn_spearman_plot <- cowplot::plot_grid(plot_cn_spearman_box, plot_cn_spearman_point, nrow = 2, rel_heights = c(1, 4))
g2p_cn_spearman_plot

# ggsave("./plots/manuscript/g2p_cn_spearman_plot.png", g2p_cn_spearman_plot, device = "png", dpi = 450, width = 12, height = 5, units = "in")
```

```{r, eval=FALSE}
plot_cn_point <- ggplot(data = g2p_cor) +
  facet_wrap(~ Dataset) +
  geom_point(mapping = aes(label = Drug_Gene, fill = InG2P, color = InG2P, shape = Spearman_CN_Sign, x = Spearman_CN_Abs, y = n, size = Spearman_Pval_CN), alpha = 0.5, size = 1.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_shape_manual(values = c(25, 24), labels = c("Negative", "Positive")) +
  scale_y_continuous(breaks = seq(0, 120, by = 10), labels = seq(0, 120, by = 10)) +
  scale_x_continuous(breaks = pretty_breaks()) +
  geom_text(data = g2p_cor_text, mapping = aes(x = 0.175, y = 115, label = Spearman_Test_Results_CN)) +
  labs(x = "Spearman correlation coefficient (r_s) between drug-gene AUC and copy number", y = "Number of level A G2P associations") +
  theme(legend.position = "bottom", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank()) +
  guides(color = FALSE, fill = FALSE) +
  labs(shape = NULL)

plot_cn_box <- ggplot(data = g2p_cor) +
  facet_wrap(~ Dataset) +
  geom_violin(mapping = aes(color = InG2P, y = Spearman_CN_Abs, x = InG2P), alpha = 0.5) +
  geom_boxplot(mapping = aes(color = InG2P, y = Spearman_CN_Abs, x = InG2P), alpha = 0.5, width = 0.25) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  coord_flip() +
  labs(x = "In G2P?") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

cn_spearman_plot <- subplot(plot_cn_box, plot_cn_point, nrows = 2, heights = c(0.2, 0.8), titleX = TRUE, titleY = TRUE) %>% layout(margin = list(l = 75), xaxis5 = list(title = "Spearman correlation coefficient between drug-gene AUC and copy number"), yaxis = list(title = "In G2P?", domain = c(0.8, 1)), yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))

saveWidgetFix(widget = cn_spearman_plot, file = "./plots/manuscript/g2p_cn_spearman_plot.html")
```

# BY LINEAGE

--------------------------------------------------------------------------------

## Mut: Grid

Summarize the Wilcoxon test p-values of the grid tests by mutation status. Grid tests were all lineage-agnostic G2P drug-gene associations tested across all screened lineages in each dataset:
```{r}
g2p_dgl_mut_tally <- g2p %>% group_by(Drug_Gene, group_general_lineage_name) %>% tally()

ccle_dgl_summ <- ccle_signif_grid_lineage
ccle_dgl_summ$Drug_Gene <- paste0(ccle_dgl_summ$Drug, "_", ccle_dgl_summ$Hugo_Symbol)
ccle_dgl_summ <- merge(g2p_dgl_mut_tally, ccle_dgl_summ[, c("Drug_Gene", "group_general_lineage_name", "p")], by = c("Drug_Gene", "group_general_lineage_name"), all.y = TRUE)
ccle_dgl_summ$Drug_Gene_Lin <- paste0(ccle_dgl_summ$Drug_Gene, "_", ccle_dgl_summ$group_general_lineage_name)
ccle_dgl_summ$n <- ifelse(is.na(ccle_dgl_summ$n), 0, ccle_dgl_summ$n)
ccle_dgl_summ$InG2P <- ifelse(ccle_dgl_summ$n != 0, "Yes", "No")
ccle_dgl_summ$Dataset <- "CCLE"

ctrp_dgl_summ <- ctrp_signif_grid_lineage
ctrp_dgl_summ$Drug_Gene <- paste0(ctrp_dgl_summ$Drug, "_", ctrp_dgl_summ$Hugo_Symbol)
ctrp_dgl_summ <- merge(g2p_dgl_mut_tally, ctrp_dgl_summ[, c("Drug_Gene", "group_general_lineage_name", "p")], by = c("Drug_Gene", "group_general_lineage_name"), all.y = TRUE)
ctrp_dgl_summ$Drug_Gene_Lin <- paste0(ctrp_dgl_summ$Drug_Gene, "_", ctrp_dgl_summ$group_general_lineage_name)
ctrp_dgl_summ$n <- ifelse(is.na(ctrp_dgl_summ$n), 0, ctrp_dgl_summ$n)
ctrp_dgl_summ$InG2P <- ifelse(ctrp_dgl_summ$n != 0, "Yes", "No")
ctrp_dgl_summ$Dataset <- "CTRP"

gdsc_dgl_summ <- gdsc_signif_grid_lineage
gdsc_dgl_summ$Drug_Gene <- paste0(gdsc_dgl_summ$Drug, "_", gdsc_dgl_summ$Hugo_Symbol)
gdsc_dgl_summ <- merge(g2p_dgl_mut_tally, gdsc_dgl_summ[, c("Drug_Gene", "group_general_lineage_name", "p")], by = c("Drug_Gene", "group_general_lineage_name"), all.y = TRUE)
gdsc_dgl_summ$Drug_Gene_Lin <- paste0(gdsc_dgl_summ$Drug_Gene, "_", gdsc_dgl_summ$group_general_lineage_name)
gdsc_dgl_summ$n <- ifelse(is.na(gdsc_dgl_summ$n), 0, gdsc_dgl_summ$n)
gdsc_dgl_summ$InG2P <- ifelse(gdsc_dgl_summ$n != 0, "Yes", "No")
gdsc_dgl_summ$Dataset <- "GDSC"

dgl_mut_grid_summ <- rbind(ccle_dgl_summ, ctrp_dgl_summ, gdsc_dgl_summ)
```

Make summary plots of mutation status tests by lineage and return a dataframe of all t-test results from comparing the p-value distributions of G2P and non-G2P drug-gene associaions:
```{r, eval=FALSE, results = 'hide'}
dgl_mut_g2p_grid_ttest_res <- lapply(as.character(unique(dgl_mut_grid_summ$group_general_lineage_name)), MutPlotsByLineage)
dgl_mut_g2p_grid_ttest_res_list <- rbindlist(dgl_mut_g2p_grid_ttest_res, use.names = TRUE)
dgl_mut_g2p_grid_ttest_res_list <- dgl_mut_g2p_grid_ttest_res_list[order(dgl_mut_g2p_grid_ttest_res_list$p_value),]
saveRDS(dgl_mut_g2p_grid_ttest_res_list, "./data_munging/rds/dgl_mut_g2p_grid_ttest_res_list.rds")
```

```{r}
dgl_mut_g2p_grid_ttest_res_list <- readRDS("./data_munging/rds/dgl_mut_g2p_grid_ttest_res_list.rds")
knitr::kable(dgl_mut_g2p_grid_ttest_res_list, caption = "Mutation status: T-test results for level A G2P drug-gene associations across lineages") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

## GE/CN: G2P

Drug-gene-lineage (DGL) overlaps between CCLE, CTRP, and GDSC:
```{r}
dgl_overlap_test_list <- list(CCLE = unique(ccle_data_g2p_dgl$Drug_Gene_Lin),
                              CTRP = unique(ctrp_data_g2p_dgl$Drug_Gene_Lin),
                              GDSC = unique(gdsc_data_g2p_dgl$Drug_Gene_Lin))

ItemsList <- venn(dgl_overlap_test_list, show.plot = TRUE)
ItemsList
```

Test the 102 level-A G2P DGL combinations:
```{r, eval=FALSE}
ccle_lineages4pearson <- as.character(unique(ccle_data_g2p_dgl$group_general_lineage_name))[!is.na(as.character(unique(ccle_data_g2p_dgl$group_general_lineage_name)))]
ccle_signif_ge <- lapply(ccle_lineages4pearson, PearsonByLineage, variable = "GE", dataset = ccle_data_g2p_dgl)
ccle_signif_ge_all_lin <- rbindlist(ccle_signif_ge, use.names = TRUE)
ccle_signif_ge_all_lin$Dataset <- "CCLE"
ccle_signif_cn <- lapply(ccle_lineages4pearson, PearsonByLineage, variable = "CN", dataset = ccle_data_g2p_dgl)
ccle_signif_cn_all_lin <- rbindlist(ccle_signif_cn, use.names = TRUE)
ccle_signif_cn_all_lin$Dataset <- "CCLE"
ccle_signif_all_lin <- merge(ccle_signif_ge_all_lin, ccle_signif_cn_all_lin, by = c("Drug_Gene", "group_general_lineage_name", "n", "Dataset"), all = TRUE)

ctrp_lineages4pearson <- as.character(unique(ctrp_data_g2p_dgl$group_general_lineage_name))[!is.na(as.character(unique(ctrp_data_g2p_dgl$group_general_lineage_name)))]
ctrp_signif_ge <- lapply(ctrp_lineages4pearson, PearsonByLineage, variable = "GE", dataset = ctrp_data_g2p_dgl)
ctrp_signif_ge_all_lin <- rbindlist(ctrp_signif_ge, use.names = TRUE)
ctrp_signif_ge_all_lin$Dataset <- "CTRP"
ctrp_signif_cn <- lapply(ctrp_lineages4pearson, PearsonByLineage, variable = "CN", dataset = ctrp_data_g2p_dgl)
ctrp_signif_cn_all_lin <- rbindlist(ctrp_signif_cn, use.names = TRUE)
ctrp_signif_cn_all_lin$Dataset <- "CTRP"
ctrp_signif_all_lin <- merge(ctrp_signif_ge_all_lin, ctrp_signif_cn_all_lin, by = c("Drug_Gene", "group_general_lineage_name", "n", "Dataset"), all = TRUE)

gdsc_lineages4pearson <- as.character(unique(gdsc_data_g2p_dgl$group_general_lineage_name))[!is.na(as.character(unique(gdsc_data_g2p_dgl$group_general_lineage_name)))]
gdsc_signif_ge <- lapply(gdsc_lineages4pearson, PearsonByLineage, variable = "GE", dataset = gdsc_data_g2p_dgl)
gdsc_signif_ge_all_lin <- rbindlist(gdsc_signif_ge, use.names = TRUE)
gdsc_signif_ge_all_lin$Dataset <- "GDSC"
gdsc_signif_cn <- lapply(gdsc_lineages4pearson, PearsonByLineage, variable = "CN", dataset = gdsc_data_g2p_dgl)
gdsc_signif_cn_all_lin <- rbindlist(gdsc_signif_cn, use.names = TRUE)
gdsc_signif_cn_all_lin$Dataset <- "GDSC"
gdsc_signif_all_lin <- merge(gdsc_signif_ge_all_lin, gdsc_signif_cn_all_lin, by = c("Drug_Gene", "group_general_lineage_name", "n", "Dataset"), all = TRUE)

signif_all_lin <- rbind(ccle_signif_all_lin, ctrp_signif_all_lin, gdsc_signif_all_lin)
signif_all_lin$n <- ifelse(is.na(signif_all_lin$n), 0, signif_all_lin$n)
signif_all_lin$InG2P <- ifelse(signif_all_lin$n != 0, "Yes", "No")
signif_all_lin$Drug_Gene_Lin <- paste0(signif_all_lin$Drug_Gene, "_", signif_all_lin$group_general_lineage_name)
saveRDS(signif_all_lin, "./data_munging/rds/dgl_signif_pearson_g2p.rds")
```

```{r}
dgl_signif_pearson_g2p <- readRDS("./data_munging/rds/dgl_signif_pearson_g2p.rds")
```

```{r}
# CCLE mut, n = 7
ccle_mut <- filter(select(ccle_signif_g2p_lineage, Drug_Gene_Lin, p), p < 0.05)
ccle_mut_n <- nrow(select(ccle_signif_g2p_lineage, Drug_Gene_Lin, p))
# CCLE GE/CN, n = 9
ccle_ge <- select(subset(filter(dgl_signif_pearson_g2p, Dataset == "CCLE"), Pearson_Pval_GE < 0.05), Drug_Gene_Lin, Pearson_Pval_GE, Pearson_Corr_GE)
ccle_cn <- select(subset(filter(dgl_signif_pearson_g2p, Dataset == "CCLE"), Pearson_Pval_CN < 0.05), Drug_Gene_Lin, Pearson_Pval_CN, Pearson_Corr_CN)
ccle_Pear_n <- nrow(filter(dgl_signif_pearson_g2p, Dataset == "CCLE"))
ccle_dgl_summ <- merge(ccle_mut, merge(ccle_ge, ccle_cn, by = "Drug_Gene_Lin", all = TRUE), all = TRUE)
ccle_dgl_summ <- with(ccle_dgl_summ, cbind(colsplit(Drug_Gene_Lin, pattern = "_", names = c("Drug", "Gene", "Lineage")), 
                                           CCLE_p = p,
                                           CCLE_p_Bonf = ifelse(p < 0.05 / ccle_mut_n, "Sig", NA),
                                           CCLE_r_GE = Pearson_Corr_GE,
                                           CCLE_p_GE = Pearson_Pval_GE,
                                           CCLE_p_GE_Bonf = ifelse(Pearson_Pval_GE < 0.05 / ccle_Pear_n, "Sig", NA),
                                           CCLE_r_CN = Pearson_Corr_CN,
                                           CCLE_p_CN = Pearson_Pval_CN,
                                           CCLE_p_CN_Bonf = ifelse(Pearson_Pval_CN < 0.05 / ccle_Pear_n, "Sig", NA)))

# CTRP mut, n = 48
ctrp_mut <- filter(select(ctrp_signif_g2p_lineage, Drug_Gene_Lin, p), p < 0.05)
ctrp_mut_n <- nrow(select(ctrp_signif_g2p_lineage, Drug_Gene_Lin, p))
# CTRP GE/CN, n = 61
ctrp_ge <- select(subset(filter(dgl_signif_pearson_g2p, Dataset == "CTRP"), Pearson_Pval_GE < 0.05), Drug_Gene_Lin, Pearson_Pval_GE, Pearson_Corr_GE)
ctrp_cn <- select(subset(filter(dgl_signif_pearson_g2p, Dataset == "CTRP"), Pearson_Pval_CN < 0.05), Drug_Gene_Lin, Pearson_Pval_CN, Pearson_Corr_CN)
ctrp_Pear_n <- nrow(filter(dgl_signif_pearson_g2p, Dataset == "CTRP"))
ctrp_dgl_summ <- merge(ctrp_mut, merge(ctrp_ge, ctrp_cn, by = "Drug_Gene_Lin", all = TRUE), all = TRUE)
ctrp_dgl_summ <- with(ctrp_dgl_summ, cbind(colsplit(Drug_Gene_Lin, pattern = "_", names = c("Drug", "Gene", "Lineage")), 
                                           CTRP_p = p,
                                           CTRP_p_Bonf = ifelse(p < 0.05 / ctrp_mut_n, "Sig", NA),
                                           CTRP_r_GE = Pearson_Corr_GE,
                                           CTRP_p_GE = Pearson_Pval_GE,
                                           CTRP_p_GE_Bonf = ifelse(Pearson_Pval_GE < 0.05 / ctrp_Pear_n, "Sig", NA),
                                           CTRP_r_CN = Pearson_Corr_CN,
                                           CTRP_p_CN = Pearson_Pval_CN,
                                           CTRP_p_CN_Bonf = ifelse(Pearson_Pval_CN < 0.05 / ctrp_Pear_n, "Sig", NA)))

# GDSC mut, n = 39
gdsc_mut <- filter(select(gdsc_signif_g2p_lineage, Drug_Gene_Lin, p), p < 0.05)
gdsc_mut_n <- nrow(select(gdsc_signif_g2p_lineage, Drug_Gene_Lin, p))
# GDSC GE/CN, n = 53
gdsc_ge <- select(subset(filter(dgl_signif_pearson_g2p, Dataset == "GDSC"), Pearson_Pval_GE < 0.05), Drug_Gene_Lin, Pearson_Pval_GE, Pearson_Corr_GE)
gdsc_cn <- select(subset(filter(dgl_signif_pearson_g2p, Dataset == "GDSC"), Pearson_Pval_CN < 0.05), Drug_Gene_Lin, Pearson_Pval_CN, Pearson_Corr_CN)
gdsc_Pear_n <- nrow(filter(dgl_signif_pearson_g2p, Dataset == "GDSC"))
gdsc_dgl_summ <- merge(gdsc_mut, merge(gdsc_ge, gdsc_cn, by = "Drug_Gene_Lin", all = TRUE), all = TRUE)
gdsc_dgl_summ <- with(gdsc_dgl_summ, cbind(colsplit(Drug_Gene_Lin, pattern = "_", names = c("Drug", "Gene", "Lineage")), 
                                           GDSC_p = p,
                                           GDSC_p_Bonf = ifelse(p < 0.05 / gdsc_mut_n, "Sig", NA),
                                           GDSC_r_GE = Pearson_Corr_GE,
                                           GDSC_p_GE = Pearson_Pval_GE,
                                           GDSC_p_GE_Bonf = ifelse(Pearson_Pval_GE < 0.05 / gdsc_Pear_n, "Sig", NA),
                                           GDSC_r_CN = Pearson_Corr_CN,
                                           GDSC_p_CN = Pearson_Pval_CN,
                                           GDSC_p_CN_Bonf = ifelse(Pearson_Pval_CN < 0.05 / gdsc_Pear_n, "Sig", NA)))

g2p_dgl_sig_summ <- merge(merge(ccle_dgl_summ, ctrp_dgl_summ, by = c("Drug", "Gene", "Lineage"), all = TRUE), gdsc_dgl_summ, by = c("Drug", "Gene", "Lineage"), all = TRUE)
```

## GE/CN: Grid

Test all combinations of G2P drugs and genes in all lineages. InG2P indicates G2P drug-gene associations test across all lineages.
```{r, eval=FALSE}
ccle_lineages4pearson <- as.character(unique(ccle_data_g2p_grid$group_general_lineage_name))[!is.na(as.character(unique(ccle_data_g2p_grid$group_general_lineage_name)))]
ccle_signif_ge <- lapply(ccle_lineages4pearson, PearsonByLineage, variable = "GE", dataset = ccle_data_g2p_grid)
ccle_signif_ge_all_lin <- rbindlist(ccle_signif_ge, use.names = TRUE)
ccle_signif_ge_all_lin$Dataset <- "CCLE"
ccle_signif_cn <- lapply(ccle_lineages4pearson, PearsonByLineage, variable = "CN", dataset = ccle_data_g2p_grid)
ccle_signif_cn_all_lin <- rbindlist(ccle_signif_cn, use.names = TRUE)
ccle_signif_cn_all_lin$Dataset <- "CCLE"
ccle_signif_all_lin <- merge(ccle_signif_ge_all_lin, ccle_signif_cn_all_lin, by = c("Drug_Gene", "group_general_lineage_name", "n", "Dataset"), all = TRUE)

ctrp_lineages4pearson <- as.character(unique(ctrp_data_g2p_grid$group_general_lineage_name))[!is.na(as.character(unique(ctrp_data_g2p_grid$group_general_lineage_name)))]
ctrp_signif_ge <- lapply(ctrp_lineages4pearson, PearsonByLineage, variable = "GE", dataset = ctrp_data_g2p_grid)
ctrp_signif_ge_all_lin <- rbindlist(ctrp_signif_ge, use.names = TRUE)
ctrp_signif_ge_all_lin$Dataset <- "CTRP"
ctrp_signif_cn <- lapply(ctrp_lineages4pearson, PearsonByLineage, variable = "CN", dataset = ctrp_data_g2p_grid)
ctrp_signif_cn_all_lin <- rbindlist(ctrp_signif_cn, use.names = TRUE)
ctrp_signif_cn_all_lin$Dataset <- "CTRP"
ctrp_signif_all_lin <- merge(ctrp_signif_ge_all_lin, ctrp_signif_cn_all_lin, by = c("Drug_Gene", "group_general_lineage_name", "n", "Dataset"), all = TRUE)

gdsc_lineages4pearson <- as.character(unique(gdsc_data_g2p_grid$group_general_lineage_name))[!is.na(as.character(unique(gdsc_data_g2p_grid$group_general_lineage_name)))]
gdsc_signif_ge <- lapply(gdsc_lineages4pearson, PearsonByLineage, variable = "GE", dataset = gdsc_data_g2p_grid)
gdsc_signif_ge_all_lin <- rbindlist(gdsc_signif_ge, use.names = TRUE)
gdsc_signif_ge_all_lin$Dataset <- "GDSC"
gdsc_signif_cn <- lapply(gdsc_lineages4pearson, PearsonByLineage, variable = "CN", dataset = gdsc_data_g2p_grid)
gdsc_signif_cn_all_lin <- rbindlist(gdsc_signif_cn, use.names = TRUE)
gdsc_signif_cn_all_lin$Dataset <- "GDSC"
gdsc_signif_all_lin <- merge(gdsc_signif_ge_all_lin, gdsc_signif_cn_all_lin, by = c("Drug_Gene", "group_general_lineage_name", "n", "Dataset"), all = TRUE)

signif_all_lin <- rbind(ccle_signif_all_lin, ctrp_signif_all_lin, gdsc_signif_all_lin)
signif_all_lin$n <- ifelse(is.na(signif_all_lin$n), 0, signif_all_lin$n)
signif_all_lin$InG2P <- ifelse(signif_all_lin$n != 0, "Yes", "No")
signif_all_lin$Drug_Gene_Lin <- paste0(signif_all_lin$Drug_Gene, "_", signif_all_lin$group_general_lineage_name)
saveRDS(signif_all_lin, "./data_munging/rds/dgl_signif_pearson_g2p_grid.rds")
```

```{r}
dgl_signif_pearson_g2p_grid <- readRDS("./data_munging/rds/dgl_signif_pearson_g2p_grid.rds")
```

```{r, results='hide', eval=FALSE}
dgl_pearson_g2p_grid_ttest_res <- lapply(as.character(unique(dgl_signif_pearson_g2p_grid$group_general_lineage_name)), PearsonPlotsByLineage)
dgl_pearson_g2p_grid_ttest_res_list <- rbindlist(dgl_pearson_g2p_grid_ttest_res, use.names = TRUE)
saveRDS(dgl_pearson_g2p_grid_ttest_res_list, "./data_munging/rds/dgl_pearson_g2p_grid_ttest_res_list.rds")
```

```{r}
dgl_pearson_g2p_grid_ttest_res_list <- readRDS("./data_munging/rds/dgl_pearson_g2p_grid_ttest_res_list.rds")

knitr::kable(dgl_pearson_g2p_grid_ttest_res_list[order(dgl_pearson_g2p_grid_ttest_res_list$p_value_GE), c("Dataset", "group_general_lineage_name", "t_statistic_GE", "df_GE", "p_value_GE")], caption = "Gene expression: T-test results for level A G2P drug-gene/gene expression associations across lineages") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")

knitr::kable(dgl_pearson_g2p_grid_ttest_res_list[order(dgl_pearson_g2p_grid_ttest_res_list$p_value_CN), c("Dataset", "group_general_lineage_name", "t_statistic_CN", "df_CN", "p_value_CN")], caption = "Copy number: T-test results for level A G2P drug-gene/gene expression associations across lineages") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

# MUT/GE/CN

--------------------------------------------------------------------------------

## Pancancer

This section contains summary plots for the Wilcoxon tests and Pearson/Spearman correlations done for only level A G2P drug-gene associations (i.e. not the grid tests) and G2P genes in the case of the CRISPR dataset.

### CRISPR

G2P genes with a significant effect on CERES score by mutation status, gene expression, and copy number. Summary plot using Pearson correlation coefficient for GE and CN:
```{r fig.height=5, fig.width=12}
crispr_res_mut <- select(crispr_signif_g2p_gene, Hugo_Symbol, p)
crispr_res_mut$corr <- 0
crispr_res_mut$Metric <- "Mutation Status"
crispr_res_mut$p_log10 <- -log10(crispr_res_mut$p)
subset(crispr_res_mut, p < 0.05)

subset(crispr_pearson_res_GE, p < 0.05)
subset(crispr_pearson_res_CN, p < 0.05)

crispr_pearson_res <- rbind(crispr_pearson_res_GE, crispr_pearson_res_CN)
crispr_pearson_res$Metric <- factor(crispr_pearson_res$Metric, levels = c("Gene Expression", "Copy Number"))
crispr_pearson_res$p_log10 <- -log10(crispr_pearson_res$p)

crispr_summ_res_mut_plot <- ggplot(data = subset(crispr_res_mut, p < 0.05)) +
  facet_grid(~ Metric) +
  scale_y_continuous(breaks = seq(0, 50, by = 5), labels = seq(0, 50, by = 5)) +
  coord_cartesian(ylim = c(0, 50)) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_label_repel(mapping = aes(y = p_log10, x = corr, label = Hugo_Symbol, color = Metric), force = 10, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  theme(legend.position = "none", panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(5.5, 1.5, 5.5, 5.5), "pt")) +
  scale_color_manual(values = c("Mutation Status" = "darkorchid4")) +
  labs(y = "-log10(p-value)")

crispr_summ_res_pearson_plot <- ggplot(data = subset(crispr_pearson_res, p < 0.05)) +
  facet_grid(~ Metric, scales = "free_x", space = "free") +
  scale_y_continuous(breaks = seq(0, 50, by = 5), labels = seq(0, 50, by = 5)) +
  scale_x_continuous(breaks = seq(-0.6, 0.3, by = 0.1), labels = formatC(seq(-0.6, 0.3, by = 0.1), digits = 2, format = "f")) +
  coord_cartesian(ylim = c(0, 50)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_label_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Hugo_Symbol), force = 10, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Pearson correlation coefficient (r)")

crispr_genomic_features_pearson_summ_plot <- plot_grid(crispr_summ_res_mut_plot, crispr_summ_res_pearson_plot, align = "h", rel_widths = c(1, 5))
crispr_genomic_features_pearson_summ_plot

# ggsave("./plots/manuscript/crispr_genomic_features_pearson_summ_plot.png", crispr_genomic_features_pearson_summ_plot, device = "png", dpi = 450, width = 12, height = 5, units = "in")
```

Summary plot using Spearman correlation coefficient for GE and CN:
```{r fig.height=5, fig.width=12}
subset(crispr_spearman_res_GE, p < 0.05)
subset(crispr_spearman_res_CN, p < 0.05)

crispr_spearman_res <- rbind(crispr_spearman_res_GE, crispr_spearman_res_CN)
crispr_spearman_res$Metric <- factor(crispr_spearman_res$Metric, levels = c("Gene Expression", "Copy Number"))
crispr_spearman_res$p_log10 <- -log10(crispr_spearman_res$p)

crispr_summ_res_spearman_plot <- ggplot(data = subset(crispr_spearman_res, p < 0.05)) +
  facet_grid(~ Metric, scales = "free_x", space = "free") +
  scale_y_continuous(breaks = seq(0, 50, by = 5), labels = seq(0, 50, by = 5)) +
  scale_x_continuous(breaks = seq(-0.6, 0.3, by = 0.1), labels = formatC(seq(-0.6, 0.3, by = 0.1), digits = 2, format = "f")) +
  coord_cartesian(ylim = c(0, 50)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_label_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Hugo_Symbol), force = 10, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Spearman's rank correlation coefficient (r_s)")

crispr_genomic_features_spearman_summ_plot <- plot_grid(crispr_summ_res_mut_plot, crispr_summ_res_spearman_plot, align = "h", rel_widths = c(1, 5))
crispr_genomic_features_spearman_summ_plot

# ggsave("./plots/manuscript/crispr_genomic_features_spearman_summ_plot.png", crispr_genomic_features_spearman_summ_plot, device = "png", dpi = 450, width = 12, height = 5, units = "in")
```

### CCLE

G2P drug-gene associations with a significant effect on AUC by mutation status, gene expression, and copy number. Summary plot using Pearson correlation coefficient for GE and CN:
```{r fig.height=4, fig.width=12}
ccle_res_mut <- select(ccle_signif_g2p, Drug_Gene, p)
ccle_res_mut$corr <- 0
ccle_res_mut$Metric <- "Mutation Status"
ccle_res_mut$p_log10 <- -log10(ccle_res_mut$p)
ccle_res_mut$Dataset <- "CCLE"

ccle_pearson_res_GE <- select(filter(ccle_cor_res, InG2P == "Yes"), Drug_Gene, Pearson_Corr_GE, Pearson_Pval_GE, Dataset)
colnames(ccle_pearson_res_GE) <- c("Drug_Gene", "corr", "p", "Dataset")
ccle_pearson_res_GE$Metric <- "Gene Expression"
subset(ccle_pearson_res_GE, p < 0.05)

ccle_pearson_res_CN <- select(filter(ccle_cor_res, InG2P == "Yes"), Drug_Gene, Pearson_Corr_CN, Pearson_Pval_CN, Dataset)
colnames(ccle_pearson_res_CN) <- c("Drug_Gene", "corr", "p", "Dataset")
ccle_pearson_res_CN$Metric <- "Copy Number"
subset(ccle_pearson_res_CN, p < 0.05)

ccle_pearson_res_filt <- rbind(ccle_pearson_res_GE, ccle_pearson_res_CN)
ccle_pearson_res_filt$Metric <- factor(ccle_pearson_res_filt$Metric, levels = c("Gene Expression", "Copy Number"))
ccle_pearson_res_filt$p_log10 <- -log10(ccle_pearson_res_filt$p)

ccle_summ_res_mut_plot <- ggplot(data = subset(ccle_res_mut, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  scale_y_continuous(breaks = seq(0, 35, by = 5), labels = seq(0, 35, by = 5)) +
  coord_cartesian(ylim = c(0, 35)) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_label_repel(mapping = aes(y = p_log10, x = corr, label = Drug_Gene, color = Metric), force = 10, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  theme(legend.position = "none", panel.grid.major.x = element_blank(), axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(5.5, 1.5, 5.5, 5.5), "pt"), strip.background.y = element_blank(), strip.text.y = element_blank()) +
  scale_color_manual(values = c("Mutation Status" = "darkorchid4")) +
  labs(y = "-log10(p-value)")

ccle_summ_res_pearson_plot <- ggplot(data = subset(ccle_pearson_res_filt, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  coord_cartesian(ylim = c(0, 35)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_label_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Drug_Gene), force = 10, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Pearson correlation coefficient (r)")

ccle_genomic_features_pearson_summ_plot <- plot_grid(ccle_summ_res_mut_plot, ccle_summ_res_pearson_plot, align = "h", rel_widths = c(1, 5))
ccle_genomic_features_pearson_summ_plot

# ggsave("./plots/manuscript/ccle_genomic_features_pearson_summ_plot.png", ccle_genomic_features_pearson_summ_plot, device = "png", dpi = 450, width = 12, height = 4, units = "in")
```

Summary plot using Spearman correlation coefficient for GE and CN:
```{r fig.height=4, fig.width=12}
ccle_spearman_res_GE <- select(filter(ccle_cor_res, InG2P == "Yes"), Drug_Gene, Spearman_Corr_GE, Spearman_Pval_GE, Dataset)
colnames(ccle_spearman_res_GE) <- c("Drug_Gene", "corr", "p", "Dataset")
ccle_spearman_res_GE$Metric <- "Gene Expression"
subset(ccle_spearman_res_GE, p < 0.05)

ccle_spearman_res_CN <- select(filter(ccle_cor_res, InG2P == "Yes"), Drug_Gene, Spearman_Corr_CN, Spearman_Pval_CN, Dataset)
colnames(ccle_spearman_res_CN) <- c("Drug_Gene", "corr", "p", "Dataset")
ccle_spearman_res_CN$Metric <- "Copy Number"
subset(ccle_spearman_res_CN, p < 0.05)

ccle_spearman_res_filt <- rbind(ccle_spearman_res_GE, ccle_spearman_res_CN)
ccle_spearman_res_filt$Metric <- factor(ccle_spearman_res_filt$Metric, levels = c("Gene Expression", "Copy Number"))
ccle_spearman_res_filt$p_log10 <- -log10(ccle_spearman_res_filt$p)

ccle_summ_res_spearman_plot <- ggplot(data = subset(ccle_spearman_res_filt, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  coord_cartesian(ylim = c(0, 35)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_label_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Drug_Gene), force = 10, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Spearman's rank correlation coefficient (r_s)")

ccle_genomic_features_spearman_summ_plot <- plot_grid(ccle_summ_res_mut_plot, ccle_summ_res_spearman_plot, align = "h", rel_widths = c(1, 5))
ccle_genomic_features_spearman_summ_plot

# ggsave("./plots/manuscript/ccle_genomic_features_spearman_summ_plot.png", ccle_genomic_features_spearman_summ_plot, device = "png", dpi = 450, width = 12, height = 4, units = "in")
```

### CTRP

G2P drug-gene associations with a significant effect on AUC by mutation status, gene expression, and copy number. Summary plot using Pearson correlation coefficient for GE and CN:
```{r fig.height=4, fig.width=12}
ctrp_res_mut <- select(ctrp_signif_g2p, Drug_Gene, p)
ctrp_res_mut$corr <- 0
ctrp_res_mut$Metric <- "Mutation Status"
ctrp_res_mut$p_log10 <- -log10(ctrp_res_mut$p)
ctrp_res_mut$Dataset <- "CTRP"

ctrp_pearson_res_GE <- select(filter(ctrp_cor_res, InG2P == "Yes"), Drug_Gene, Pearson_Corr_GE, Pearson_Pval_GE, Dataset)
colnames(ctrp_pearson_res_GE) <- c("Drug_Gene", "corr", "p", "Dataset")
ctrp_pearson_res_GE$Metric <- "Gene Expression"
subset(ctrp_pearson_res_GE, p < 0.05)

ctrp_pearson_res_CN <- select(filter(ctrp_cor_res, InG2P == "Yes"), Drug_Gene, Pearson_Corr_CN, Pearson_Pval_CN, Dataset)
colnames(ctrp_pearson_res_CN) <- c("Drug_Gene", "corr", "p", "Dataset")
ctrp_pearson_res_CN$Metric <- "Copy Number"
subset(ctrp_pearson_res_CN, p < 0.05)

ctrp_pearson_res_filt <- rbind(ctrp_pearson_res_GE, ctrp_pearson_res_CN)
ctrp_pearson_res_filt$Metric <- factor(ctrp_pearson_res_filt$Metric, levels = c("Gene Expression", "Copy Number"))
ctrp_pearson_res_filt$p_log10 <- -log10(ctrp_pearson_res_filt$p)

ctrp_summ_res_mut_plot <- ggplot(data = subset(ctrp_res_mut, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  scale_y_continuous(breaks = seq(0, 35, by = 5), labels = seq(0, 35, by = 5)) +
  coord_cartesian(ylim = c(0, 35)) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_text_repel(mapping = aes(y = p_log10, x = corr, label = Drug_Gene, color = Metric), force = 15, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.9) +
  theme(legend.position = "none", panel.grid.major.x = element_blank(), axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(5.5, 1.5, 5.5, 5.5), "pt"), strip.background.y = element_blank(), strip.text.y = element_blank()) +
  scale_color_manual(values = c("Mutation Status" = "darkorchid4")) +
  labs(y = "-log10(p-value)")

ctrp_summ_res_pearson_plot <- ggplot(data = subset(ctrp_pearson_res_filt, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  coord_cartesian(ylim = c(0, 35)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_text_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Drug_Gene), force = 10, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Pearson correlation coefficient (r)")

ctrp_genomic_features_pearson_summ_plot <- plot_grid(ctrp_summ_res_mut_plot, ctrp_summ_res_pearson_plot, align = "h", rel_widths = c(1, 5))
ctrp_genomic_features_pearson_summ_plot

# ggsave("./plots/manuscript/ctrp_genomic_features_pearson_summ_plot.png", ctrp_genomic_features_pearson_summ_plot, device = "png", dpi = 450, width = 12, height = 4, units = "in")
```

Summary plot using Spearman correlation coefficient for GE and CN:
```{r fig.height=4, fig.width=12}
ctrp_spearman_res_GE <- select(filter(ctrp_cor_res, InG2P == "Yes"), Drug_Gene, Spearman_Corr_GE, Spearman_Pval_GE, Dataset)
colnames(ctrp_spearman_res_GE) <- c("Drug_Gene", "corr", "p", "Dataset")
ctrp_spearman_res_GE$Metric <- "Gene Expression"
subset(ctrp_spearman_res_GE, p < 0.05)

ctrp_spearman_res_CN <- select(filter(ctrp_cor_res, InG2P == "Yes"), Drug_Gene, Spearman_Corr_CN, Spearman_Pval_CN, Dataset)
colnames(ctrp_spearman_res_CN) <- c("Drug_Gene", "corr", "p", "Dataset")
ctrp_spearman_res_CN$Metric <- "Copy Number"
subset(ctrp_spearman_res_CN, p < 0.05)

ctrp_spearman_res_filt <- rbind(ctrp_spearman_res_GE, ctrp_spearman_res_CN)
ctrp_spearman_res_filt$Metric <- factor(ctrp_spearman_res_filt$Metric, levels = c("Gene Expression", "Copy Number"))
ctrp_spearman_res_filt$p_log10 <- -log10(ctrp_spearman_res_filt$p)

ctrp_summ_res_spearman_plot <- ggplot(data = subset(ctrp_spearman_res_filt, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  coord_cartesian(ylim = c(0, 35)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_text_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Drug_Gene), force = 20, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Spearman's rank correlation coefficient (r_s)")

ctrp_genomic_features_spearman_summ_plot <- plot_grid(ctrp_summ_res_mut_plot, ctrp_summ_res_spearman_plot, align = "h", rel_widths = c(1, 5))
ctrp_genomic_features_spearman_summ_plot

# ggsave("./plots/manuscript/ctrp_genomic_features_spearman_summ_plot.png", ctrp_genomic_features_spearman_summ_plot, device = "png", dpi = 450, width = 12, height = 4, units = "in")
```

### GDSC

G2P drug-gene associations with a significant effect on AUC by mutation status, gene expression, and copy number. Summary plot using Pearson correlation coefficient for GE and CN:
```{r fig.height=4, fig.width=12}
gdsc_res_mut <- select(gdsc_signif_g2p, Drug_Gene, p)
gdsc_res_mut$corr <- 0
gdsc_res_mut$Metric <- "Mutation Status"
gdsc_res_mut$p_log10 <- -log10(gdsc_res_mut$p)
gdsc_res_mut$Dataset <- "GDSC"

gdsc_pearson_res_GE <- select(filter(gdsc_cor_res, InG2P == "Yes"), Drug_Gene, Pearson_Corr_GE, Pearson_Pval_GE, Dataset)
colnames(gdsc_pearson_res_GE) <- c("Drug_Gene", "corr", "p", "Dataset")
gdsc_pearson_res_GE$Metric <- "Gene Expression"
subset(gdsc_pearson_res_GE, p < 0.05)

gdsc_pearson_res_CN <- select(filter(gdsc_cor_res, InG2P == "Yes"), Drug_Gene, Pearson_Corr_CN, Pearson_Pval_CN, Dataset)
colnames(gdsc_pearson_res_CN) <- c("Drug_Gene", "corr", "p", "Dataset")
gdsc_pearson_res_CN$Metric <- "Copy Number"
subset(gdsc_pearson_res_CN, p < 0.05)

gdsc_pearson_res_filt <- rbind(gdsc_pearson_res_GE, gdsc_pearson_res_CN)
gdsc_pearson_res_filt$Metric <- factor(gdsc_pearson_res_filt$Metric, levels = c("Gene Expression", "Copy Number"))
gdsc_pearson_res_filt$p_log10 <- -log10(gdsc_pearson_res_filt$p)

gdsc_summ_res_mut_plot <- ggplot(data = subset(gdsc_res_mut, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  scale_y_continuous(breaks = seq(0, 16, by = 1), labels = seq(0, 16, by = 1)) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_text_repel(mapping = aes(y = p_log10, x = corr, label = Drug_Gene, color = Metric), force = 15, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.9) +
  theme(legend.position = "none", panel.grid.major.x = element_blank(), axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(5.5, 1.5, 5.5, 5.5), "pt"), strip.background.y = element_blank(), strip.text.y = element_blank()) +
  scale_color_manual(values = c("Mutation Status" = "darkorchid4")) +
  labs(y = "-log10(p-value)")

gdsc_summ_res_pearson_plot <- ggplot(data = subset(gdsc_pearson_res_filt, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  scale_y_continuous(breaks = seq(0, 100, by = 10), labels = seq(0, 100, by = 10)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_text_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Drug_Gene), force = 10, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Pearson correlation coefficient (r)")

gdsc_genomic_features_pearson_summ_plot <- plot_grid(gdsc_summ_res_mut_plot, gdsc_summ_res_pearson_plot, align = "h", rel_widths = c(1, 5))
gdsc_genomic_features_pearson_summ_plot

# ggsave("./plots/manuscript/gdsc_genomic_features_pearson_summ_plot.png", gdsc_genomic_features_pearson_summ_plot, device = "png", dpi = 450, width = 12, height = 4, units = "in")
```

Summary plot using Spearman correlation coefficient for GE and CN:
```{r fig.height=4, fig.width=12}
gdsc_spearman_res_GE <- select(filter(gdsc_cor_res, InG2P == "Yes"), Drug_Gene, Spearman_Corr_GE, Spearman_Pval_GE, Dataset)
colnames(gdsc_spearman_res_GE) <- c("Drug_Gene", "corr", "p", "Dataset")
gdsc_spearman_res_GE$Metric <- "Gene Expression"
subset(gdsc_spearman_res_GE, p < 0.05)

gdsc_spearman_res_CN <- select(filter(gdsc_cor_res, InG2P == "Yes"), Drug_Gene, Spearman_Corr_CN, Spearman_Pval_CN, Dataset)
colnames(gdsc_spearman_res_CN) <- c("Drug_Gene", "corr", "p", "Dataset")
gdsc_spearman_res_CN$Metric <- "Copy Number"
subset(gdsc_spearman_res_CN, p < 0.05)

gdsc_spearman_res_filt <- rbind(gdsc_spearman_res_GE, gdsc_spearman_res_CN)
gdsc_spearman_res_filt$Metric <- factor(gdsc_spearman_res_filt$Metric, levels = c("Gene Expression", "Copy Number"))
gdsc_spearman_res_filt$p_log10 <- -log10(gdsc_spearman_res_filt$p)

gdsc_summ_res_spearman_plot <- ggplot(data = subset(gdsc_spearman_res_filt, p < 0.05)) +
  facet_grid(Dataset ~ Metric) +
  scale_y_continuous(breaks = seq(0, 50, by = 5), labels = seq(0, 50, by = 5)) +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Metric), alpha = 0.5) +
  geom_text_repel(mapping = aes(y = p_log10, x = corr, color = Metric, label = Drug_Gene), force = 20, nudge_y = 2, nudge_x = 0.05, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("Gene Expression" = "darkorchid1", "Copy Number" = "mediumorchid1")) +
  theme(legend.position = "none", axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Spearman's rank correlation coefficient (r_s)")

gdsc_genomic_features_spearman_summ_plot <- plot_grid(gdsc_summ_res_mut_plot, gdsc_summ_res_spearman_plot, align = "h", rel_widths = c(1, 5))
gdsc_genomic_features_spearman_summ_plot

# ggsave("./plots/manuscript/gdsc_genomic_features_spearman_summ_plot.png", gdsc_genomic_features_spearman_summ_plot, device = "png", dpi = 450, width = 12, height = 4, units = "in")
```

### Drug: Mut

G2P drug-gene associations with a significant effect on AUC by mutation status:
```{r fig.height=4, fig.width=8}
all_mut_res <- rbind(ccle_res_mut, ctrp_res_mut, gdsc_res_mut)

all_mut_summ_plot <- ggplot(data = subset(all_mut_res, p < 0.05)) +
  facet_grid(~ Dataset) +
  geom_point(aes(x = corr, y = p_log10, color = Dataset), alpha = 0.5) +
  scale_y_continuous(breaks = seq(0, 12, by = 2), labels = seq(0, 12, by = 2)) +
  scale_x_continuous(breaks = seq(0, 0.1, by = 0.1), labels = seq(0, 0.1, by = 0.1)) +
  coord_cartesian(xlim = c(0, 0.05)) +
  geom_text_repel(mapping = aes(y = p_log10, x = corr, color = Dataset, label = Drug_Gene), nudge_x = 0.05, direction = "y", hjust = 0, size = 2.5, segment.size = 0.1, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  theme(legend.position = "none", plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt"), panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), axis.text.x = element_blank()) +
  labs(y = "-log10(p-value)")
all_mut_summ_plot

# ggsave("./plots/manuscript/all_mut_summ_plot.png", all_mut_summ_plot, device = "png", dpi = 450, width = 12, height = 6, units = "in")
```

### Drug: GE and CN

G2P drug-gene associations with a significant effect on AUC by gene expression and copy number:
```{r fig.height=12, fig.width=12}
all_pearson_res <- rbind(ccle_pearson_res_filt, ctrp_pearson_res_filt, gdsc_pearson_res_filt)
all_pearson_summ_plot <- ggplot(data = subset(all_pearson_res, p < 0.05)) +
  facet_grid(Dataset ~ Metric, scales = "free", space = "free") +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Dataset), alpha = 0.5) +
  scale_y_continuous(breaks = seq(0, 60, by = 10), labels = seq(0, 60, by = 10)) +
  geom_text_repel(mapping = aes(y = p_log10, x = corr, color = Dataset, label = Drug_Gene), force = 10, nudge_y = 3, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  theme(legend.position = "none", plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Pearson correlation coefficient (r)", y = "-log10(p-value)")
all_pearson_summ_plot
# ggsave("./plots/manuscript/all_pearson_summ_plot.png", all_pearson_summ_plot, device = "png", dpi = 450, width = 12, height = 10, units = "in")

all_spearman_res <- rbind(ccle_spearman_res_filt, ctrp_spearman_res_filt, gdsc_spearman_res_filt)
all_spearman_summ_plot <- ggplot(data = subset(all_spearman_res, p < 0.05)) +
  facet_grid(Dataset ~ Metric, scales = "free", space = "free") +
  geom_vline(xintercept = 0, color = "grey50", lty = 2, lwd = 0.3) +
  geom_point(aes(x = corr, y = p_log10, color = Dataset), alpha = 0.5) +
  scale_y_continuous(breaks = seq(0, 60, by = 10), labels = seq(0, 60, by = 10)) +
  geom_text_repel(mapping = aes(y = p_log10, x = corr, color = Dataset, label = Drug_Gene), force = 10, nudge_y = 1, size = 2.5, segment.size = 0.2, segment.colour = "grey50", box.padding = 0.6) +
  scale_color_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  theme(legend.position = "none", plot.margin = unit(c(5.5, 5.5, 5.5, 1.5), "pt")) +
  labs(x = "Pearson correlation coefficient (r)", y = "-log10(p-value)")
all_spearman_summ_plot
# ggsave("./plots/manuscript/all_spearman_summ_plot.png", all_spearman_summ_plot, device = "png", dpi = 450, width = 12, height = 10, units = "in")
```

### Drug: P-values

```{r fig.height=6, fig.width=12}
pval_summ_mut <- g2p_mut_all_summ
pval_summ_mut$Measure <- "Mutation Status"
pval_summ_mut$Drug_Gene_Ordered <- NULL

pval_summ_ge <- select(g2p_cor, Drug_Gene, n, Pearson_Pval_GE, InG2P, Dataset)
colnames(pval_summ_ge)[colnames(pval_summ_ge) == "Pearson_Pval_GE"] <- "p"
pval_summ_ge$p_log10 <- -log10(pval_summ_ge$p)
pval_summ_ge$Measure <- "Gene Expression"

pval_summ_cn <- select(g2p_cor, Drug_Gene, n, Pearson_Pval_CN, InG2P, Dataset)
colnames(pval_summ_cn)[colnames(pval_summ_cn) == "Pearson_Pval_CN"] <- "p"
pval_summ_cn$p_log10 <- -log10(pval_summ_cn$p)
pval_summ_cn$Measure <- "Copy Number"

pval_summ <- rbind(pval_summ_mut, pval_summ_ge, pval_summ_cn)
pval_summ$Measure <- factor(pval_summ$Measure, levels = c("Mutation Status", "Gene Expression", "Copy Number"))

pval_summ_plot <- ggplot(data = pval_summ, aes(x = p_log10, y = n, color = Dataset, shape = Dataset, label = Drug_Gene)) +
  facet_wrap(~ Measure, scales = "free_x") +
  geom_point(alpha = 0.7) +
  scale_color_manual(values = c("green3", "darkslategray4", "darkturquoise")) +
  geom_text_repel(data = filter(pval_summ, n > 20 & Measure == "Mutation Status"), mapping = aes(x = p_log10, y = n, color = Dataset, label = Drug_Gene), direction = "y", nudge_x = 9, hjust = 0, size = 2.5, segment.size = 0.1, segment.colour = "grey50", box.padding = 0.6, show.legend = FALSE) +
  geom_text_repel(data = filter(pval_summ, n > 20 & Measure == "Gene Expression"), mapping = aes(x = p_log10, y = n, color = Dataset, label = Drug_Gene), direction = "y", nudge_x = 50, hjust = 0, size = 2.5, segment.size = 0.1, segment.colour = "grey50", box.padding = 0.6, show.legend = FALSE) +
  geom_text_repel(data = filter(pval_summ, n > 20 & Measure == "Copy Number"), mapping = aes(x = p_log10, y = n, color = Dataset, label = Drug_Gene), direction = "y", nudge_x = 20, hjust = 0, size = 2.5, segment.size = 0.1, segment.colour = "grey50", box.padding = 0.6, show.legend = FALSE) +
  scale_y_continuous(breaks = seq(0, 120, by = 10), labels = seq(0, 120, by = 10)) +
  labs(x = "-log10(p-value)", y = "Number of level A G2P associations") +
  theme(legend.position = "bottom")
pval_summ_plot

# ggsave("./plots/manuscript/all_pval_summ_plot.png", pval_summ_plot, device = "png", dpi = 450, width = 12, height = 6, units = "in")

pval_summ_plot2 <- ggplot(data = pval_summ, aes(x = p_log10, y = n, color = Dataset, shape = Dataset, label = Drug_Gene)) +
  facet_wrap(~ Measure, scales = "free_x") +
  geom_point(alpha = 0.7) +
  scale_color_manual(values = c("green3", "darkslategray4", "darkturquoise")) +
  geom_text_repel(data = filter(pval_summ, p_log10 > 5 & Measure == "Mutation Status"), mapping = aes(x = p_log10, y = n, color = Dataset, label = Drug_Gene), direction = "y", nudge_y = 10, hjust = 0, size = 2.5, segment.size = 0.1, segment.colour = "grey50", box.padding = 0.6, show.legend = FALSE) +
  geom_text_repel(data = filter(pval_summ, p_log10 > 20 & Measure == "Gene Expression"), mapping = aes(x = p_log10, y = n, color = Dataset, label = Drug_Gene), direction = "y", nudge_y = 10, hjust = 0, size = 2.5, segment.size = 0.1, segment.colour = "grey50", box.padding = 0.6, show.legend = FALSE) +
  geom_text_repel(data = filter(pval_summ, p_log10 > 7.5 & Measure == "Copy Number"), mapping = aes(x = p_log10, y = n, color = Dataset, label = Drug_Gene), direction = "y", nudge_y = 10, hjust = 0, size = 2.5, segment.size = 0.1, segment.colour = "grey50", box.padding = 0.6, show.legend = FALSE) +
  scale_y_continuous(breaks = seq(0, 120, by = 10), labels = seq(0, 120, by = 10)) +
  labs(x = "-log10(p-value)", y = "Number of level A G2P associations") +
  theme(legend.position = "bottom")
pval_summ_plot2

# ggsave("./plots/manuscript/all_pval_summ_plot2.png", pval_summ_plot2, device = "png", dpi = 450, width = 12, height = 6, units = "in")
```

### Summary stats

```{r}
# Mutation status
g2p_mut_all_summ_stats <- g2p_mut_all_summ
g2p_mut_all_summ_stats$n_true <- TRUE
g2p_mut_all_summ_stats <- g2p_mut_all_summ_stats %>% drop_na(p) %>% group_by(Dataset, InG2P) %>% summarize(p_mean = round(mean(p), 2), p_sd = round(sd(p), 2), n = sum(n_true), n_signif = sum(p < 0.05), n_sig_perc = n_signif / n * 100)

# Pearson correlations
g2p_cor_stats <- g2p_cor
g2p_cor_stats$n_true <- TRUE
g2p_cor_stats <- g2p_cor_stats %>% group_by(Dataset, InG2P) %>% drop_na() %>% summarize(
          ge_mean = round(mean(Pearson_GE_Abs), 5),
          ge_sd = round(sd(Pearson_GE_Abs), 5),
          ge_n = sum(n_true),
          ge_n_signif = sum(Pearson_Pval_GE < 0.05),
          ge_n_sig_perc = ge_n_signif / ge_n * 100,
          cn_mean = round(mean(Pearson_CN_Abs), 3),
          cn_sd = round(sd(Pearson_CN_Abs), 5),
          cn_n = sum(n_true),
          cn_n_signif = sum(Pearson_Pval_CN < 0.05),
          cn_n_sig_perc = cn_n_signif / cn_n * 100)
```

### CTRP example derivation

Example p-value and Pearson correlation coefficient derivations for afatinib and ERBB2 in the CTRP dataset:
```{r}
ctrp_grid_summary_data <- filter(ctrp_data_g2p_grid, Drug_Gene == "afatinib_ERBB2")

ctrp_grid_summary_mut_corr <- wilcox.test(AUC_zscore ~ Mutation_Status_Nonsilent, data = ctrp_grid_summary_data[!is.na(ctrp_grid_summary_data$Mutation_Status_Nonsilent),], paired = FALSE)
ctrp_grid_summary_mut_corr_text <- paste0("W = ", formatC(ctrp_grid_summary_mut_corr$statistic, format = "d", big.mark = ","), "\np = ", round(ctrp_grid_summary_mut_corr$p.value, 3))

ctrp_grid_summary_ge_corr <- cor.test(y = ctrp_grid_summary_data$AUC, x = ctrp_grid_summary_data$RPKM, method = "pearson", use = "complete.obs")
ctrp_grid_summary_ge_corr_text <- paste0("r(", ctrp_grid_summary_ge_corr$parameter, ") = ", round(ctrp_grid_summary_ge_corr$estimate, 2), "\np ", ifelse(round(ctrp_grid_summary_ge_corr$p.value, 3) == 0, "< 0.001", paste0("= ", round(ctrp_grid_summary_ge_corr$p.value, 3))))

ctrp_grid_summary_cn_corr <- cor.test(y = ctrp_grid_summary_data$AUC, x = ctrp_grid_summary_data$CN, method = "pearson", use = "complete.obs")
ctrp_grid_summary_cn_corr_text <- paste0("r(", ctrp_grid_summary_cn_corr$parameter, ") = ", formatC(ctrp_grid_summary_cn_corr$estimate, digits = 2, format = "f"), "\np ", ifelse(round(ctrp_grid_summary_cn_corr$p.value, 3) == 0, "< 0.001", paste0("= ", round(ctrp_grid_summary_cn_corr$p.value, 3))))

ctrp_grid_summary_mut_plot <- ggplot(data = ctrp_grid_summary_data[!is.na(ctrp_grid_summary_data$Mutation_Status_Nonsilent),], aes(y = AUC_zscore, x = Mutation_Status_Nonsilent, fill = Mutation_Status_Nonsilent)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
  geom_jitter(width = 0.2, shape = 21) +
  scale_fill_manual(values = c("Wildtype" = "lightsteelblue1", "Mutant" = "steelblue4")) +
  annotate(geom = "text", label = ctrp_grid_summary_mut_corr_text, x = 1.5, y = 1.3) +
  guides(fill = FALSE) +
  labs(x = "ERBB2 mutation status", y = "Afatinib AUC (z-score)") +
  geom_segment(aes(x = 1, y = 1.5, xend = 1.1, yend = 1.5), lwd = 0.2) +
  geom_segment(aes(x = 1.9, y = 1.5, xend = 2, yend = 1.5), lwd = 0.2) +
  geom_segment(aes(x = 1, y = 1.5, xend = 1, yend = 1.35), lwd = 0.2) +
  geom_segment(aes(x = 2, y = 1.5, xend = 2, yend = 1.35), lwd = 0.2)
ctrp_grid_summary_mut_plot

ctrp_grid_summary_ge_plot <- ggplot(data = ctrp_grid_summary_data, aes(y = AUC_zscore, x = RPKM)) +
  geom_point(alpha = 0.25, size = 1.25, color = "gray4") +
  geom_smooth(method = "lm", se = FALSE, lwd = 0.5, lty = 2) +
  annotate(geom = "text", label = ctrp_grid_summary_ge_corr_text, x = 2000, y = 1.3) +
  labs(x = "ERBB2 gene expression (RPKM)", y = "Afatinib AUC (z-score)") +
  theme(axis.title.y = element_blank()) +
  geom_curve(aes(x = 750, y = -1.9, xend = 1720, yend = 1.1), lwd = 0.2, curvature = -0.2, arrow = arrow(length = unit(0.03, "npc")))
ctrp_grid_summary_ge_plot

ctrp_grid_summary_cn_plot <- ggplot(data = ctrp_grid_summary_data, aes(y = AUC_zscore, x = CN)) +
  geom_point(alpha = 0.25, size = 1.25, color = "gray4") +
  geom_smooth(method = "lm", se = FALSE, lwd = 0.5, lty = 2) +
  annotate(geom = "text", label = ctrp_grid_summary_cn_corr_text, x = 5, y = 1.3) +
  labs(x = "ERBB2 copy number (log2 ratio)") +
  theme(axis.title.y = element_blank()) +
  geom_curve(aes(x = 4, y = -2.3, xend = 5, yend = 0.8), lwd = 0.2, curvature = 0.2, arrow = arrow(length = unit(0.03, "npc")))
ctrp_grid_summary_cn_plot

# ggsave("./plots/manuscript/ctrp_grid_summary_mut_plot.png", ctrp_grid_summary_mut_plot, device = "png", dpi = 450, width = 3, height = 4, units = "in")
# ggsave("./plots/manuscript/ctrp_grid_summary_ge_plot.png", ctrp_grid_summary_ge_plot, device = "png", dpi = 450, width = 5, height = 4, units = "in")
# ggsave("./plots/manuscript/ctrp_grid_summary_cn_plot.png", ctrp_grid_summary_cn_plot, device = "png", dpi = 450, width = 5, height = 4, units = "in")
```

## By lineage

### Kolmogorov-Smirnov tests

```{r}
MakeCorrResSummaryTable <- function(df_list, dataset_name) {
  res_full <- data.frame(matrix(ncol = 8, nrow = 0))
  for(i in 1:length(df_list)) {
    corr_res <- df_list[[i]] %>% group_by(G2P_DGL) %>% summarize(Dataset = dataset_name,
                Drug_Gene = names(df_list)[i],
                Lineages = paste(as.character(unique(group_general_lineage_name)), collapse = ", "),
                N_Lineages = length(as.character(unique(group_general_lineage_name))),
                Pearson_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs")$estimate,
                Pearson_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs")$estimate,
                Spearman_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs")$estimate,
                Spearman_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs")$estimate)
    corr_res <- merge(corr_res, compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, data = df_list[[i]], group.by = "G2P_DGL", method = "wilcox.test")[, c("G2P_DGL", "p")], by = "G2P_DGL", all = TRUE)
    corr_res$Mut_Signif <- ifelse(corr_res$p < 0.05, 1, 0)
    res_full <- rbind(res_full, corr_res)
  }
  return(res_full)
}
```

```{r}
ccle_data_g2p$G2P_DGL <- ifelse(ccle_data_g2p$Drug_Gene_Lin %in% g2p_dgls, "Yes", "No")
ccle_data_g2p_split <- split(ccle_data_g2p, f = ccle_data_g2p$Drug_Gene)
ccle_corr_all <- MakeCorrResSummaryTable(ccle_data_g2p_split, dataset_name = "CCLE")

ctrp_data_g2p$G2P_DGL <- ifelse(ctrp_data_g2p$Drug_Gene_Lin %in% g2p_dgls, "Yes", "No")
ctrp_data_g2p_split <- split(ctrp_data_g2p, f = ctrp_data_g2p$Drug_Gene)
ctrp_corr_all <- MakeCorrResSummaryTable(ctrp_data_g2p_split, dataset_name = "CTRP")

gdsc_data_g2p$G2P_DGL <- ifelse(gdsc_data_g2p$Drug_Gene_Lin %in% g2p_dgls, "Yes", "No")
gdsc_data_g2p_split <- split(gdsc_data_g2p, f = gdsc_data_g2p$Drug_Gene)
gdsc_corr_all <- MakeCorrResSummaryTable(gdsc_data_g2p_split, dataset_name = "GDSC")

g2p_corr_all <- rbind(ccle_corr_all, ctrp_corr_all, gdsc_corr_all)
```

```{r}
G2PCorr_KStest <- function(dataset, dataset_name) {
  Metric <- colnames(dataset)[6:9]
  D_stat <- NULL
  pval <- NULL
  signif <- NULL

  yes <- dataset[dataset$G2P_DGL == "Yes",]
  no <- dataset[dataset$G2P_DGL == "No",]

  for(i in 6:9) {
    # print(yes[[colnames(yes)[i]]])
    # print(no[[colnames(no)[i]]])
    test <- ks.test(yes[[colnames(yes)[i]]], no[[colnames(no)[i]]])
    D_stat <- c(D_stat, test$statistic)
    pval <- c(pval, test$p.value)
    signif <- c(signif, ifelse(test$p.value < 0.001, "***", ifelse(test$p.value < 0.01, "**", ifelse(test$p.value < 0.05, "*", NA))))
  }
  df <- data.frame(Metric, D_stat, pval, signif)
  df$Dataset <- dataset_name
  return(df)
}
```

```{r}
ccle_corr_stats <- G2PCorr_KStest(ccle_corr_all, "CCLE")
ctrp_corr_stats <- G2PCorr_KStest(ctrp_corr_all, "CTRP")
gdsc_corr_stats <- G2PCorr_KStest(gdsc_corr_all, "GDSC")
g2p_corr_stats <- rbind(ccle_corr_stats, ctrp_corr_stats, gdsc_corr_stats)
g2p_corr_stats

g2p_mut_stats <- g2p_corr_all %>% group_by(Dataset, G2P_DGL) %>% drop_na(Mut_Signif) %>% summarize(N_Signif_Muts = sum(Mut_Signif),
                                                                                                   Total_Tests = length(Mut_Signif),
                                                                                                   Perc_Signif = N_Signif_Muts / Total_Tests * 100)
g2p_mut_stats
```

### Grid t-test summaries

Figure #. Histograms of p-values from t-tests comparing distributions of summary statistics for CCLE, CTRP, and GDSC drug screen results in the context of G2P level A evidence and lineage. Each p-value is derived from comparisons similar to those shown in Figure 2, but for these p-values, the data is grouped by lineage prior to conducting the t-test rather than the test being lineage-agnostic. To produce these p-values, AUC was correlated with each genomic feature for all possible pairwise combinations of G2P drugs and genes in all lineages represented in each dataset. For each lineage, t-tests were used to compare distributions of p-values (mutation status) or Pearson correlation coefficients (gene expression and copy number) that were grouped by whether or not the underlying drug-gene combination had level-A G2P evidence. For each dataset, if there were two or more correlation results in each G2P presence/absence group in a specific lineage, a t-test was conducted and the p-value appears in the appropriate histogram. Mutation status histograms are sparse due to limited mutation annotations for many of the G2P genes in all screened cell lines.
```{r fig.height=4, fig.width=12}
ttest_summ_res_mut <- dgl_mut_g2p_grid_ttest_res_list[, c("p_value", "Dataset", "group_general_lineage_name")]
ttest_summ_res_mut$Metric <- "Mutation Status"

ttest_summ_res_ge <- dgl_pearson_g2p_grid_ttest_res_list[, c("p_value_GE", "Dataset", "group_general_lineage_name")]
colnames(ttest_summ_res_ge) <- c("p_value", "Dataset", "group_general_lineage_name")
ttest_summ_res_ge$Metric <- "Gene Expression"

ttest_summ_res_cn <- dgl_pearson_g2p_grid_ttest_res_list[, c("p_value_CN", "Dataset", "group_general_lineage_name")]
colnames(ttest_summ_res_cn) <- c("p_value", "Dataset", "group_general_lineage_name")
ttest_summ_res_cn$Metric <- "Copy Number"

ttest_summ_res <- rbind(ttest_summ_res_mut, ttest_summ_res_ge, ttest_summ_res_cn)
ttest_summ_res$Metric <- factor(ttest_summ_res$Metric, levels = c("Mutation Status", "Gene Expression", "Copy Number"))

ttest_mut_summ_facet <- ggplot(data = ttest_summ_res, aes(x = p_value, fill = Dataset, color = Dataset)) +
  facet_grid(Metric ~ Dataset) +
  geom_histogram(bins = 40, binwidth = 0.025, boundary = 0, closed = "left") +
  scale_y_continuous(breaks = c(0:6), labels = c(0:6)) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.1), expand = c(0.01, 0.01)) +
  scale_fill_manual(values = c("CCLE" = "palegreen3", "CTRP" = "slategray3", "GDSC" = "paleturquoise3")) +
  scale_color_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  theme(legend.position = c(0.5, 1.7), legend.direction = "horizontal", panel.grid.minor.y = element_blank()) +
  labs(x = "P-value", y = "Frquency")
ttest_mut_summ_facet

# ggsave("./plots/manuscript/all_dgl_ttest_summ_facet_plot.png", ttest_mut_summ_facet, device = "png", dpi = 450, width = 12, height = 4, units = "in")
```

Big grid plot by lineages:
```{r fig.height=20, fig.width=12}
dgl_mut <- dgl_mut_grid_summ
dgl_mut$Sign <- "None"
dgl_mut$Measure <- "Mutation Status"
colnames(dgl_mut)[colnames(dgl_mut) == "p"] <- "Statistic"
dgl_mut$Drug_Gene_Lin <- NULL

dgl_ge <- select(dgl_signif_pearson_g2p_grid, Drug_Gene, group_general_lineage_name, n, Pearson_GE_Abs, InG2P, Dataset, Pearson_GE_Sign)
dgl_ge$Measure <- "Gene Expression"
colnames(dgl_ge)[colnames(dgl_ge) == "Pearson_GE_Abs"] <- "Statistic"
colnames(dgl_ge)[colnames(dgl_ge) == "Pearson_GE_Sign"] <- "Sign"

dgl_cn <- select(dgl_signif_pearson_g2p_grid, Drug_Gene, group_general_lineage_name, n, Pearson_CN_Abs, InG2P, Dataset, Pearson_CN_Sign)
dgl_cn$Measure <- "Copy Number"
colnames(dgl_cn)[colnames(dgl_cn) == "Pearson_CN_Abs"] <- "Statistic"
colnames(dgl_cn)[colnames(dgl_cn) == "Pearson_CN_Sign"] <- "Sign"

dgl_all <- rbind(dgl_mut, dgl_ge, dgl_cn)
dgl_all$Sign <- ifelse(is.na(dgl_all$Sign), "None", dgl_all$Sign)

dgl_all_plot <- ggplot(data = dgl_all, aes(x = InG2P, y = Statistic)) +
  facet_grid(group_general_lineage_name ~ Measure * Dataset) +
  geom_jitter(mapping = aes(shape = Sign, fill = InG2P, color = InG2P), width = 0.4, alpha = 0.5, size = 1) +
  geom_boxplot(width = 0.25, fill = NA, outlier.shape = NA) +
  scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
  scale_shape_manual(values = c(25, 24, 4), labels = c("Negative", "Positive", "NA")) +
  theme(legend.position = "top") +
  guides(color = FALSE, fill = FALSE)

# ggsave("./plots/manuscript/all_dgl_plot.png", dgl_all_plot, device = "png", dpi = 450, width = 12, height = 40, units = "in")
```

## Cell line overlap

```{r, eval=FALSE}
crispr_ccl <- unique(select(crispr_data, DepMap_ID, CCLE_Name))
ccle_ccl <- unique(select(ccle_data, DepMap_ID, CCLE_Name, accession_id))
ctrp_ccl <- unique(select(ctrp_data, DepMap_ID, CCLE_Name, accession_id))
gdsc_ccl <- unique(select(gdsc_data, DepMap_ID, CCLE_Name, accession_id))
# all_ccl <- merge(ccle_ccl, ctrp_ccl, by = c("DepMap_ID", "CCLE_Name", "accession_id"), all = TRUE)
```

# OLD: BOXPLOTS

--------------------------------------------------------------------------------

Mutation status boxplots for genes identified in G2P.

## G2P genes

Boxplots ordered by gene based on frequency of G2P associations.

### CRISPR

```{r fig.height=5, fig.width=20}
crispr_data_g2p$Hugo_Symbol <- factor(crispr_data_g2p$Hugo_Symbol, levels = g2p_genes)
crispr_data_g2p_color <- as.character(crispr_data_g2p$Color_Nonsilent)
names(crispr_data_g2p_color) <- crispr_data_g2p$Mutation_Status_Nonsilent
crispr_label_text <- data.frame(p.signif.adj = crispr_signif_g2p_gene$p.signif.adj, p.signif = crispr_signif_g2p_gene$p.signif, Hugo_Symbol = crispr_signif_g2p_gene$Hugo_Symbol)
crispr_label_text <- filter(crispr_label_text, Hugo_Symbol %in% g2p_genes)
crispr_label_text$Hugo_Symbol <- factor(crispr_label_text$Hugo_Symbol, levels = g2p_genes)
crispr_data_g2p_plot <- ggplot(data = crispr_data_g2p, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, nrow = 1) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85)) +
  scale_fill_manual(values = crispr_data_g2p_color) +
  guides(color = FALSE) +
  geom_text(data = crispr_label_text, mapping = aes(x = 1.5, y = 2, label = p.signif), nudge_y = 0.1) +
  theme(strip.text.x = element_text(size = 10, angle = 45), legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(title = "CRISPR", fill = "Mutation Status", y = "CERES Score", x = "Mutation Status")
crispr_data_g2p_plot

# ggsave("./plots/manuscript/crispr_data_g2p_plot.png", crispr_data_g2p_plot, device = "png", dpi = 450, width = 20, height = 5, units = "in")
```

### CCLE

Plot code:
```{r, eval=FALSE}
ccle_drugs <- as.character(unique(ccle_data_ptmuts$Drug))
ccle_g2p_drugboxplots <- lapply(ccle_drugs, makeDrugBoxplots, dataset = "CCLE")
ccle_g2p_drugboxplots_paths <- paste0(ccle_drugs, "_ccle_drugboxplots.png")
pwalk(list(ccle_g2p_drugboxplots_paths, ccle_g2p_drugboxplots), ggsave, path = "./plots/manuscript/ccle_g2p_drugboxplots", dpi = 450, width = 20, height = 5, units = "in")
```

Data management:
```{r, eval=FALSE}
ccle_g2p_drugboxplots <- paste0(list.files("./plots/manuscript/ccle_g2p_drugboxplots", full.names = TRUE))
names(ccle_g2p_drugboxplots) <- str_replace_all(ccle_g2p_drugboxplots, c("_ccle_drugboxplots.png" = "", "./plots/manuscript/ccle_g2p_drugboxplots/" = ""))
bsselect(ccle_g2p_drugboxplots, type = "img", live_search = TRUE, show_tick = TRUE, height = 100)
```

### CTRP

Plot code:
```{r, eval=FALSE}
ctrp_drugs <- as.character(unique(ctrp_data_ptmuts$Drug))
ctrp_g2p_drugboxplots <- lapply(ctrp_drugs, makeDrugBoxplots, dataset = "ctrp")
ctrp_g2p_drugboxplots_paths <- paste0(ctrp_drugs, "_ctrp_drugboxplots.png")
pwalk(list(ctrp_g2p_drugboxplots_paths, ctrp_g2p_drugboxplots), ggsave, path = "./plots/manuscript/ctrp_g2p_drugboxplots", dpi = 450, width = 20, height = 5, units = "in")
```

Data management:
```{r, eval=FALSE}
ctrp_g2p_drugboxplots <- paste0(list.files("./plots/manuscript/ctrp_g2p_drugboxplots", full.names = TRUE))
names(ctrp_g2p_drugboxplots) <- str_replace_all(ctrp_g2p_drugboxplots, c("_ctrp_drugboxplots.png" = "", "./plots/manuscript/ctrp_g2p_drugboxplots/" = ""))
bsselect(ctrp_g2p_drugboxplots, type = "img", live_search = TRUE, show_tick = TRUE, height = 100)
```

### GDSC

Plot code:
```{r, eval=FALSE}
gdsc_drugs <- as.character(unique(gdsc_data_ptmuts$Drug))
gdsc_g2p_drugboxplots <- lapply(gdsc_drugs, makeDrugBoxplots, dataset = "gdsc")
gdsc_g2p_drugboxplots_paths <- paste0(gdsc_drugs, "_gdsc_drugboxplots.png")
pwalk(list(gdsc_g2p_drugboxplots_paths, gdsc_g2p_drugboxplots), ggsave, path = "./plots/manuscript/gdsc_g2p_drugboxplots", dpi = 450, width = 20, height = 5, units = "in")
```

Data management:
```{r, eval=FALSE}
gdsc_g2p_drugboxplots <- paste0(list.files("./plots/manuscript/gdsc_g2p_drugboxplots", full.names = TRUE))
names(gdsc_g2p_drugboxplots) <- str_replace_all(gdsc_g2p_drugboxplots, c("_gdsc_drugboxplots.png" = "", "./plots/manuscript/gdsc_g2p_drugboxplots/" = ""))
bsselect(gdsc_g2p_drugboxplots, type = "img", live_search = TRUE, show_tick = TRUE, height = 100)
```

## G2P level A drug-gene associations only

### CCLE

```{r fig.height=4, fig.width=12}
ccle_g2p_order <- as.character(ccle_signif_g2p$Drug_Gene)
ccle_data_g2p$Drug_Gene_Ordered <- factor(ccle_data_g2p$Drug_Gene, levels = ccle_g2p_order, labels = sapply(strsplit(x = ccle_g2p_order, split = "_"), function(x) paste0(x[1], "\n", x[2])))
ccle_signif_g2p$Drug_Gene_Ordered <- factor(ccle_signif_g2p$Drug_Gene, levels = ccle_g2p_order, labels = sapply(strsplit(x = ccle_g2p_order, split = "_"), function(x) paste0(x[1], "\n", x[2])))
ccle_data_g2p_color <- as.character(ccle_data_g2p$Color_Nonsilent)
names(ccle_data_g2p_color) <- ccle_data_g2p$Mutation_Status_Nonsilent
ccle_label_text <- data.frame(p.signif.adj = ccle_signif_g2p$p.signif.adj, p.signif = ccle_signif_g2p$p.signif, Drug_Gene_Ordered = ccle_signif_g2p$Drug_Gene_Ordered)
ccle_data_g2p_plot <- ggplot(data = ccle_data_g2p, aes(x = Mutation_Status_Nonsilent, y = AUC)) +
  facet_wrap(~ Drug_Gene_Ordered, drop = FALSE, nrow = 1) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85), outlier.shape = 3, outlier.size = 0.5) +
  scale_fill_manual(values = ccle_data_g2p_color) +
  guides(color = FALSE) +
  geom_text(data = ccle_label_text, mapping = aes(x = 1.5, y = 7, label = p.signif), nudge_y = 0.1) +
  theme(legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(title = "CCLE", fill = "Mutation Status", y = "AUC", x = "Mutation Status")
ccle_data_g2p_plot

# ggsave("./plots/manuscript/ccle_data_g2p_plot.png", ccle_data_g2p_plot, device = "png", dpi = 450, width = 12, height = 4, units = "in")
```

### CTRP

```{r fig.height=7, fig.width=12}
ctrp_g2p_order <- as.character(ctrp_signif_g2p$Drug_Gene)
ctrp_data_g2p$Drug_Gene_Ordered <- factor(ctrp_data_g2p$Drug_Gene, levels = ctrp_g2p_order, labels = sapply(strsplit(x = ctrp_g2p_order, split = "_"), function(x) paste0(x[1], "\n", x[2])))
ctrp_signif_g2p$Drug_Gene_Ordered <- factor(ctrp_signif_g2p$Drug_Gene, levels = ctrp_g2p_order, labels = sapply(strsplit(x = ctrp_g2p_order, split = "_"), function(x) paste0(x[1], "\n", x[2])))
ctrp_data_g2p_color <- as.character(ctrp_data_g2p$Color_Nonsilent)
names(ctrp_data_g2p_color) <- ctrp_data_g2p$Mutation_Status_Nonsilent
ctrp_label_text <- data.frame(p.signif.adj = ctrp_signif_g2p$p.signif.adj, p.signif = ctrp_signif_g2p$p.signif, Drug_Gene_Ordered = ctrp_signif_g2p$Drug_Gene_Ordered)
ctrp_data_g2p_plot <- ggplot(data = ctrp_data_g2p, aes(x = Mutation_Status_Nonsilent, y = AUC)) +
  facet_wrap(~ Drug_Gene_Ordered, drop = FALSE, nrow = 3) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85), outlier.shape = 3, outlier.size = 0.5) +
  scale_fill_manual(values = ctrp_data_g2p_color) +
  guides(color = FALSE) +
  geom_text(data = ctrp_label_text, mapping = aes(x = 1.5, y = 22.5, label = p.signif), nudge_y = 0.1) +
  theme(legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(title = "CTRP", fill = "Mutation Status", y = "AUC", x = "Mutation Status")
ctrp_data_g2p_plot

# ggsave("./plots/manuscript/ctrp_data_g2p_plot.png", ctrp_data_g2p_plot, device = "png", dpi = 450, width = 12, height = 7, units = "in")
```

### GDSC

```{r fig.height=7, fig.width=12}
gdsc_g2p_order <- as.character(gdsc_signif_g2p$Drug_Gene)
gdsc_data_g2p$Drug_Gene_Ordered <- factor(gdsc_data_g2p$Drug_Gene, levels = gdsc_g2p_order, labels = sapply(strsplit(x = gdsc_g2p_order, split = "_"), function(x) paste0(x[1], "\n", x[2])))
gdsc_signif_g2p$Drug_Gene_Ordered <- factor(gdsc_signif_g2p$Drug_Gene, levels = gdsc_g2p_order, labels = sapply(strsplit(x = gdsc_g2p_order, split = "_"), function(x) paste0(x[1], "\n", x[2])))
gdsc_data_g2p_color <- as.character(gdsc_data_g2p$Color_Nonsilent)
names(gdsc_data_g2p_color) <- gdsc_data_g2p$Mutation_Status_Nonsilent
gdsc_label_text <- data.frame(p.signif.adj = gdsc_signif_g2p$p.signif.adj, p.signif = gdsc_signif_g2p$p.signif, Drug_Gene_Ordered = gdsc_signif_g2p$Drug_Gene_Ordered)
gdsc_data_g2p_plot <- ggplot(data = gdsc_data_g2p, aes(x = Mutation_Status_Nonsilent, y = AUC)) +
  facet_wrap(~ Drug_Gene_Ordered, drop = FALSE, nrow = 3) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85), outlier.shape = 3, outlier.size = 0.5) +
  scale_fill_manual(values = gdsc_data_g2p_color) +
  guides(color = FALSE) +
  geom_text(data = gdsc_label_text, mapping = aes(x = 1.5, y = 1, label = p.signif), nudge_y = 0.1) +
  theme(legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(title = "GDSC", fill = "Mutation Status", y = "AUC", x = "Mutation Status")
gdsc_data_g2p_plot

# ggsave("./plots/manuscript/gdsc_data_g2p_plot.png", gdsc_data_g2p_plot, device = "png", dpi = 450, width = 12, height = 7, units = "in")
```

# OLD: CRISPR POINT MUT MAPPING

--------------------------------------------------------------------------------

Match specific point mutations.
```{r fig.height=5, fig.width=20, eval=FALSE}
g2p_indications <- filter(read.delim("./data_munging/data_mutation_associations_appended.csv", sep = "\t", header = TRUE), Evidence.Level == "A")
maf_g2p_indications <- filter(maf_raw, Genome_Change %in% g2p_indications$MutationName)
crispr_g2p_indications <- merge(crispr_data_ptmuts, maf_g2p_indications, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
crispr_g2p_indications$Mutation_Status_Nonsilent <- ifelse(is.na(crispr_g2p_indications$Mutation_Status_Nonsilent), "Wildtype", crispr_g2p_indications$Mutation_Status_Nonsilent)
dup_g2p_indications <- filter(crispr_g2p_indications[, c("Hugo_Symbol", "CCLE_Name", "DepMap_ID")] %>% group_by(Hugo_Symbol, CCLE_Name, DepMap_ID) %>% tally(), n > 1)
crispr_g2p_indications <- merge(crispr_g2p_indications, dup_g2p_indications, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
crispr_g2p_indications$Genome_Change <- factor(crispr_g2p_indications$Genome_Change, levels = unique(crispr_g2p_indications$Genome_Change))
crispr_g2p_indications_nonNA <- filter(crispr_g2p_indications, is.na(n))
crispr_g2p_indications_plot <- ggplot(crispr_g2p_indications, aes(x = Hugo_Symbol, y = Score)) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(alpha = 0.5, color = "lightgray", outlier.shape = NA) +
  geom_jitter(width = 0.3, mapping = aes(color = Mutation_Status_Nonsilent), size = 1) +
  theme_light() +
  theme(legend.title = element_blank()) +
  labs(x = "Gene", y = "CERES Score")
crispr_g2p_indications_plot
crispr_g2p_indications_muts_plot <- ggplot(crispr_g2p_indications, aes(x = Protein_Change, y = Score)) +
  facet_grid(~ Hugo_Symbol, scales = "free_x", space = "free") +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(alpha = 0.5, color = "lightgray", outlier.shape = NA) +
  geom_jitter(width = 0.1, mapping = aes(color = Mutation_Status_Nonsilent), size = 1) +
  theme_light() +
  theme(legend.title = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Gene", y = "CERES Score")
crispr_g2p_indications_muts_plot

# ggsave("./plots/manuscript/crispr_g2p_indications_plot_test.png", crispr_g2p_indications_plot, width = 12, height = 5, units = "in")
# ggsave("./plots/manuscript/crispr_g2p_indications_muts_plot_test.png", crispr_g2p_indications_muts_plot, width = 20, height = 5, units = "in")
```

# REFERENCES

--------------------------------------------------------------------------------

Barretina, J., Caponigro, G., Stransky, N., Venkatesan, K., Margolin, A. A., Kim, S., … Garraway, L. A. (2012). The Cancer Cell Line Encyclopedia enables predictive modelling of anticancer drug sensitivity. Nature, 483(7391), 603–607. https://doi.org/10.1038/nature11003

Broad Institute Cancer Dependency Map; Cancer Data Science (2018): Cancer Dependency Map, CRISPR Avana dataset 18Q3 (Avana_public_18Q3). figshare. Fileset. doi:10.6084/m9.figshare.6931364.v1

Consortium, T. C. C. L. E., & Consortium, T. G. of D. S. in C. (2015). Pharmacogenomic agreement between two cancer cell line data sets. Nature, 528(7580), 84–87. https://doi.org/10.1038/nature15736

Data Science, Cancer (2018): DEMETER2 data. figshare. Fileset. doi:10.6084/m9.figshare.6025238.v2

Doench, J. G., Fusi, N., Sullender, M., Hegde, M., Vaimberg, E. W., Donovan, K. F., … Root, D. E. (2016). Optimized sgRNA design to maximize activity and minimize off-target effects of CRISPR-Cas9. Nature Biotechnology, 34(2), 184–191. https://doi.org/10.1038/nbt.3437

Meyers, R. M., Bryan, J. G., McFarland, J. M., Weir, B. A., Sizemore, A. E., Xu, H., … Tsherniak, A. (2017). Computational correction of copy-number effect improves specificity of CRISPR-Cas9 essentiality screens in cancer cells. Nature Genetics, 49(12), 1779–1784. https://doi.org/10.1038/ng.3984

McFarland, J. M., Ho, Z. V., Kugener, G., Dempster, J. M., Montgomery, P. G., Bryan, J. G., … Tsherniak, A. (2018). Improved estimation of cancer dependencies from large-scale RNAi screens using model-based normalization and data integration. https://doi.org/10.1101/305656

# SESSION INFORMATION

--------------------------------------------------------------------------------

```{r}
print(sessionInfo())
```